### 2014-04-15 paper.tex
initial commit

\documentclass[conference]{IEEEtran}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

\addtolength{\parskip}{-0.06in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{Cluster'14,} {September 17, 2014, Madrid, Spain}
%\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{An Innovative Storage Stack Addressing the Requirements of Extreme Scale Platforms}

\author{
\IEEEauthorblockN{Jay Lofstead}
\IEEEauthorblockA{Sandia National Laboratories\\gflofst@sandia.gov}
\and
\IEEEauthorblockN{Ivo Jimenez}
\IEEEauthorblockA{University of California, Santa Cruz\\ivo@cs.ucsc.edu}
\and
\IEEEauthorblockN{Carlos Maltzahn}
\IEEEauthorblockA{University of California, Santa Cruz\\carlosm@soe.ucsc.edu}
\and
\IEEEauthorblockN{Quincey Koziol}
\IEEEauthorblockA{HDF Group\\koziol@hdfgroup.org}
\and
\IEEEauthorblockN{John Bent}
\IEEEauthorblockA{EMC\\john.bent@emc.com}
\and
\IEEEauthorblockN{Eric Barton}
\IEEEauthorblockA{Intel\\eric.barton@intel.com}
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper not only provides a timely
summary of important aspects of the design specifications but also capture the
underlying reasoning that is not available elsewhere.

While many of the design elements are best practices from current research,
there are many new contributions as well as different integration approaches
offering different performance and quality of service guarantees. The specifics
of how the different layers interact offer a new approach for managing
scientific data for extreme scale platforms.

The integration of the HDF5 interface as the user interface gives a familiar
way to imteract with all of the new functionality while largely insulating 
end-users from the intracisies of dealing with the advanced functionality
offered by the stack.

The optional components aid in the scalability and simplifying implementation
across a wide range of platforms.

The data storage layer offers support both for local machine and data center
wide storage arrays as part of the design.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

\section{Introduction}

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, a joint effort between the US Department of Energy's
Office of Advanced Simulation and Computing and Advanced Scientific Computing
Research commissioned an effort to develop a design and prototype for an IO
stack suitable for the extreme scale environment. This is a joint effort led by
Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL. Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase recently completed with a second phase
currently getting underway as of early 2014. The core focus of the first phase
was basic functionality and design. Overriding many of the decisions during
this and any subsequent phases is the reality of budgets. Placing GBs of NVRAM
on every node, while a potentially advantageous approach, is not financially
feasible. With this in mind, the second phase will refine this design
incorporating fault recovery and other features missing from the first phase.

The overall design seeks to offer high availability, byte-granular,
multi-version concurrency control. Through the use of a copy-on-write style
mechanism, multiple versions of an object can be stored in potentially greatly
reduced space. It assumes the client interface will be through an IO library
affording a more complicated interface that offers richer functionality
requiring only minimal end-user code changes.  Managing most data access in a
platform-local layer rather than requiring writing to centralized storage will
better support the performance and energy requirements of extreme scale
integrated application workflows. At a more detailed view, the various layers
of the IO stack each contribute different functionality and performance
implications.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library~\cite{hdf5}. The intent is
to only have access to the storage stack through such an API to manage the
complexity of working with the lower layers. Below the user API is an IO
forwarding layer that redirects IO calls from the compute nodes to the IO
dispatching layer.  This IO forwarding layer is analogous to the function of
the IO nodes in a BlueGene machine. The next two layers have considerable
functionality. The IO dispatcher (IOD) serves as the primary storage interface
for the IO stack and is the only way to reach the persistent storage array in
lower layers. The Distributed Application Object Storage (DAOS) layer serves as
the persistent storage interface and is intended to be the foundation on which
everything else is built with no dependence on any technologies specified above
it. For example, the IOD layer and burst buffers are not required for DAOS to
operate properly. At the bottom is the Versioning Object Storage Device (VOSD).
It serves as the interface for storing objects of all types efficiently.

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Most of the key
features explored in this paper all have a strong presence in the IOD layer
motivating the focus of this examination.

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as data rearrangement, can be moved to the IOD layer
offloading the communication load from the compute nodes. IOD has three main
purposes. First, if the optional burst buffer is available, it works as a fast
cache absorbing write operations for the slower trickle out to the central
storage array. It can also be used to retrieve objects from the central storage
array for more efficient read operations and offers data filtering to make
client reads more efficient.  Second, it offers the transaction mechanism for
controlling data set visibility and to manage faults that would prevent a data
set from being used. Third, data processing operations can be placed in the
IOD. These operations are intended to offer data rearrangement, filtering, and
similar operations prior to data reaching the central storage array.

While these ideas are not necessarily new, they are new twists on best of class
efforts for these technologies. For example, offloading the collective
two-phase data sieving from the compute nodes to reorganize data has proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
Beyond these broad items, there are many important details some of which are
examined in more detail below.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product. These insights were gained based on
personal conversations with the team
members~\cite{JohnBent,QuinceyKoziol,EricBarton} discussing some of the
potential challenges with the design as written. These ideas are presented to
give a fuller picture of where the project is going rather than dwelling on the
limitations of the published documents.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incorporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:summary} discusses the system
overall with recommendations on what design elements should be considered based
on broader issues with current HPC data centers. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Burst Buffers}
\label{sec:burst}

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated as part of the IO
stack~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
The current Fast Forward IOD design recommends incorporating SSDs, but
specifically lists these devices as optional. Unfortunately, not incorporating
burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
the IOD design currently is written assuming burst buffers. This means that the
bulk of IO operations will only hit the IOD layer and proposed functionality,
such as the function shipping, do not discuss the impact on the DAOS layer or
the operators themselves should a burst buffer not be available.  Consider the
important functionality of data rearrangement and doing things like changing
the fast array dimension on shared, spinning media.  The design of DAOS assumes
that it will only be involved when persisting a completed transaction and only
for a fraction of the total transactions created.  Transaction frequency at the
IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
is the process of transferring the set of copy-on-write changes from
transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
mirrors the copy-on-write functionality, but stores the differences between
epochs rather than individual transactions. The primary difference is that at
the DAOS layer, data is stored in large, logically contiguous chunks to manage
the metadata load and hopefully improve read performance.

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from PreDatA. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the written IOD design, it describes a fixed-sized staging area
that is partitioned on a per-application basis.  This is unlikely to be useful
because of the limited compute and communication capacity to spare to perform
these operations at a bottleneck in the IO path.  The use of a separate data
staging area intentionally separate from the IO path allows using operators on
limited resources leaving the IO path clear for strictly data movement. A
nuance of this design is discussed in the Design Philosophy below.

By concentrating the Burst Buffers and function shipping into the storage
stack, three problems arise.  First, the amount of network bandwidth, IO
bandwidth, and compute power consumed for example operations from a single
application is likely to completely monopolize the IOD processes. Second, if
space and time partitioning is used instead, the functionality risks being too
small to be useful. Third, the long-term hardware performance advantage for
SSDs is questionable. Recent studies have shown that the erase-before-write
and interference between reading and writing with flash-based SSDs can cause
severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
of an optional use SSD layer in the new Trinity machine at Los Alamos will
offer a test bed to determine how likely these observed problems would affect a
production extreme scale platform.

%Current NAND-based flash devices top out at around 400 MB/sec.
%The key spec that is missing from this number is that 400 MB/sec is a measure
%of the fixed number of available IOPS multiplied by the block size. This
%represents the ideal streaming performance possible. The problem is that it
%costs an IOP to read 1 byte or 1 block (4KB or 8KB, depending on the device).
%It costs 1 IOP to write a full block--usually. In some cases, it will cost 2
%IOPs.  This accounts for the required pre-erase write prior to writing to a
%reused block. In the worst case, it can be 3 IOPS per write. This would be for
%a partial block write (read the old block, erase, write the modified block).
%One-third of 400 MB/sec, about 133 MB/sec, is well below the streaming
%performance of HDDs.  Granted, there is still rotational and seek latency to
%deal with for HDDs, but the advantage for SSDs has evaporated and potentially
%turned into a penalty at a considerable cost premium.  There are faster SSD
%solutions on the market that incorporate DRAM for caching and using the PCIe
%bus, for example, but their price precludes them from use in an extreme scale
%platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimately relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

\subsection{Design Philosophy}
The burst buffers design, as presented in the IOD documents, limits the
placement of the function operators and SSD buffers to the IO nodes. The team
does acknowledge the limitations of this design and intend to ultimately focus
on spreading the IOD layer from the IO nodes into the compute area as well.
This is intended both to help address the limitations of the IO bandwidth and
compute capability of these few nodes for data processing, but also to take
advantage of new layers in the storage hierarchy. By incorporating NVRAM into
compute nodes, new options for buffering data prior to being moved to
centralized storage become available and addresses some of the concerns about
SSD performance. For example, including a small amount of Phase Change memory
into many or most compute nodes offers a way to move data outside of both the
compute and IO path for data and communication intensive operations. Other 
projects~\cite{zheng:2010:predata} have suggested this will have value, but the
cost will have to be considered as part of the overall platform budget. This
lessens the impact of some operators while offering additional options for
places to store data.

Burst buffers being optional is a high level goal, but not one considered at a
detailed level within the design. For example, if there is no burst buffer, all
of the advanced functionality proposed for the IOD layer would have to work
against the DAOS layer instead. For example, function shipping assumes it will
operate on fast, local data within the IOD layer rather than against the
globally shared DAOS layer. With the additional desire to support using compute
node resources for these operations, serious work will be required to make a
fully functional end-to-end IOD layer implementation for a production system.

Another concern that is acknowledged, but no thought has been applied to, is
the requirement that a single IOD process of the set assigned to an application
is the master for any operation.  Should the number of concurrent applications
exceed the available nodes, sharing an IOD process will be required. The
requirements both in terms of scheduling and resource management were cut from
the project due to funding limitations. Since partitioning of the IOD processes
for exclusive use by particular applications is the assumed operating mode,
should insufficient IOD resources be available, either a job could be delayed
or IOD resources could be reallocated from a different process could be
redeployed for use by the new job. Handling resilience concerns for the IOD
processes must also be address.  These sorts of considerations still need to be
made for a full production system.

\section{Transactions and Epochs}
\label{sec:transactions}

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not a set of distributed,
asynchronous modifications across a set of related objects is complete or not.
It is also used to control access by treating the transaction ID of committed
transaction as a version identifier.  At the DAOS layer, they are called epochs
and represent persisted (durable) transactions from the IOD layer. Each of
these offers different functionality, but are connected as is explained below.
How these differ from the \DDT approach is also explored.  While IOD's and
\DDTns's transactions are seemingly very different, they use a similar
high-level design, but very different implementation, to solve the same
problem.

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container provides the single access context through which to access a
collection of objects. Transactions are the way that a series of modifications
to the objects within a container are treated atomically. Conceptually,
containers corresponds to a something akin to an HDF5 file in a traditional
file system. The objects in each container represent different data within a
file.  The three initially defined object types are key-value stores,
multi-dimensional arrays, and blobs.  The easiest way to understand these types
is to evaluate these from the perspective of an HDF-5 file, the initial user
interface layer. The key-value store represents a collection of attributes or
groups. The array represents a potentially multi-dimensional array.  The blob
represents a byte stream of arbitrary contents.  The fundamental difference
between an array and a blob is that the array has metadata specifying the
dimension(s). At the physical layer within the IO nodes, all of these objects
may be striped across multiple IO nodes.  Given this context, the transactions
come in two forms.

First is a single leader transaction where the IOD manages based on calls from
a single client. The underlying assumption is that the client side will manage
the transactional operations itself and the single client is capable of
reporting to the IOD how to evolve the transaction state.

The second form is called multi-leader and has the IOD layer manage the
transactions. In this case, when the transaction is created, a count of clients
is provided to the IOD layer. As clients commit their changes to the container,
the reference count is reduced. Once the count reaches 0, the transaction is
automatically committed.

\subsubsection{Design Philosophy}
Undocumented, but inherent in the design of these transactions is how faults
are detected. The initial design assumes the current Lustre fault detection
mechanism that can determine if a process or node is no longer reachable. This
detection happens at the DAOS layer and when a fault is detected, the rollback
process is pushed up to the IOD layer for all non-persisted or non-committed
transactions. This defines how a fault will be detected and what will trigger a
passive fault recovery (i.e., transaction abort).

There are two steps for beginning a transaction on a container. The first step
is for one or more process to open the container. This handle can be shared
eliminating the need for every participating process to hit the IOD layer to
open the file. The second step is a call to determine how many leaders will
participate in the transaction. In the single leader case, there is no
aggregation of success/fail statuses to determine the final transaction state.
Instead, it is assumed that the client will fully manage the transaction. In
the multi-leader model, some subset from 2 to $n$ where $n$ is the count of all
processes, declare themselves a leader for this container operation to the IOD
layer. Any number of processes can participate in modifying container without
regard to whether or not they are a leader. Once each leader has finished, with
the assumption that any clients they may be responsible for are finished as
well, the IOD layer aggregates those responses to either commit or abort the
transaction.

Ultimately, with the passive detection of faults for transaction leaders, the
transaction mechanism can work very well. A mostly unstated restriction that is
being relaxed is that every sequential transaction on a container is considered
dependent on the earlier transaction. Should one output be delayed and the
subsequent five succeed, when the delayed process finally fails, all six
transactions are rolled back. The thought of using this mechanism to store
subsequent checkpoint outputs in the same container to both save space, but not
care if one fails, cannot work in the current form. This has been acknowledged
and is being relaxed requiring a new parameter to the creation of a transaction
determining if it will be dependent or not.

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, when an IOD transaction is persisted to DAOS, the IOD transaction ID is
the used as the epoch ID. The key difference is that at the DAOS layer, some
transaction (epoch) IDs will not be represented since not all IOD transactions
are necessarily persisted.

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
This is a hybrid approach to metadata management that is half-way between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

Eliminating metadata as a core component of a file system is not new. It has
been explored as part of the Light Weight File Systems
project~\cite{oldfield:lwfs}. In LWFS, the metadata service is explicitly
limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. Using this hybrid approach is less common~\cite{weil:2006:ceph} and
introduces other issues.

IOD and DAOS both share a philosophy that they will have to maintain the
metadata about how the physical pieces of the logical objects are striped and
where they are placed. The primary metadata management is done at the DAOS
layer with the IOD layer relying on the DAOS layer for all authoritative
information about containers and objects. The only place where the IOD layer
manages metadata for itself is to manage how the different objects are striped
across the IO nodes.

\subsubsection{Design Philosophy}
While the metadata design is not fully defined, there are a few things that
are intended. For example, there is a standard, well-known container that is
the system metadata. This includes the list of all other containers. This
container is treated like any other data in the system and striped as
appropriate. Unfortunately, this still couples the metadata to a single object
that must serialize access. If the metadata, including information about
striping and other data layout operations were separated completely from the
data path, more scalable throughput could be achieved. The real challenge of
this is introduced by the IOD, DAOS, and VOSD layers collectively. Each of
these requires some different metadata storage and the migration is transparent
to the user.  Supporting fully independent metadata with this model is
difficult. Serious thought on how to do this effectively outside the data path
should be considered for phase two.

Based on the lessons from the \DDT metadata
service~\cite{lofstead:2012:txn-metadata} construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services.  Users can
manage everything by maintaining the metadata including the list of objects
themselves. However, there are drawbacks to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks are introduced.  First, a client could
create a dangling entry in the metadata service that does not correspond to any
objects in the object store. Second, a client could create orphaned objects
that have no associated metadata entries. Third, updates to the metadata or
object store service should be an atomic operation, but due to a lack of
coordination, a window where the system is inconsistent appears.

Ultimately, the consistency semantics required must be determined. If a
metadata service is required and it must be in sync with the object storage
service, then additional work must be performed. In traditional file systems,
the metadata and object storage updates are atomic. With decoupling metadata
from object storage, should this atomicity still be desired, it requires both
the ability for the services to participate in a task that is part of a larger
atomic operation and a higher-level mechanism to manage the atomic operation.

Overall, while additional work is required to maintain a client-side only
metadata service, it eliminates any potential bottlenecks related to updating
metadata related to the object storage. The burden of tracking striping and
other metadata that has traditionally been part of the metadata associated with
the file system will have to be maintained by the object storage service. The
lack of a centralized, serialized bottleneck to store that information improves
concurrency.

\subsection{Comparison to \DDTns}
The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
approach for handling ACID-style transactions in an environment with parallel
clients and multiple servers (doubly distributed). Rather than being aimed
solely at data movement operations, \DDT seeks to address the general problem
of managing any operation with multiple clients and servers.  Consider the
management of the analysis/visualization area, potentially similar to the IOD
concept. The transaction protocol is used to help manage resizing of the
resource allocation to the various analysis and visualization components.  For
the purposes of this discussion, \DDT could also be used to manage changing how
IOD processes and/or nodes are used without exposing these changes to the
client processes prematurely.  This has been described and analyzed
previously~\cite{dayal:2013:io-containers}.

The example metadata and data storage services created as part of the \DDT
project have no dependencies between transactions that prevent visibility
should an older version be incomplete. This additional, intentional requirement
by IOD offers different functionality than \DDTns's example services. In the
case of \DDTns, the functionality is more minimal, but also avoids some of the
concerns outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
%The
%breakdown of the number of participants in each role is shown in
%Table~\ref{tab:scaling}. For comparison, consider the Number of
%Sub-Coordinators equivalent to IOD processes. The Processes Per Sub-Coordinator
%represents the number of clients that use a particular IOD process. For these
%tests, we maintained a balanced distribution and always used at least two
%sub-coordinators to slow down the processing.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
%\vspace{-0.15in}
\end{figure}

%\begin{table}[ht]
%    \vspace{-0.15in}
%    \centering
%    \caption[Scaling Configuration]{Performance Tests Scaling Configuration}
%    \bigskip
%    \vspace{-0.15in}
%
%    \begin{tabular}{|r|r|r|}
%\hline
%Processes & \vtop{\hbox{\strut Number of}\hbox{\strut Sub-Coordinators}} & \vtop{\hbox{\strut Processes Per} \hbox{\strut Sub-Coordinator}}\\
%\hline
%%8 & 2 & 4 \\
%%16 & 2 & 8 \\
%%32 & 2 & 16 \\
%%64 & 2 & 32 \\
%%128 & 2 & 64 \\
%256 & 2 & 128 \\
%512 & 2 & 256 \\
%1024 & 4 & 256 \\
%2048 & 8 & 256 \\
%4096 & 16 & 256 \\
%8192 & 32 & 256 \\
%16384 & 64 & 256 \\
%32768 & 128 & 256 \\
%65536 & 256 & 256 \\
%\hline
%    \end{tabular}
%    \label{tab:scaling}
%\end{table}

At a high level, both \DDT and the IOD transactions have the same design. In
both cases, a hierarchical model is employed. In the case of \DDTns, it is a
purely client-side tree using semi-synchronous messaging. The messaging itself,
in the current implementation, uses asynchronous MPI messages. The synchronous
component comes from the timeout mechanism used to detect faults.  It forces a
level of coordination and synchronization for the protocol. For IOD, it is a
server-side tree and fully asynchronous relying on the existing Lustre fault
detection mechanism for failure detection. In both cases, there is a master in
charge of managing the transaction and a collection of workers that aggregate
into the master through second-level leaders. Beyond that, there are some
significant differences. Some of the different choices made by IOD raise some
possible concern.

The multi-leader model introduces the possibility of forcing a rollback of an
entire transaction when a partial retry might be sufficient for success. Since
the transactions are managed at a high level rather than the individual
tasks, a failure in a limited distributed task can cause the entire transaction
to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
10 have an additional shared task to complete. If the task shared by the 3
processes fails on any of the three, the entire transaction would roll back
because it is a coarse-grained success/failure. If a concept like
sub-transactions at a task granularity were used, then it would be possible for
the one process that failed to report just that failure. Then the transaction
manager could reassign the resources for these 3 processes and try just that
operation again. If it now succeeds, then the overall transaction can be marked
successful only redoing the minimum amount of work required.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
%\vspace{-0.25in}
\end{figure}

\DDT has addressed theses issues in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, whichever process takes over
that role knows to expect a completion message for that sub-transaction or the
overall transaction cannot complete. Global sub-transactions can be defined at
any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

The advantages of eliminating these messages is not performance as demonstrated
by the performance of \DDTns. Instead, it offers a much less synchronous model
that matches with different programming models, such as Charm++ or other
task-based approaches. Since it can work for the bulk-synchronous model also,
it is a more broadly applicable approach. This assumes that the observed
potential issues can be addressed successfully.

%\subsection{Comparison to Other Protocols}
%Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
%ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
%unsuitable for this environment. First, the distributed servers in Paxos
%systems are all distributed copies of each other that eventually become
%consistent. Given the scale we wish to address, a single node's memory is
%unlikely to be able to hold all of the data necessary for many operations at
%scale. They also do not have a guarantee for when consensus will be achieved
%without using slower synchronous calls. For the tight timing we wish to
%support, we need guarantees of when a consistent state has been achieved.
%Second, these systems also all assume that updates are initiated from a single
%client process rather than a parallel set of processes as is the standard in
%HPC environments.
%
%\DDT uses a second layer of coordination on the client side that greatly
%increases the scalability by consolidating messages from clients into unique
%sets prior to sending to the overall coordinator. A gossip
%protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
%purpose, but the delay of eventual consistency is strictly avoided with this
%protocol to ensure guarantees at particular states in the code. For example, if
%a failure occurs, the global understanding of the role of all processes is
%required in order for effective communication to occur for operations like
%creating sub-transactions or voting. In this case, the protocol can offer
%stronger statements about consistency than these protocols offer.  These
%features offer a way to easily scale the transaction protocol given the
%guarantees we wish to offer. The IOD approach does nothing to address these
%concerns.
%
%Another effort to offer consistency and data integrity for the ZFS file
%system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
%focus on the processes all having a notion of completion as a transaction, this
%work focuses on the integrity of the data movement operations. We view this
%work as something that should be considered hand-in-hand with a transaction
%approach to ensure the integrity of the movement of the data in addition to the
%agreement of processes about the successful completion of a parallel operation.

\section{Broader Design}
\label{sec:summary}

At a broader level, there are some concerns that were partially clarified
through conversations with the team.  Consider a shared file system across an
HPC data center. The current design maintains the metadata in its own
container. Since copying data from the IOD layer to the DAOS layer requires an
explicit persist call, how and when synchronizing the metadata across the
layers and potentially across machines occurs is undefined. Delaying
synchronization until an explicit persist is called will reduce the update
frequency, but delays the data visibility on other platforms. Ideally, the
metadata object would need to be automatically persisted every time a container
transaction is persisted to the DAOS layer.

The implication of this is that every transaction persist is double operation
to account for the metadata persist. More importantly, the IOD-layer version of
the metadata container may contain readable transactions that have not been
persisted to the DAOS layer. How to handle this inconsistency between the two
layers still needs to be explored.

A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of objects across a set of shards. For the IOD an object may be a
shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
Having the same names with locally correct, globally conflicting definitions
serves to confuse how the system should work.

\section{Conclusions}
\label{sec:conclusion}

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the consensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our understanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDTns, Lightweight File Systems, and
other efforts to explore the requirements for this space, along with the
analysis presented in this paper will prove useful for the next phase of the
Fast Forward project.

\section{Acknowledgements}
The authors would like to thank Eric Barton, John Bent, Gary Grider, and
Quincey Koziol for their early review comments and discussions that clarified
the details of the design, the intent of the design, and the future plans.

\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}

### 2014-04-15 paper.tex
Adding description of ACG and HDF5 layer

@@
 %\usepackage{subfig}
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage[pdftex]{}
-\usepackage{todonotes}
+\usepackage{graphicx}
+%\usepackage{todonotes}
 \usepackage{listings}
@@
 It serves as the interface for storing objects of all types efficiently.
+
+%%%%%%%%%%%%%%%%%%%%%%%%% <<<<<<<<<<<<<<<<
+
+The ``I/O nodes'' handle 
+requests forwarded by the scientific applications, integrate a tier of
+solid-state devices to absorb the burst of random operations, and
+organize/re-format the data so that transfers from the staging area to
+the traditional parallel file system can be done more efficiently. It
+also has the capacity of executing analysis on the fresh data that has
+just been generated by simulation applications running at the compute
+nodes. As the data arrives, re-organization and data preparation can be
+applied in order to anticipate the execution of analytical tasks
+{[}3{]}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-arch.png}
+\caption{Exascale Architecture. \label{exa-arch}}
+\end{figure}
+
+A common cluster configuration for this type of deployment is shown in
+Figure \ref{exa-arch}. The designated I/O nodes (IONs) are connected to
+the compute nodes (CNs) through the same fast fabric (eg. Infiniband),
+while the connection to the external storage cluster is through a
+secondary slower channel (eg. 10Gb ethernet). I/O staging thus
+introduces a new layer in the application-to-storage stack, as shown in
+Figure \ref{exa-stack} (left).
+
+In current proposals, the stack is managed by middleware that seats
+between the application and the parallel file system {[}2,4{]}. Since
+most of the applications assume a POSIX interface to storage {[}5{]},
+existing middleware either transparently handle the I/O operations,
+appearing as regular POSIX calls to applications; or modify the I/O API
+as little as possible in order to minimize the impact on production
+codes.
+
+As we move towards the exascale goal, this way of interfacing with
+storage becomes more of an obstacle. Many features provided by the stack
+are hidden by the middleware/POSIX layers. If these were visible to the
+applications, and if the applications were able to control how the
+underlying layers behave, they would benefit greatly since the domain
+knowledge that is available at the application level could be used to
+execute storage operations efficiently. One more implication of the
+25-year-old POSIX API is in cumbersome middleware, since metadata has to
+be part of the same uni-dimensional byte array being stored.
+
+The Fast Forward Storage I/O project {[}6{]} is aimed at merging the
+features of existing middleware into a next-generation parallel file
+system stack. This new storage API termed the I/O dispatcher (IOD) is
+object-based, transactional, asynchronous and I/O-staging-aware (Figure
+\ref{exa-stack} (right)).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-stack.png}
+\caption{Top-to-bottom view of the component stack. \label{exa-stack}}
+\end{figure}
+
+The prototype stack, from a top-down point of view, has the following
+components:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Arbitrarily Connected Graphs (ACG).
+\item
+  HDF5 extensions.
+\item
+  I/O Dispatcher (IOD).
+\item
+  Distributed Application Object Storage (DAOS).
+\item
+  Versioning Object Storage Device (VOSD).
+\end{enumerate}
+
+Applications running on the compute nodes will be written in Python
+scripts or in C++ using the HDF5 APIs. These applications will execute
+analysis on graph-based data (ACG). The point is to demonstrate both the
+HPC and BigData use cases of the Exascale architecture. The data
+structures will be stored in HDF5, for which new API extensions will be
+implemented. These new API calls will expose the transactional,
+asynchronous and function-forwarding semantics of the underlying stack.
+
+As mentioned before, the stack provides \emph{non-POSIX, object-based,
+transactional and asynchronous active-storage}, meaning that POSIX is
+supplanted by new object interfaces that reach up to the HDF5 layer
+(Application I/O in the stack diagram); transactional semantics are
+present, from HDF down to the VOSD layer (Storage layer); clients don't
+need to wait for any blocking operation; and analysis can be shipped and
+executed on I/O or DAOS nodes.
+
+We give a high-level description of each of layer.
+
+\subsection{ACG}\label{acg}
+
+This will exemplify how applications make use of the Exascale stack. For
+FF, support for
+GraphLab 
+and GraphBuilder will be prototyped, which are graph-processing
+frameworks. GraphBuilder is a set of MapReduce tasks that extract,
+normalize, partition and serialize a graph out of unstructured data, and
+writes graph-specific formats into HDFS. These files are later consumed
+by GraphLab, a vertex-centric, asynchronous execution engine that runs
+directly on top of HDFS (i.e. non-MapReduce). The following illustrates
+the architecture of both frameworks:
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{images/graphlab-and-graphbuilder.png}
+\caption{Figure 3. GraphLab and GraphBuilder stacks (taken from
+{[}4{]}).}
+\end{figure}
+
+In order to make both work on top of the exascale stack, both have to be
+modified. After these modifications are implemented, GraphBuilder will
+be able to write the partitioned graph in (the newly proposed) HDF5
+files which will thus be stored in the IOD nodes (or IONs) in a
+parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
+the library to perform at high speeds by benefiting from the new
+features (see next section). In general both frameworks will be modified
+so that calls to HDFS-based formats are replaced by the proposed HDF5
+ones. This is referred to as the HDF Adaptation Layer or HAL and will
+provide, from the GraphBuilder/GraphLab point of view {[}5{]}:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  capability for storing the newly proposed HDF5 format
+\item
+  association of network information to vertices/edges
+\item
+  shipping computation to the IONs
+\item
+  asynchronous vertex updates
+\item
+  efficient data sharing ammong CNs
+\item
+  computation over versioned datasets
+\end{itemize}
+
+\subsection{HDF5 extensions}\label{hdf5-extensions}
+
+The extensions done to HDF5 allows an application to take full advantage
+of the new exascale features. The additions comprise {[}6{]} (Figure 4):
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Object-storage API based on HDF5 to support high-level data models.
+  This exposes asynchronous, transactional semantics to the application,
+  as well as end-to-end data integrity. It will also allows the usage of
+  pointer data types, passing of hints down to the storage and support
+  for asynchronous index building, maintenance and querying.
+\item
+  Virtual Object Layer (VOL) plugin that translates HDF5 API requests
+  from applications to IOD API calls.
+\item
+  Function shipping from CN to IONs. This provides the application
+  developer with the capacity of sending computation down to the IONs
+  and get back results.
+\item
+  Analysis Shipping from CN to IONs or DAOS nodes. This is similar to 3
+  but instead of returning the result over the network, it gets stored
+  on the nodes and pointers to it are returned.
+\end{enumerate}
+
+In terms of the CN-ION communication model, a client/server architecture
+is implemented {[}8{]}: every ION runs an IOFSL (I/O Function Shipping
+Layer) server {[}9{]}; the IOFSL client is integrated into the HDF5
+library which runs on each CN. A client can forward requests to any
+number of IONs. Every I/O operation issued by HDF5 is asynchronously
+shipped to the IOFSL server and asynchronously executed.
+
+HDF5 {[}7{]} is a versatile data model containing complex data objects
+and metadata. Its information set is a collection of datasets, groups,
+datatypes and metadata objects. The data model defines mechanisms for
+creating associations between various information items. The main
+components of HDF5 are described below.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/hdf5.png}
+\caption{Sample file illustrating the hierarchical structure of HDF5
+files. \label{hdf5-sample}}
+\end{figure}
+
+\begin{itemize}
+\item
+  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
+  is represented by a file. It is a collection of objects that also
+  explains the relationship between them. Every file begins with a root
+  group ''/'', which serves as the ''starting-point'' in the object
+  hierarchy.
+\item
+  \textbf{Dataset}: HDF5 datasets are objects that represent actual data
+  or content. Datasets are arrays which can have multiple dimensions. A
+  dataset is characterized by a dataspace and a datatype. The dataspace
+  captures the rank (number of dimensions), and the current and maximum
+  extent in each dimension. The datatype describes the type of its data
+  elements.
+\item
+  \textbf{Group}: A group is an explicit association between HDF5
+  objects. It is synonymous with directories in a file system. A group
+  could contain multiple other groups, datasets or datatypes within it.
+\item
+  \textbf{Attribute}: Attributes are used for annotating datasets,
+  groups, and datatype objects. They are datasets themselves, and are
+  attached to existing objects they annotate.
+\end{itemize}
+
+For example, as shown in Figure \ref{hdf5-sample}, the file
+Sample.h5 contains the root group which itself contains a
+group G1 and two datasets, D1 and D2. Group
+G1 contains a dataset D3. Attribute A1 is
+linked to dataset D1. The objects and the relationships between
+them can be represented as a B-tree, which is used internally by HDF5 to
+index its objects.
+
+The Fast Forward project is working on the adaptation of the HDF5
+libraries in order to expose the storage stack to applications.
+
+\subsection{Virtual Object Layer}\label{virtual-object-layer}
+
+The Virtual Object Layer is an abstraction mechanism internal to the
+HDF5 library {[}\textbf{???}{]}. As shown in Figure \ref{vol-arch} it is
+implemented just below the public API. The VOL exports an interface that
+allows writing plugins for HDF5, thereby enabling developers to store
+objects in a format different from the default HDF5 file format. Plugin
+writers provide an implementation for a set of functions that access
+data on disk or on a parallel file system.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.20]{images/vol-arch.png}
+\caption{Architectural view of the VOL abstraction mechanism.
+\label{vol-arch}}
+\end{figure}
+
+\subsection{IOD VOL plugin}\label{iod-vol-plugin}
+
+The IOD VOL plugin serves as the bridge between HDF5 and the IOD API. It
+is also shown in Figure \ref{vol-arch}. The application calls the HDF5
+library while running on the system's compute nodes. Using the VOL
+architecture, the IOD VOL plugin uses a function shipper (RPC library)
+to forward the VOL calls to a server component running on the I/O nodes
+(IONs). At that point, the VOL calls are translated into I/O Dispatcher
+(IOD) API calls and executed at the IONs.
+
+The mapping between HDF5 objects and IOD objects is the following:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  file = container
+\item
+  dataset = array
+\item
+  group = key-value store
+\item
+  attribute = key-value store
+\end{itemize}
+
+\subsection{Fast Forward's I/O Dispatcher
+API}\label{fast-forwards-io-dispatcher-api}
+
+The Fast Forward I/O and Storage project is aimed at integrating many of
+the services existing currently in middleware libraries in a
+next-generation storage stack, exposed through a non-POSIX,
+transactional, asynchronous, object-based API.
+
+From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
+interface allows the user to:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  transactions
+\item
+  asynchrony
+\item
+  objects
+\item
+  layout
+\item
+  placement
+\item
+  formats
+\end{itemize}
+
+Type of objects:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  container
+\item
+  key-value store
+\item
+  blobs
+\item
+  multi-dimensional arrays
+\end{itemize}
+
+As mentioned previously, IOD allows to store distinct types of objects.
+From here on we will limit our discussion to multidimensional arrays
+(HDF5 datasets) and leave the treatment of other object types as future
+work. A multidimensional array can be sharded among the distinct I/O
+nodes in the staging area. The IOD API supports the following
+strategies:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{contiguous}. fixed chunking, distributed in a round-robin
+  fashion accross the I/O nodes.
+\item
+  \textbf{chunked}. same as above but with irregular (sparse) chunking.
+\item
+  \textbf{user-defined}. either contiguous or chunked but user specifies
+  where to place each individual shard.
+\end{itemize}
+
+It is possible to request the transformation of an object's physical
+layout to other formats, having multiple copies of the same objects in
+multiple formats if desired. Also, the user can pre-fetch objects from
+the storage cluster into the I/O nodes or read them directly from the
+storage cluster. At the semantic level (HDF5), indexes can be created
+for datasets, which results in being able to read through an index
+instead of directly from the base array.
+
+{[}sharding. \label{sharding}{]}{[}sharding{]}
+
+All these distinct alternatives result in having many different ways of
+executing the same analysis task. In the future, we plan to extend our
+design to incorporate an optimization phase that will prepare a task by
+taking into account the distinct possible ways of executing and select
+the most promising one. In our subsequent discussions, we consider only
+data-movement optimization, i.e.~sending the analysis code as close as
+possible to the data. In practice, this means we focus on identifying
+sharding of datasets and execute code accordingly over the appropriate
+shards.
+
+%%%%%%%%%%%%%%%%%%%%%%%%% >>>>>>>>>>>>>>
 

### 2014-04-17 paper.tex
Adding description of ACG and HDF5 layer
Fixed display of the author list and edited abstract

--- 2014-04-15 paper.tex
@@
 
-\title{An Innovative Storage Stack Addressing the Requirements of Extreme Scale Platforms}
+\title{An Innovative Storage Stack Addressing the Requirements of Extreme Scale Platforms and Big Data Applications}
 
 \author{
-\IEEEauthorblockN{Jay Lofstead}
-\IEEEauthorblockA{Sandia National Laboratories\\gflofst@sandia.gov}
-\and
-\IEEEauthorblockN{Ivo Jimenez}
-\IEEEauthorblockA{University of California, Santa Cruz\\ivo@cs.ucsc.edu}
-\and
-\IEEEauthorblockN{Carlos Maltzahn}
-\IEEEauthorblockA{University of California, Santa Cruz\\carlosm@soe.ucsc.edu}
-\and
-\IEEEauthorblockN{Quincey Koziol}
-\IEEEauthorblockA{HDF Group\\koziol@hdfgroup.org}
-\and
-\IEEEauthorblockN{John Bent}
-\IEEEauthorblockA{EMC\\john.bent@emc.com}
-\and
-\IEEEauthorblockN{Eric Barton}
-\IEEEauthorblockA{Intel\\eric.barton@intel.com}
+\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
+\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
+\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
 }
@@
 wide storage arrays as part of the design.
+
+In addition to these features supporting HPC platforms, the stack also addresses
+the need of big data applications as demonstrated through the Arbitrary Connected Graphs tool.
 

### 2014-04-17 paper.tex

--- 2014-04-15 paper.tex
@@
 \centering
-\includegraphics{images/graphlab-and-graphbuilder.png}
+\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
 \caption{Figure 3. GraphLab and GraphBuilder stacks (taken from

### 2014-04-17 paper.tex
fixed up the title and intro a bit. re-ordered the pieces Ivo added a bit to make it more sections by layer. More still needs to be done.

@@
 
-\title{An Innovative Storage Stack Addressing the Requirements of Extreme Scale Platforms and Big Data Applications}
+\title{An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications}
 
@@
 Current production HPC IO stack design is unlikely to offer sufficient features
-and performance to adequately serve the needs of an extreme scale platform. To
-address these limitations, a joint effort between the US Department of Energy's
-Office of Advanced Simulation and Computing and Advanced Scientific Computing
-Research commissioned an effort to develop a design and prototype for an IO
-stack suitable for the extreme scale environment. This is a joint effort led by
-Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
-leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
+and performance to adequately serve the needs of an extreme scale platform.
+Adding to the complexity of the problem is the variety of Big Data problems
+that are taking advantage of these platforms. Trying to address both of these
+domains is a daunting problem.
+
+To address these challenges, a joint effort between the US Department of
+Energy's Office of Advanced Simulation and Computing and Advanced Scientific
+Computing Research commissioned an effort to develop a design and prototype for
+an IO stack suitable for the extreme scale environment called the Fast Forward
+Storage and IO (FFSIO) project. This is a joint effort led by Lawrence
+Livermore National Laboratory, with the DOE Data Management Nexus leads Rob
+Ross and Gary Grider as coordinators and contract lead Mark Gary. The
 participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
@@
 
-The basic architecture incorporates five layers. The top layer is a high level
-IO library, such as the demonstration HDF-5 library~\cite{hdf5}. The intent is
-to only have access to the storage stack through such an API to manage the
-complexity of working with the lower layers. Below the user API is an IO
-forwarding layer that redirects IO calls from the compute nodes to the IO
-dispatching layer.  This IO forwarding layer is analogous to the function of
-the IO nodes in a BlueGene machine. The next two layers have considerable
-functionality. The IO dispatcher (IOD) serves as the primary storage interface
-for the IO stack and is the only way to reach the persistent storage array in
-lower layers. The Distributed Application Object Storage (DAOS) layer serves as
-the persistent storage interface and is intended to be the foundation on which
-everything else is built with no dependence on any technologies specified above
-it. For example, the IOD layer and burst buffers are not required for DAOS to
-operate properly. At the bottom is the Versioning Object Storage Device (VOSD).
-It serves as the interface for storing objects of all types efficiently.
-
-%%%%%%%%%%%%%%%%%%%%%%%%% <<<<<<<<<<<<<<<<
-
-The ``I/O nodes'' handle 
-requests forwarded by the scientific applications, integrate a tier of
-solid-state devices to absorb the burst of random operations, and
-organize/re-format the data so that transfers from the staging area to
-the traditional parallel file system can be done more efficiently. It
-also has the capacity of executing analysis on the fresh data that has
-just been generated by simulation applications running at the compute
-nodes. As the data arrives, re-organization and data preparation can be
-applied in order to anticipate the execution of analytical tasks
-{[}3{]}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/exa-arch.png}
-\caption{Exascale Architecture. \label{exa-arch}}
-\end{figure}
-
-A common cluster configuration for this type of deployment is shown in
-Figure \ref{exa-arch}. The designated I/O nodes (IONs) are connected to
-the compute nodes (CNs) through the same fast fabric (eg. Infiniband),
-while the connection to the external storage cluster is through a
-secondary slower channel (eg. 10Gb ethernet). I/O staging thus
-introduces a new layer in the application-to-storage stack, as shown in
-Figure \ref{exa-stack} (left).
-
-In current proposals, the stack is managed by middleware that seats
-between the application and the parallel file system {[}2,4{]}. Since
-most of the applications assume a POSIX interface to storage {[}5{]},
-existing middleware either transparently handle the I/O operations,
-appearing as regular POSIX calls to applications; or modify the I/O API
-as little as possible in order to minimize the impact on production
-codes.
-
-As we move towards the exascale goal, this way of interfacing with
-storage becomes more of an obstacle. Many features provided by the stack
-are hidden by the middleware/POSIX layers. If these were visible to the
-applications, and if the applications were able to control how the
-underlying layers behave, they would benefit greatly since the domain
-knowledge that is available at the application level could be used to
-execute storage operations efficiently. One more implication of the
-25-year-old POSIX API is in cumbersome middleware, since metadata has to
-be part of the same uni-dimensional byte array being stored.
-
-The Fast Forward Storage I/O project {[}6{]} is aimed at merging the
-features of existing middleware into a next-generation parallel file
-system stack. This new storage API termed the I/O dispatcher (IOD) is
-object-based, transactional, asynchronous and I/O-staging-aware (Figure
-\ref{exa-stack} (right)).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/exa-stack.png}
-\caption{Top-to-bottom view of the component stack. \label{exa-stack}}
-\end{figure}
-
-The prototype stack, from a top-down point of view, has the following
-components:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Arbitrarily Connected Graphs (ACG).
-\item
-  HDF5 extensions.
-\item
-  I/O Dispatcher (IOD).
-\item
-  Distributed Application Object Storage (DAOS).
-\item
-  Versioning Object Storage Device (VOSD).
-\end{enumerate}
-
-Applications running on the compute nodes will be written in Python
-scripts or in C++ using the HDF5 APIs. These applications will execute
-analysis on graph-based data (ACG). The point is to demonstrate both the
-HPC and BigData use cases of the Exascale architecture. The data
-structures will be stored in HDF5, for which new API extensions will be
-implemented. These new API calls will expose the transactional,
-asynchronous and function-forwarding semantics of the underlying stack.
-
-As mentioned before, the stack provides \emph{non-POSIX, object-based,
-transactional and asynchronous active-storage}, meaning that POSIX is
-supplanted by new object interfaces that reach up to the HDF5 layer
-(Application I/O in the stack diagram); transactional semantics are
-present, from HDF down to the VOSD layer (Storage layer); clients don't
-need to wait for any blocking operation; and analysis can be shipped and
-executed on I/O or DAOS nodes.
-
-We give a high-level description of each of layer.
-
-\subsection{ACG}\label{acg}
-
-This will exemplify how applications make use of the Exascale stack. For
-FF, support for
-GraphLab 
-and GraphBuilder will be prototyped, which are graph-processing
-frameworks. GraphBuilder is a set of MapReduce tasks that extract,
-normalize, partition and serialize a graph out of unstructured data, and
-writes graph-specific formats into HDFS. These files are later consumed
-by GraphLab, a vertex-centric, asynchronous execution engine that runs
-directly on top of HDFS (i.e. non-MapReduce). The following illustrates
-the architecture of both frameworks:
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
-\caption{Figure 3. GraphLab and GraphBuilder stacks (taken from
-{[}4{]}).}
-\end{figure}
-
-In order to make both work on top of the exascale stack, both have to be
-modified. After these modifications are implemented, GraphBuilder will
-be able to write the partitioned graph in (the newly proposed) HDF5
-files which will thus be stored in the IOD nodes (or IONs) in a
-parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
-the library to perform at high speeds by benefiting from the new
-features (see next section). In general both frameworks will be modified
-so that calls to HDFS-based formats are replaced by the proposed HDF5
-ones. This is referred to as the HDF Adaptation Layer or HAL and will
-provide, from the GraphBuilder/GraphLab point of view {[}5{]}:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  capability for storing the newly proposed HDF5 format
-\item
-  association of network information to vertices/edges
-\item
-  shipping computation to the IONs
-\item
-  asynchronous vertex updates
-\item
-  efficient data sharing ammong CNs
-\item
-  computation over versioned datasets
-\end{itemize}
-
-\subsection{HDF5 extensions}\label{hdf5-extensions}
-
-The extensions done to HDF5 allows an application to take full advantage
-of the new exascale features. The additions comprise {[}6{]} (Figure 4):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Object-storage API based on HDF5 to support high-level data models.
-  This exposes asynchronous, transactional semantics to the application,
-  as well as end-to-end data integrity. It will also allows the usage of
-  pointer data types, passing of hints down to the storage and support
-  for asynchronous index building, maintenance and querying.
-\item
-  Virtual Object Layer (VOL) plugin that translates HDF5 API requests
-  from applications to IOD API calls.
-\item
-  Function shipping from CN to IONs. This provides the application
-  developer with the capacity of sending computation down to the IONs
-  and get back results.
-\item
-  Analysis Shipping from CN to IONs or DAOS nodes. This is similar to 3
-  but instead of returning the result over the network, it gets stored
-  on the nodes and pointers to it are returned.
-\end{enumerate}
-
-In terms of the CN-ION communication model, a client/server architecture
-is implemented {[}8{]}: every ION runs an IOFSL (I/O Function Shipping
-Layer) server {[}9{]}; the IOFSL client is integrated into the HDF5
-library which runs on each CN. A client can forward requests to any
-number of IONs. Every I/O operation issued by HDF5 is asynchronously
-shipped to the IOFSL server and asynchronously executed.
-
-HDF5 {[}7{]} is a versatile data model containing complex data objects
-and metadata. Its information set is a collection of datasets, groups,
-datatypes and metadata objects. The data model defines mechanisms for
-creating associations between various information items. The main
-components of HDF5 are described below.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/hdf5.png}
-\caption{Sample file illustrating the hierarchical structure of HDF5
-files. \label{hdf5-sample}}
-\end{figure}
-
-\begin{itemize}
-\item
-  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
-  is represented by a file. It is a collection of objects that also
-  explains the relationship between them. Every file begins with a root
-  group ''/'', which serves as the ''starting-point'' in the object
-  hierarchy.
-\item
-  \textbf{Dataset}: HDF5 datasets are objects that represent actual data
-  or content. Datasets are arrays which can have multiple dimensions. A
-  dataset is characterized by a dataspace and a datatype. The dataspace
-  captures the rank (number of dimensions), and the current and maximum
-  extent in each dimension. The datatype describes the type of its data
-  elements.
-\item
-  \textbf{Group}: A group is an explicit association between HDF5
-  objects. It is synonymous with directories in a file system. A group
-  could contain multiple other groups, datasets or datatypes within it.
-\item
-  \textbf{Attribute}: Attributes are used for annotating datasets,
-  groups, and datatype objects. They are datasets themselves, and are
-  attached to existing objects they annotate.
-\end{itemize}
-
-For example, as shown in Figure \ref{hdf5-sample}, the file
-Sample.h5 contains the root group which itself contains a
-group G1 and two datasets, D1 and D2. Group
-G1 contains a dataset D3. Attribute A1 is
-linked to dataset D1. The objects and the relationships between
-them can be represented as a B-tree, which is used internally by HDF5 to
-index its objects.
-
-The Fast Forward project is working on the adaptation of the HDF5
-libraries in order to expose the storage stack to applications.
-
-\subsection{Virtual Object Layer}\label{virtual-object-layer}
-
-The Virtual Object Layer is an abstraction mechanism internal to the
-HDF5 library {[}\textbf{???}{]}. As shown in Figure \ref{vol-arch} it is
-implemented just below the public API. The VOL exports an interface that
-allows writing plugins for HDF5, thereby enabling developers to store
-objects in a format different from the default HDF5 file format. Plugin
-writers provide an implementation for a set of functions that access
-data on disk or on a parallel file system.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.20]{images/vol-arch.png}
-\caption{Architectural view of the VOL abstraction mechanism.
-\label{vol-arch}}
-\end{figure}
-
-\subsection{IOD VOL plugin}\label{iod-vol-plugin}
-
-The IOD VOL plugin serves as the bridge between HDF5 and the IOD API. It
-is also shown in Figure \ref{vol-arch}. The application calls the HDF5
-library while running on the system's compute nodes. Using the VOL
-architecture, the IOD VOL plugin uses a function shipper (RPC library)
-to forward the VOL calls to a server component running on the I/O nodes
-(IONs). At that point, the VOL calls are translated into I/O Dispatcher
-(IOD) API calls and executed at the IONs.
-
-The mapping between HDF5 objects and IOD objects is the following:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  file = container
-\item
-  dataset = array
-\item
-  group = key-value store
-\item
-  attribute = key-value store
-\end{itemize}
-
-\subsection{Fast Forward's I/O Dispatcher
-API}\label{fast-forwards-io-dispatcher-api}
-
-The Fast Forward I/O and Storage project is aimed at integrating many of
-the services existing currently in middleware libraries in a
-next-generation storage stack, exposed through a non-POSIX,
-transactional, asynchronous, object-based API.
-
-From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
-interface allows the user to:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  transactions
-\item
-  asynchrony
-\item
-  objects
-\item
-  layout
-\item
-  placement
-\item
-  formats
-\end{itemize}
-
-Type of objects:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  container
-\item
-  key-value store
-\item
-  blobs
-\item
-  multi-dimensional arrays
-\end{itemize}
-
-As mentioned previously, IOD allows to store distinct types of objects.
-From here on we will limit our discussion to multidimensional arrays
-(HDF5 datasets) and leave the treatment of other object types as future
-work. A multidimensional array can be sharded among the distinct I/O
-nodes in the staging area. The IOD API supports the following
-strategies:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{contiguous}. fixed chunking, distributed in a round-robin
-  fashion accross the I/O nodes.
-\item
-  \textbf{chunked}. same as above but with irregular (sparse) chunking.
-\item
-  \textbf{user-defined}. either contiguous or chunked but user specifies
-  where to place each individual shard.
-\end{itemize}
-
-It is possible to request the transformation of an object's physical
-layout to other formats, having multiple copies of the same objects in
-multiple formats if desired. Also, the user can pre-fetch objects from
-the storage cluster into the I/O nodes or read them directly from the
-storage cluster. At the semantic level (HDF5), indexes can be created
-for datasets, which results in being able to read through an index
-instead of directly from the base array.
-
-{[}sharding. \label{sharding}{]}{[}sharding{]}
-
-All these distinct alternatives result in having many different ways of
-executing the same analysis task. In the future, we plan to extend our
-design to incorporate an optimization phase that will prepare a task by
-taking into account the distinct possible ways of executing and select
-the most promising one. In our subsequent discussions, we consider only
-data-movement optimization, i.e.~sending the analysis code as close as
-possible to the data. In practice, this means we focus on identifying
-sharding of datasets and execute code accordingly over the appropriate
-shards.
-
-%%%%%%%%%%%%%%%%%%%%%%%%% >>>>>>>>>>>>>>
-
-The focus of this paper is primarily the IOD layer given the critical role it
-has in the performance and functionality of the entire stack. Most of the key
-features explored in this paper all have a strong presence in the IOD layer
-motivating the focus of this examination.
-
-The core idea for IOD is to provide a way to manage the IO load that is
-separate from the compute nodes and the storage array. Communication intensive
-activities, such as data rearrangement, can be moved to the IOD layer
-offloading the communication load from the compute nodes. IOD has three main
-purposes. First, if the optional burst buffer is available, it works as a fast
-cache absorbing write operations for the slower trickle out to the central
-storage array. It can also be used to retrieve objects from the central storage
-array for more efficient read operations and offers data filtering to make
-client reads more efficient.  Second, it offers the transaction mechanism for
-controlling data set visibility and to manage faults that would prevent a data
-set from being used. Third, data processing operations can be placed in the
-IOD. These operations are intended to offer data rearrangement, filtering, and
-similar operations prior to data reaching the central storage array.
-
-While these ideas are not necessarily new, they are new twists on best of class
-efforts for these technologies. For example, offloading the collective
-two-phase data sieving from the compute nodes to reorganize data has proven
-effective at reducing the total time for writing data due to fewer participants
-involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
-Beyond these broad items, there are many important details some of which are
-examined in more detail below.
+Overall, the architecture shifts from the idea of files and directories to
+containers of objects. This shift avoids the bottlenecks related to the POSIX
+files and directories structure such as the serialization of file creation, the
+limitation and impact of the number of files in a directory, and the limited
+semantics of a byte stream. Instead, the new interface focuses on high-level
+data models and their properties and relationships. This concept permeates the
+entire IO stack.
+
+In addition to addressing the traditional scientific workload, this project
+seeks to expand functionality to better support Big Data type applications. The
+key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
+Map-Reduce systems. Key system features are introduced to efficiently support
+these computing models in addition to the typically bursty IO loads of more
+traditional HPC applications.
+
+The basic architecture incorporates five layers some of which are potentially
+optional. The top layer is a high level IO library, such as the demonstration
+HDF-5 library~\cite{hdf5}. The intent is to only have access to the storage
+stack through such an API to manage the complexity of working with the lower
+layers. These changes incorporate the necessary features for ACGs.
+
+Below the user API is an IO forwarding layer that redirects IO calls from the
+compute nodes to the IO dispatching layer.  This IO forwarding layer is
+analogous to the function of the IO nodes in a BlueGene machine. The next two
+layers have considerable functionality. The IO dispatcher (IOD) serves as the
+primary storage interface for the IO stack and is the only way to reach the
+persistent storage array in lower layers. Ideally, the IOD layer can be
+optional based on the available hardware. Much of the functionality offered at
+this layer would shift either up or down the stack as discussed in detail
+below. The Distributed Application Object
+Storage (DAOS) layer serves as the persistent storage interface and is intended
+to be the foundation on which everything else is built with no dependence on
+any technologies specified above it. For example, the IOD layer and burst
+buffers are not required for DAOS to operate properly. At the bottom is the
+Versioning Object Storage Device (VOSD).  It serves as the interface for
+storing objects of all types efficiently.
+
+%The core idea for IOD is to provide a way to manage the IO load that is
+%separate from the compute nodes and the storage array. Communication intensive
+%activities, such as data rearrangement, can be moved to the IOD layer
+%offloading the communication load from the compute nodes. IOD has three main
+%purposes. First, if the optional burst buffer is available, it works as a fast
+%cache absorbing write operations for the slower trickle out to the central
+%storage array. It can also be used to retrieve objects from the central storage
+%array for more efficient read operations and offers data filtering to make
+%client reads more efficient.  Second, it offers the transaction mechanism for
+%controlling data set visibility and to manage faults that would prevent a data
+%set from being used. Third, data processing operations can be placed in the
+%IOD. These operations are intended to offer data rearrangement, filtering, and
+%similar operations prior to data reaching the central storage array.
+
+%While these ideas are not necessarily new, they are new twists on best of class
+%efforts for these technologies. For example, offloading the collective
+%two-phase data sieving from the compute nodes to reorganize data has proven
+%effective at reducing the total time for writing data due to fewer participants
+%involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
+%Beyond these broad items, there are many important details some of which are
+%examined in more detail below.
 
@@
 personal conversations with the team
-members~\cite{JohnBent,QuinceyKoziol,EricBarton} discussing some of the
-potential challenges with the design as written. These ideas are presented to
+members~\cite{JohnBent,QuinceyKoziol,EricBarton} understanding the current
+design as written. These ideas are presented to
 give a fuller picture of where the project is going rather than dwelling on the
@@
 paper.
+
+\section{HDF Layer}
+The API has been maintained as much as possible to make backwards compatibility
+easier. Instead, the Virtual Object Layer (VOL), a new layer under development
+for several years has been leveraged to incorporate the needed support.
+
+Some additional API functions have been introduced to support....
+
+The extensions done to HDF5 allows an application to take full advantage
+of the new exascale features. The additions comprise {[}6{]} (Figure 4):
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Object-storage API based on HDF5 to support high-level data models.
+  This exposes asynchronous, transactional semantics to the application,
+  as well as end-to-end data integrity. It will also allows the usage of
+  pointer data types, passing of hints down to the storage and support
+  for asynchronous index building, maintenance and querying.
+\item
+  Virtual Object Layer (VOL) plugin that translates HDF5 API requests
+  from applications to IOD API calls.
+\item
+  Function shipping from CN to IONs. This provides the application
+  developer with the capacity of sending computation down to the IONs
+  and get back results.
+\item
+  Analysis Shipping from CN to IONs or DAOS nodes. This is similar to 3
+  but instead of returning the result over the network, it gets stored
+  on the nodes and pointers to it are returned.
+\end{enumerate}
+
+In terms of the CN-ION communication model, a client/server architecture
+is implemented {[}8{]}: every ION runs an IOFSL (I/O Function Shipping
+Layer) server {[}9{]}; the IOFSL client is integrated into the HDF5
+library which runs on each CN. A client can forward requests to any
+number of IONs. Every I/O operation issued by HDF5 is asynchronously
+shipped to the IOFSL server and asynchronously executed.
+
+HDF5 {[}7{]} is a versatile data model containing complex data objects
+and metadata. Its information set is a collection of datasets, groups,
+datatypes and metadata objects. The data model defines mechanisms for
+creating associations between various information items. The main
+components of HDF5 are described below.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/hdf5.png}
+\caption{Sample file illustrating the hierarchical structure of HDF5
+files. \label{hdf5-sample}}
+\end{figure}
+
+\begin{itemize}
+\item
+  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
+  is represented by a file. It is a collection of objects that also
+  explains the relationship between them. Every file begins with a root
+  group ''/'', which serves as the ''starting-point'' in the object
+  hierarchy.
+\item
+  \textbf{Dataset}: HDF5 datasets are objects that represent actual data
+  or content. Datasets are arrays which can have multiple dimensions. A
+  dataset is characterized by a dataspace and a datatype. The dataspace
+  captures the rank (number of dimensions), and the current and maximum
+  extent in each dimension. The datatype describes the type of its data
+  elements.
+\item
+  \textbf{Group}: A group is an explicit association between HDF5
+  objects. It is synonymous with directories in a file system. A group
+  could contain multiple other groups, datasets or datatypes within it.
+\item
+  \textbf{Attribute}: Attributes are used for annotating datasets,
+  groups, and datatype objects. They are datasets themselves, and are
+  attached to existing objects they annotate.
+\end{itemize}
+
+For example, as shown in Figure \ref{hdf5-sample}, the file
+Sample.h5 contains the root group which itself contains a
+group G1 and two datasets, D1 and D2. Group
+G1 contains a dataset D3. Attribute A1 is
+linked to dataset D1. The objects and the relationships between
+them can be represented as a B-tree, which is used internally by HDF5 to
+index its objects.
+
+The Fast Forward project is working on the adaptation of the HDF5
+libraries in order to expose the storage stack to applications.
+
+\section{I/O Nodes}
+The ``I/O nodes'' handle 
+requests forwarded by the scientific applications, integrate a tier of
+solid-state devices to absorb the burst of random operations, and
+organize/re-format the data so that transfers from the staging area to
+the traditional parallel file system can be done more efficiently. It
+also has the capacity of executing analysis on the fresh data that has
+just been generated by simulation applications running at the compute
+nodes. As the data arrives, re-organization and data preparation can be
+applied in order to anticipate the execution of analytical tasks
+{[}3{]}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-arch.png}
+\caption{Exascale Architecture. \label{exa-arch}}
+\end{figure}
+
+A common cluster configuration for this type of deployment is shown in
+Figure \ref{exa-arch}. The designated I/O nodes (IONs) are connected to
+the compute nodes (CNs) through the same fast fabric (eg. Infiniband),
+while the connection to the external storage cluster is through a
+secondary slower channel (eg. 10Gb ethernet). I/O staging thus
+introduces a new layer in the application-to-storage stack, as shown in
+Figure \ref{exa-stack} (left).
+
+In current proposals, the stack is managed by middleware that seats
+between the application and the parallel file system {[}2,4{]}. Since
+most of the applications assume a POSIX interface to storage {[}5{]},
+existing middleware either transparently handle the I/O operations,
+appearing as regular POSIX calls to applications; or modify the I/O API
+as little as possible in order to minimize the impact on production
+codes.
+
+As we move towards the exascale goal, this way of interfacing with
+storage becomes more of an obstacle. Many features provided by the stack
+are hidden by the middleware/POSIX layers. If these were visible to the
+applications, and if the applications were able to control how the
+underlying layers behave, they would benefit greatly since the domain
+knowledge that is available at the application level could be used to
+execute storage operations efficiently. One more implication of the
+25-year-old POSIX API is in cumbersome middleware, since metadata has to
+be part of the same uni-dimensional byte array being stored.
+
+The Fast Forward Storage I/O project {[}6{]} is aimed at merging the
+features of existing middleware into a next-generation parallel file
+system stack. This new storage API termed the I/O dispatcher (IOD) is
+object-based, transactional, asynchronous and I/O-staging-aware (Figure
+\ref{exa-stack} (right)).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-stack.png}
+\caption{Top-to-bottom view of the component stack. \label{exa-stack}}
+\end{figure}
+
+The prototype stack, from a top-down point of view, has the following
+components:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Arbitrarily Connected Graphs (ACG).
+\item
+  HDF5 extensions.
+\item
+  I/O Dispatcher (IOD).
+\item
+  Distributed Application Object Storage (DAOS).
+\item
+  Versioning Object Storage Device (VOSD).
+\end{enumerate}
+
+Applications running on the compute nodes will be written in Python
+scripts or in C++ using the HDF5 APIs. These applications will execute
+analysis on graph-based data (ACG). The point is to demonstrate both the
+HPC and BigData use cases of the Exascale architecture. The data
+structures will be stored in HDF5, for which new API extensions will be
+implemented. These new API calls will expose the transactional,
+asynchronous and function-forwarding semantics of the underlying stack.
+
+As mentioned before, the stack provides \emph{non-POSIX, object-based,
+transactional and asynchronous active-storage}, meaning that POSIX is
+supplanted by new object interfaces that reach up to the HDF5 layer
+(Application I/O in the stack diagram); transactional semantics are
+present, from HDF down to the VOSD layer (Storage layer); clients don't
+need to wait for any blocking operation; and analysis can be shipped and
+executed on I/O or DAOS nodes.
+
+We give a high-level description of each of layer.
+
+\subsection{ACG}\label{acg}
+
+This will exemplify how applications make use of the Exascale stack. For
+FF, support for
+GraphLab 
+and GraphBuilder will be prototyped, which are graph-processing
+frameworks. GraphBuilder is a set of MapReduce tasks that extract,
+normalize, partition and serialize a graph out of unstructured data, and
+writes graph-specific formats into HDFS. These files are later consumed
+by GraphLab, a vertex-centric, asynchronous execution engine that runs
+directly on top of HDFS (i.e. non-MapReduce). The following illustrates
+the architecture of both frameworks:
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
+\caption{Figure 3. GraphLab and GraphBuilder stacks (taken from
+{[}4{]}).}
+\end{figure}
+
+In order to make both work on top of the exascale stack, both have to be
+modified. After these modifications are implemented, GraphBuilder will
+be able to write the partitioned graph in (the newly proposed) HDF5
+files which will thus be stored in the IOD nodes (or IONs) in a
+parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
+the library to perform at high speeds by benefiting from the new
+features (see next section). In general both frameworks will be modified
+so that calls to HDFS-based formats are replaced by the proposed HDF5
+ones. This is referred to as the HDF Adaptation Layer or HAL and will
+provide, from the GraphBuilder/GraphLab point of view {[}5{]}:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  capability for storing the newly proposed HDF5 format
+\item
+  association of network information to vertices/edges
+\item
+  shipping computation to the IONs
+\item
+  asynchronous vertex updates
+\item
+  efficient data sharing ammong CNs
+\item
+  computation over versioned datasets
+\end{itemize}
+
+\subsection{Virtual Object Layer}\label{virtual-object-layer}
+
+The Virtual Object Layer is an abstraction mechanism internal to the
+HDF5 library {[}\textbf{???}{]}. As shown in Figure \ref{vol-arch} it is
+implemented just below the public API. The VOL exports an interface that
+allows writing plugins for HDF5, thereby enabling developers to store
+objects in a format different from the default HDF5 file format. Plugin
+writers provide an implementation for a set of functions that access
+data on disk or on a parallel file system.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.20]{images/vol-arch.png}
+\caption{Architectural view of the VOL abstraction mechanism.
+\label{vol-arch}}
+\end{figure}
+
+\subsection{IOD VOL plugin}\label{iod-vol-plugin}
+
+The IOD VOL plugin serves as the bridge between HDF5 and the IOD API. It
+is also shown in Figure \ref{vol-arch}. The application calls the HDF5
+library while running on the system's compute nodes. Using the VOL
+architecture, the IOD VOL plugin uses a function shipper (RPC library)
+to forward the VOL calls to a server component running on the I/O nodes
+(IONs). At that point, the VOL calls are translated into I/O Dispatcher
+(IOD) API calls and executed at the IONs.
+
+The mapping between HDF5 objects and IOD objects is the following:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  file = container
+\item
+  dataset = array
+\item
+  group = key-value store
+\item
+  attribute = key-value store
+\end{itemize}
+
+\subsection{Fast Forward's I/O Dispatcher
+API}\label{fast-forwards-io-dispatcher-api}
+
+The Fast Forward I/O and Storage project is aimed at integrating many of
+the services existing currently in middleware libraries in a
+next-generation storage stack, exposed through a non-POSIX,
+transactional, asynchronous, object-based API.
+
+From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
+interface allows the user to:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  transactions
+\item
+  asynchrony
+\item
+  objects
+\item
+  layout
+\item
+  placement
+\item
+  formats
+\end{itemize}
+
+Type of objects:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  container
+\item
+  key-value store
+\item
+  blobs
+\item
+  multi-dimensional arrays
+\end{itemize}
+
+As mentioned previously, IOD allows to store distinct types of objects.
+From here on we will limit our discussion to multidimensional arrays
+(HDF5 datasets) and leave the treatment of other object types as future
+work. A multidimensional array can be sharded among the distinct I/O
+nodes in the staging area. The IOD API supports the following
+strategies:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{contiguous}. fixed chunking, distributed in a round-robin
+  fashion accross the I/O nodes.
+\item
+  \textbf{chunked}. same as above but with irregular (sparse) chunking.
+\item
+  \textbf{user-defined}. either contiguous or chunked but user specifies
+  where to place each individual shard.
+\end{itemize}
+
+It is possible to request the transformation of an object's physical
+layout to other formats, having multiple copies of the same objects in
+multiple formats if desired. Also, the user can pre-fetch objects from
+the storage cluster into the I/O nodes or read them directly from the
+storage cluster. At the semantic level (HDF5), indexes can be created
+for datasets, which results in being able to read through an index
+instead of directly from the base array.
+
+{[}sharding. \label{sharding}{]}{[}sharding{]}
+
+All these distinct alternatives result in having many different ways of
+executing the same analysis task. In the future, we plan to extend our
+design to incorporate an optimization phase that will prepare a task by
+taking into account the distinct possible ways of executing and select
+the most promising one. In our subsequent discussions, we consider only
+data-movement optimization, i.e.~sending the analysis code as close as
+possible to the data. In practice, this means we focus on identifying
+sharding of datasets and execute code accordingly over the appropriate
+shards.
 

### 2014-04-22 paper.tex
major restructuring. title and abstract completely rewritten. Into largely changed. Rest of the paper reorganized to talk about each layer and then the cross-cutting ideas.

--- 2014-04-17 paper.tex
@@
 summary of important aspects of the design specifications but also capture the
-underlying reasoning that is not available elsewhere.
+underlying reasoning that is not available elsewhere. With this information, we
+encourate the broader community to understand the design, intent, and future
+directions to foster discussion guiding phase 2 and the ultimate production
+storage stack based on this work.
 
@@
 offering different performance and quality of service guarantees. The specifics
-of how the different layers interact offer a new approach for managing
-scientific data for extreme scale platforms.
-
-The integration of the HDF5 interface as the user interface gives a familiar
-way to imteract with all of the new functionality while largely insulating 
+of how the different layers interact offer a new approach for managing data for
+extreme scale platforms and workloads.
+
+For the current demonstration system, the HDF5 user API gives a familiar
+way to interact with all of the new functionality while largely insulating 
 end-users from the intracisies of dealing with the advanced functionality
-offered by the stack.
-
-The optional components aid in the scalability and simplifying implementation
-across a wide range of platforms.
-
-The data storage layer offers support both for local machine and data center
-wide storage arrays as part of the design.
-
-In addition to these features supporting HPC platforms, the stack also addresses
-the need of big data applications as demonstrated through the Arbitrary Connected Graphs tool.
+offered by the stack.  The various layers below the user API include optional
+components that may aid in scalability and simplify implementation across a
+wide range of platforms.  The data storage layer offers support both
+for local machine and data center wide storage arrays as part of the design.
+
+In addition to these features supporting HPC platforms, the stack also
+addresses the need of big data applications as demonstrated through the
+Arbitrary Connected Graphs tool.
 
@@
 currently getting underway as of early 2014. The core focus of the first phase
-was basic functionality and design. Overriding many of the decisions during
-this and any subsequent phases is the reality of budgets. Placing GBs of NVRAM
-on every node, while a potentially advantageous approach, is not financially
-feasible. With this in mind, the second phase will refine this design
-incorporating fault recovery and other features missing from the first phase.
-
-The overall design seeks to offer high availability, byte-granular,
+was basic functionality and design. While an idealized potential system would
+be the perfect target architecture, the reality of budgets has tempered many of
+the decisions. For exmaple, extensive availability of NVRAN or SSDs on all of
+the compute nodes is currently not economically feasible limiting some of the
+potential design choices.  With this in mind, the second phase will refine this
+design incorporating fault recovery and other features missing from the first
+phase.
+
+The complete design seeks to offer high availability, byte-granular,
 multi-version concurrency control. Through the use of a copy-on-write style
-mechanism, multiple versions of an object can be stored in potentially greatly
-reduced space. It assumes the client interface will be through an IO library
-affording a more complicated interface that offers richer functionality
-requiring only minimal end-user code changes.  Managing most data access in a
-platform-local layer rather than requiring writing to centralized storage will
-better support the performance and energy requirements of extreme scale
-integrated application workflows. At a more detailed view, the various layers
-of the IO stack each contribute different functionality and performance
-implications.
-
-The basic architecture incorporates five layers. The top layer is a high level
-IO library, such as the demonstration HDF-5 library~\cite{hdf5}. The intent is
-to only have access to the storage stack through such an API to manage the
-complexity of working with the lower layers. Below the user API is an IO
-forwarding layer that redirects IO calls from the compute nodes to the IO
-dispatching layer.  This IO forwarding layer is analogous to the function of
-the IO nodes in a BlueGene machine. The next two layers have considerable
-functionality. The IO dispatcher (IOD) serves as the primary storage interface
-for the IO stack and is the only way to reach the persistent storage array in
-lower layers. The Distributed Application Object Storage (DAOS) layer serves as
-the persistent storage interface and is intended to be the foundation on which
-everything else is built with no dependence on any technologies specified above
-it. For example, the IOD layer and burst buffers are not required for DAOS to
-operate properly. At the bottom is the Versioning Object Storage Device (VOSD).
-It serves as the interface for storing objects of all types efficiently.
-
-%%%%%%%%%%%%%%%%%%%%%%%%% <<<<<<<<<<<<<<<<
-
-The ``I/O nodes'' handle 
-requests forwarded by the scientific applications, integrate a tier of
-solid-state devices to absorb the burst of random operations, and
-organize/re-format the data so that transfers from the staging area to
-the traditional parallel file system can be done more efficiently. It
-also has the capacity of executing analysis on the fresh data that has
-just been generated by simulation applications running at the compute
-nodes. As the data arrives, re-organization and data preparation can be
-applied in order to anticipate the execution of analytical tasks
-{[}3{]}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/exa-arch.png}
-\caption{Exascale Architecture. \label{exa-arch}}
-\end{figure}
-
-A common cluster configuration for this type of deployment is shown in
-Figure \ref{exa-arch}. The designated I/O nodes (IONs) are connected to
-the compute nodes (CNs) through the same fast fabric (eg. Infiniband),
-while the connection to the external storage cluster is through a
-secondary slower channel (eg. 10Gb ethernet). I/O staging thus
-introduces a new layer in the application-to-storage stack, as shown in
-Figure \ref{exa-stack} (left).
-
-In current proposals, the stack is managed by middleware that seats
-between the application and the parallel file system {[}2,4{]}. Since
-most of the applications assume a POSIX interface to storage {[}5{]},
-existing middleware either transparently handle the I/O operations,
-appearing as regular POSIX calls to applications; or modify the I/O API
-as little as possible in order to minimize the impact on production
-codes.
-
-As we move towards the exascale goal, this way of interfacing with
-storage becomes more of an obstacle. Many features provided by the stack
-are hidden by the middleware/POSIX layers. If these were visible to the
-applications, and if the applications were able to control how the
-underlying layers behave, they would benefit greatly since the domain
-knowledge that is available at the application level could be used to
-execute storage operations efficiently. One more implication of the
-25-year-old POSIX API is in cumbersome middleware, since metadata has to
-be part of the same uni-dimensional byte array being stored.
-
-The Fast Forward Storage I/O project {[}6{]} is aimed at merging the
-features of existing middleware into a next-generation parallel file
-system stack. This new storage API termed the I/O dispatcher (IOD) is
-object-based, transactional, asynchronous and I/O-staging-aware (Figure
-\ref{exa-stack} (right)).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/exa-stack.png}
-\caption{Top-to-bottom view of the component stack. \label{exa-stack}}
-\end{figure}
-
-The prototype stack, from a top-down point of view, has the following
-components:
+mechanism reducing space and time requirements, multiple versions of an object
+can be stored efficiently. By assuming the client interface will be through an
+IO library, a more complicated interface that offers richer functionality can
+be incorporated while requiring only minimal end-user code changes.  Managing
+most data access in a platform-local layer rather than requiring writing to
+centralized storage will better support the performance and energy requirements
+of extreme scale integrated application workflows. At a more detailed view, the
+various layers of the IO stack each contribute different functionality and
+performance implications.
+
+Overall, the architecture shifts from the idea of files and directories to
+containers of objects. This shift avoids the bottlenecks related to the POSIX
+files and directories structure such as the serialization of file creation, the
+limitation and impact of the number of files in a directory, and the limited
+semantics of a byte stream. Instead, the new interface focuses on high-level
+data models and their properties and relationships. This concept permeates the
+entire IO stack.
+
+In addition to addressing the traditional scientific workload, this project
+seeks to expand functionality to better support Big Data type applications. The
+key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
+Map-Reduce systems. Key system features are introduced to efficiently support
+these computing models in addition to the typically bursty IO loads of more
+traditional HPC applications.
+
+The architecture incorporates five layers, some of which are potentially
+optional. The top layer is a high level IO library, such as the demonstration
+HDF-5 library~\cite{hdf5}. The intent is to only have access to the storage
+stack through such an API to manage the complexity of working with the lower
+layers and to enable advanced functionality that would require more direct user
+intervention. This layer incorporate the necessary features for ACGs from a
+end-user's perspective.
+
+Below the user API is an IO forwarding layer that redirects IO calls from the
+compute nodes to the IO dispatching layer.  This IO forwarding layer is
+analogous to the function of the IO nodes in a BlueGene machine or the passive
+data staging processes demonstrated
+previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
+have considerable functionality. The IO dispatcher (IOD) serves as the primary
+storage interface for the IO stack and is the only way to reach the persistent
+storage array in lower layers. Ideally, the IOD layer's functionality can be
+optional based on available hardware. Much of the functionality offered at this
+layer would shift either up or down the stack as discussed in detail below. The
+Distributed Asynchronous Object Storage (DAOS) layer serves as the persistent
+storage interface and is intended to be the traditional file system-like
+foundation on which everything else is built with no dependence on any
+technologies specified above it. For example, the IOD layer with or without
+burst buffers is not required for DAOS to operate properly.  Instead, the DAOS
+layer can handle all of the IO operations from the user API layer, albiet with
+the potential performance penalty of manipulating the shared, persistent
+storage array. At the bottom is the Versioning Object Storage Device (VOSD).
+It serves as the interface for storing objects of all types efficiently. Think
+of this layer as the physical disk interface layer. In terms of Lustre, this
+would replace the file system on individual storage targets with an interface
+friendlier to the containers of objects concept used in the higher layers.
+
+Along with the analysis of the published design documents, a discussion of the
+design philosophy representing the overall intent is presented. This
+information represents information that may or may not have been written down,
+but is the intent of ultimate product. These insights were gained based on
+personal conversations with the team
+members~\cite{JohnBent,QuinceyKoziol,EricBarton} understanding the current
+design as written. These ideas are presented to
+give a fuller picture of where the project is going rather than dwelling on the
+limitations of the published documents.
+
+The rest of the paper is organized as follows. A brief overview of related work
+is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
+discusses the programmatic interface end users will see when interacting with
+the storage array. This will be discussed in the contect of the HDF-5 based
+example library used for the functionality demonstration. Section~\ref{sec:iof}
+briefly discusses the motivation and proposal for the IO forwarding layer.
+Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
+functionality it offers. This will detail the pieces of the layer that are
+potentially optional and mention the cross-cutting features discussed in a
+later, cross-cutting section. Section~\ref{sec:daos} discusses how the
+persistent storage array interface itself functions. As with the IOD layer, the
+disucssion of cross-cutting features will be mentioned, but disucssed more
+fully in the cross-cutting section. The VOSD layer is disucssed in
+Section~\ref{sec:vosd}. In particular, the mapping between the DAOS and VOSD
+layers are explored as it pertains to the physical storage. Next is an
+exploration of cross-cutting features like transactions and metadata management
+in Section~\ref{sec:broader}. Since these and other features are spread across
+multiple layers, it makes more sense to discuss them independently once an
+understanding of the overall structure has been presented.  The paper is
+concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
+covered in the paper.
+
+\section{Related Work}
+\label{sec:related}
+
+Many projects over the last couple of decades have sought to address some
+challenging aspect of parallel file system design. The recent rise of the ``Big
+Data'' applications with different characteristic IO patterns have somewhat
+complicated the picture. Extreme scale machines will be expected to handle both
+the traditional simulation-related workloads as well as applications more
+squarely in the Big Data arena. This will require some adjustments to the
+underlying system for good performance for both scenarios.
+
+The major previous work is really limited to full file systems rather than the
+mountain of file system adjustments made over the years. The major players
+related to this work include Ceph, PVFS, Lustre, GPFS, and Panasas.
+
+EXPLAIN A BIT ABOUT HOW EACH CANNOT ADDRESS THE PROBLEMS WITH A CITATION FOR
+EACH. DID I MISS A MAJOR FILE SYSTEM OF NOTE?
+
+\section{End-User API Layer}
+\label{sec:end-user}
+Since the proposal specifies a high-level IO API will be the primary end-user
+interface for programmatically interacting with the FFSIO stack, the team used
+the HDF-5 API and leverated the Virtual Object Layer (VOL) for the initial
+design and implementation demonstration. The VOL provides a way to intercept
+HDF-5 API calls just below the API layer to implement alternative
+functionality. For this project, rather than the default writing to disk in the
+HDF-5 format, the VOL layer is used to interact with the IOD layer and the new
+concepts it offers, such as transactions, completely hidden from the user.
+This both offers a straightforward way for existing applications to test
+against the FFSIO stack and to ensure proper semantics are maintained at the
+lower levels. Some additional API functions have been introduced to support
+new functionality, such as the ACG support for Hadoop-style applications.  The
+additions comprise (Figure~\ref{fig:vol-arch}):
 
@@
 \itemsep1pt\parskip0pt\parsep0pt
-\item
-  Arbitrarily Connected Graphs (ACG).
-\item
-  HDF5 extensions.
-\item
-  I/O Dispatcher (IOD).
-\item
-  Distributed Application Object Storage (DAOS).
-\item
-  Versioning Object Storage Device (VOSD).
-\end{enumerate}
-
-Applications running on the compute nodes will be written in Python
-scripts or in C++ using the HDF5 APIs. These applications will execute
-analysis on graph-based data (ACG). The point is to demonstrate both the
-HPC and BigData use cases of the Exascale architecture. The data
-structures will be stored in HDF5, for which new API extensions will be
-implemented. These new API calls will expose the transactional,
-asynchronous and function-forwarding semantics of the underlying stack.
-
-As mentioned before, the stack provides \emph{non-POSIX, object-based,
-transactional and asynchronous active-storage}, meaning that POSIX is
-supplanted by new object interfaces that reach up to the HDF5 layer
-(Application I/O in the stack diagram); transactional semantics are
-present, from HDF down to the VOSD layer (Storage layer); clients don't
-need to wait for any blocking operation; and analysis can be shipped and
-executed on I/O or DAOS nodes.
-
-We give a high-level description of each of layer.
-
-\subsection{ACG}\label{acg}
-
-This will exemplify how applications make use of the Exascale stack. For
-FF, support for
-GraphLab 
-and GraphBuilder will be prototyped, which are graph-processing
-frameworks. GraphBuilder is a set of MapReduce tasks that extract,
-normalize, partition and serialize a graph out of unstructured data, and
-writes graph-specific formats into HDFS. These files are later consumed
-by GraphLab, a vertex-centric, asynchronous execution engine that runs
-directly on top of HDFS (i.e. non-MapReduce). The following illustrates
-the architecture of both frameworks:
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
-\caption{Figure 3. GraphLab and GraphBuilder stacks (taken from
-{[}4{]}).}
-\end{figure}
-
-In order to make both work on top of the exascale stack, both have to be
-modified. After these modifications are implemented, GraphBuilder will
-be able to write the partitioned graph in (the newly proposed) HDF5
-files which will thus be stored in the IOD nodes (or IONs) in a
-parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
-the library to perform at high speeds by benefiting from the new
-features (see next section). In general both frameworks will be modified
-so that calls to HDFS-based formats are replaced by the proposed HDF5
-ones. This is referred to as the HDF Adaptation Layer or HAL and will
-provide, from the GraphBuilder/GraphLab point of view {[}5{]}:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  capability for storing the newly proposed HDF5 format
-\item
-  association of network information to vertices/edges
-\item
-  shipping computation to the IONs
-\item
-  asynchronous vertex updates
-\item
-  efficient data sharing ammong CNs
-\item
-  computation over versioned datasets
-\end{itemize}
-
-\subsection{HDF5 extensions}\label{hdf5-extensions}
-
-The extensions done to HDF5 allows an application to take full advantage
-of the new exascale features. The additions comprise {[}6{]} (Figure 4):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
+
 \item
@@
   for asynchronous index building, maintenance and querying.
+
 \item
@@
   from applications to IOD API calls.
-\item
-  Function shipping from CN to IONs. This provides the application
-  developer with the capacity of sending computation down to the IONs
-  and get back results.
-\item
-  Analysis Shipping from CN to IONs or DAOS nodes. This is similar to 3
-  but instead of returning the result over the network, it gets stored
-  on the nodes and pointers to it are returned.
+
+\item\label{item:fs}
+  Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
+  the application developer with the capacity of sending computation down to the
+  IONs and get back results.
+
+\item
+  Analysis Shipping from CN to IONs or DAOS nodes. This is similar
+  to~\ref{item:fs} but instead of returning the result over the network, it
+  gets stored on the nodes and pointers to it are returned.
+
 \end{enumerate}
 
-In terms of the CN-ION communication model, a client/server architecture
-is implemented {[}8{]}: every ION runs an IOFSL (I/O Function Shipping
-Layer) server {[}9{]}; the IOFSL client is integrated into the HDF5
-library which runs on each CN. A client can forward requests to any
-number of IONs. Every I/O operation issued by HDF5 is asynchronously
-shipped to the IOFSL server and asynchronously executed.
-
-HDF5 {[}7{]} is a versatile data model containing complex data objects
+Function and Analysis Shipping are part of the cross-cutting features and are
+discussed in Section~\ref{sec:fn-shipping}.
+
+HDF5~\cite{hdf5} is a versatile data model containing complex data objects
 and metadata. Its information set is a collection of datasets, groups,
@@
 \includegraphics[scale=0.30]{images/hdf5.png}
-\caption{Sample file illustrating the hierarchical structure of HDF5
-files. \label{hdf5-sample}}
+\caption{Sample file illustrating the hierarchical structure of HDF5 files}
+\label{fig:hdf5-sample}
 \end{figure}
@@
 \begin{itemize}
+
 \item
@@
   hierarchy.
+
 \item
@@
   elements.
+
 \item
@@
   could contain multiple other groups, datasets or datatypes within it.
+
 \item
@@
   attached to existing objects they annotate.
+
 \end{itemize}
 
-For example, as shown in Figure \ref{hdf5-sample}, the file
-Sample.h5 contains the root group which itself contains a
-group G1 and two datasets, D1 and D2. Group
-G1 contains a dataset D3. Attribute A1 is
-linked to dataset D1. The objects and the relationships between
-them can be represented as a B-tree, which is used internally by HDF5 to
-index its objects.
-
-The Fast Forward project is working on the adaptation of the HDF5
-libraries in order to expose the storage stack to applications.
-
-\subsection{Virtual Object Layer}\label{virtual-object-layer}
-
-The Virtual Object Layer is an abstraction mechanism internal to the
-HDF5 library {[}\textbf{???}{]}. As shown in Figure \ref{vol-arch} it is
-implemented just below the public API. The VOL exports an interface that
-allows writing plugins for HDF5, thereby enabling developers to store
-objects in a format different from the default HDF5 file format. Plugin
-writers provide an implementation for a set of functions that access
-data on disk or on a parallel file system.
+For example, as shown in Figure \ref{fig:hdf5-sample}, the file Sample.h5
+contains the root group which itself contains a group G1 and two datasets, D1
+and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1.
+The objects and the relationships between them can be represented as a B-tree,
+which is used internally by HDF5 to index its objects.
+
+The Fast Forward project is working on the adaptation of the HDF5 libraries in
+order to expose the storage stack to applications.
+
+\subsection{Virtual Object Layer}
+\label{virtual-object-layer}
+
+The Virtual Object Layer is an abstraction mechanism internal to the HDF5
+library~\cite{hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
+just below the public API. The VOL exports an interface that allows writing
+plugins for HDF5, thereby enabling developers to handle the data in ways other
+than necessarily writing to storage in an HDF5 format.  Plugin writers provide
+an implementation for a set of functions that are trusted to provide the
+proper semantics represented by the function for the new environment. For
+example, data staging could be implemented in the VOL layer by replacing
+writing to disk in the HDF5 format to sending data to a data staging area using
+some messaging mechanism.
 
@@
 \includegraphics[scale=0.20]{images/vol-arch.png}
-\caption{Architectural view of the VOL abstraction mechanism.
-\label{vol-arch}}
+\caption{Architectural view of the VOL abstraction mechanism}
+\label{fig:vol-arch}
 \end{figure}
 
-\subsection{IOD VOL plugin}\label{iod-vol-plugin}
+\subsection{IOD VOL plugin}
+\label{iod-vol-plugin}
 
@@
 
-\subsection{Fast Forward's I/O Dispatcher
-API}\label{fast-forwards-io-dispatcher-api}
-
-The Fast Forward I/O and Storage project is aimed at integrating many of
-the services existing currently in middleware libraries in a
-next-generation storage stack, exposed through a non-POSIX,
-transactional, asynchronous, object-based API.
-
-From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
-interface allows the user to:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  transactions
-\item
-  asynchrony
-\item
-  objects
-\item
-  layout
-\item
-  placement
-\item
-  formats
-\end{itemize}
-
-Type of objects:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  container
-\item
-  key-value store
-\item
-  blobs
-\item
-  multi-dimensional arrays
-\end{itemize}
-
-As mentioned previously, IOD allows to store distinct types of objects.
-From here on we will limit our discussion to multidimensional arrays
-(HDF5 datasets) and leave the treatment of other object types as future
-work. A multidimensional array can be sharded among the distinct I/O
-nodes in the staging area. The IOD API supports the following
-strategies:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{contiguous}. fixed chunking, distributed in a round-robin
-  fashion accross the I/O nodes.
-\item
-  \textbf{chunked}. same as above but with irregular (sparse) chunking.
-\item
-  \textbf{user-defined}. either contiguous or chunked but user specifies
-  where to place each individual shard.
-\end{itemize}
-
-It is possible to request the transformation of an object's physical
-layout to other formats, having multiple copies of the same objects in
-multiple formats if desired. Also, the user can pre-fetch objects from
-the storage cluster into the I/O nodes or read them directly from the
-storage cluster. At the semantic level (HDF5), indexes can be created
-for datasets, which results in being able to read through an index
-instead of directly from the base array.
-
-{[}sharding. \label{sharding}{]}{[}sharding{]}
-
-All these distinct alternatives result in having many different ways of
-executing the same analysis task. In the future, we plan to extend our
-design to incorporate an optimization phase that will prepare a task by
-taking into account the distinct possible ways of executing and select
-the most promising one. In our subsequent discussions, we consider only
-data-movement optimization, i.e.~sending the analysis code as close as
-possible to the data. In practice, this means we focus on identifying
-sharding of datasets and execute code accordingly over the appropriate
-shards.
-
-%%%%%%%%%%%%%%%%%%%%%%%%% >>>>>>>>>>>>>>
-
-The focus of this paper is primarily the IOD layer given the critical role it
-has in the performance and functionality of the entire stack. Most of the key
-features explored in this paper all have a strong presence in the IOD layer
-motivating the focus of this examination.
+\section{IO Forwarding Layer}
+\label{sec:iof}
+
+This section will briefly describe the limited functionality of the IO
+Forwarding layer and why it was included in the design.  It will not exceed
+probably half of a column.
+
+\section{IOD Layer}
+\label{sec:iod}
 
@@
 
-Along with the analysis of the published design documents, a discussion of the
-design philosophy representing the overall intent is presented. This
-information represents information that may or may not have been written down,
-but is the intent of ultimate product. These insights were gained based on
-personal conversations with the team
-members~\cite{JohnBent,QuinceyKoziol,EricBarton} discussing some of the
-potential challenges with the design as written. These ideas are presented to
-give a fuller picture of where the project is going rather than dwelling on the
-limitations of the published documents.
-
-The rest of the paper is organized as follows. Section~\ref{sec:burst}
-discusses some of the features of incorporating burst buffers as designed and
-suggests some considerations and alternatives for the next generation of this
-project. Section~\ref{sec:transactions} discusses the transactions approach
-offered in the IOD layer and the corresponding epochs in the DAOS layer. It
-also offers a comparison to the \DDT system given the very similar high-level
-design and motivating use case.  Section~\ref{sec:summary} discusses the system
-overall with recommendations on what design elements should be considered based
-on broader issues with current HPC data centers. The paper is concluded in
-Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
-paper.
-
-\section{Burst Buffers}
+\subsection{I/O Dispatcher API}
+\label{sec:fast-forwards-io-dispatcher-api}
+
+The Fast Forward I/O and Storage project is aimed at integrating many of
+the services existing currently in middleware libraries in a
+next-generation storage stack, exposed through a non-POSIX,
+transactional, asynchronous, object-based API.
+
+From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
+interface allows the user to:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  transactions
+\item
+  asynchrony
+\item
+  objects
+\item
+  layout
+\item
+  placement
+\item
+  formats
+\end{itemize}
+
+Type of objects:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  container
+\item
+  key-value store
+\item
+  blobs
+\item
+  multi-dimensional arrays
+\end{itemize}
+
+As mentioned previously, IOD allows to store distinct types of objects.
+From here on we will limit our discussion to multidimensional arrays
+(HDF5 datasets) and leave the treatment of other object types as future
+work. A multidimensional array can be sharded among the distinct I/O
+nodes in the staging area. The IOD API supports the following
+strategies:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{contiguous}. fixed chunking, distributed in a round-robin
+  fashion accross the I/O nodes.
+\item
+  \textbf{chunked}. same as above but with irregular (sparse) chunking.
+\item
+  \textbf{user-defined}. either contiguous or chunked but user specifies
+  where to place each individual shard.
+\end{itemize}
+
+It is possible to request the transformation of an object's physical
+layout to other formats, having multiple copies of the same objects in
+multiple formats if desired. Also, the user can pre-fetch objects from
+the storage cluster into the I/O nodes or read them directly from the
+storage cluster. At the semantic level (HDF5), indexes can be created
+for datasets, which results in being able to read through an index
+instead of directly from the base array.
+
+All these distinct alternatives result in having many different ways of
+executing the same analysis task. In the future, we plan to extend our
+design to incorporate an optimization phase that will prepare a task by
+taking into account the distinct possible ways of executing and select
+the most promising one. In our subsequent discussions, we consider only
+data-movement optimization, i.e., sending the analysis code as close as
+possible to the data. In practice, this means we focus on identifying
+sharding of datasets and execute code accordingly over the appropriate
+shards.
+
+\subsubsection{Sharding}
+\label{sec:sharding}
+
+The idea of sharding is to store a portion of a data set on a single storage
+target. This is similiar in concept to the HDF-5 hyperslab. The FFSIO stack
+supports sharding the data in the default or some other structured way as well
+as ``re-sharding'' based on application needs. For example, reordering the
+data so that a different dimension is the ``fast'' dimension may greatly
+improve the performance of a subsequent data analytics task. A common scenario
+where this is common is a Fortran code (column-major) writes data for a C code
+(row-major) to analyze.
+
+\subsection{I/O Nodes}
+The ``I/O nodes'' handle requests forwarded by the scientific applications,
+potentially integrate a tier of solid-state devices to absorb the burst of
+random operations, and organize/re-format the data so that transfers to/from
+the staging area from/to the traditional parallel file system can be done more
+efficiently. It also has the capacity to executing analysis on the fresh data
+that has just been generated by simulation applications running at the compute
+nodes. As the data arrives, re-organization and data preparation can be applied
+in order to anticipate the execution of analytical tasks.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-arch.png}
+\caption{Exascale Architecture}
+\label{fig:exa-arch}
+\end{figure}
+
+A common cluster configuration for this type of deployment is shown in
+Figure~\ref{fig:exa-arch}. The designated I/O nodes (IONs) are connected to the
+compute nodes (CNs) through the same fast fabric (e.g., InfiniBand), while the
+connection to the external storage cluster is through a secondary slower
+channel (e.g., 10Gb Ethernet). I/O staging thus introduces a new layer in the
+application-to-storage stack, as shown in Figure~\ref{fig:exa-stack} (left).
+
+In current proposals, the stack is managed by middleware that sits between the
+application and the parallel file system. Since most of the applications assume
+a POSIX interface to storage, existing middleware either transparently handle
+the I/O operations, appearing as regular POSIX calls to applications; or modify
+the I/O API as little as possible in order to minimize the impact on production
+codes.
+
+As we move towards the exascale goal, this way of interfacing with storage
+becomes more of an obstacle. Many features provided by the stack are hidden by
+the middleware/POSIX layers. If these were visible to the applications, and if
+the applications were able to control how the underlying layers behave, they
+would benefit greatly since the domain knowledge that is available at the
+application level could be used to execute storage operations efficiently. One
+more implication of the 25-year-old POSIX API is in cumbersome middleware,
+since metadata has to be part of the same uni-dimensional byte array being
+stored.
+
+The Fast Forward Storage I/O project is aimed at merging the features of
+existing middleware into a next-generation parallel file system stack. This new
+storage API termed the I/O dispatcher (IOD) is object-based, transactional,
+asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.30]{images/exa-stack.png}
+\caption{Top-to-bottom view of the component stack}
+\label{fig:exa-stack}
+\end{figure}
+
+The prototype stack, from a top-down point of view, has the following
+components:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Arbitrarily Connected Graphs (ACG).
+\item
+  HDF5 extensions.
+\item
+  I/O Dispatcher (IOD).
+\item
+  Distributed Asynchronous Object Storage (DAOS).
+\item
+  Versioning Object Storage Device (VOSD).
+\end{enumerate}
+
+Applications running on the compute nodes will be written in Python scripts or
+in C++ using the HDF5 APIs. These applications will execute analysis on
+graph-based data (ACG). The point is to demonstrate both the HPC and BigData
+use cases of the Exascale architecture. The data structures will be stored in
+HDF5, for which new API extensions will be implemented. These new API calls
+will expose the transactional, asynchronous and function-forwarding semantics
+of the underlying stack.
+
+As mentioned before, the stack provides \emph{non-POSIX, object-based,
+transactional and asynchronous active-storage}, meaning that POSIX is
+supplanted by new object interfaces that reach up to the HDF5 layer
+(Application I/O in the stack diagram); transactional semantics are present,
+from HDF down to the VOSD layer (Storage layer); clients don't need to wait for
+any blocking operation; and analysis can be shipped and executed on I/O or DAOS
+nodes.
+
+We give a high-level description of each of layer.
+
+\subsection{Burst Buffers}
 \label{sec:burst}
@@
 
-\subsection{Design Philosophy}
+\noindent\textbf{Design Philosophy}
+
 The burst buffers design, as presented in the IOD documents, limits the
@@
 
-\section{Transactions and Epochs}
+\section{DAOS Layer}\label{sec:daos}
+The Distributed Asynchronous Object Storage layer serves as the traditional
+file-system interface layer for the storage devices.
+
+\section{VOSD Layer}\label{sec:vosd}
+The Versioning Object Storage Device layer operates as the layer for each
+persistent storage device used to support the parallel storage array.
+
+\section{Broader Design}
+\label{sec:broader}
+
+At a broader level, there are some concerns that were partially clarified
+through conversations with the team.  Consider a shared file system across an
+HPC data center. The current design maintains the metadata in its own
+container. Since copying data from the IOD layer to the DAOS layer requires an
+explicit persist call, how and when synchronizing the metadata across the
+layers and potentially across machines occurs is undefined. Delaying
+synchronization until an explicit persist is called will reduce the update
+frequency, but delays the data visibility on other platforms. Ideally, the
+metadata object would need to be automatically persisted every time a container
+transaction is persisted to the DAOS layer.
+
+The implication of this is that every transaction persist is double operation
+to account for the metadata persist. More importantly, the IOD-layer version of
+the metadata container may contain readable transactions that have not been
+persisted to the DAOS layer. How to handle this inconsistency between the two
+layers still needs to be explored.
+
+A point of confusion rather than a potential design challenge is the change in
+definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
+container is a collection of objects. For the DAOS layer, a container is a
+collection of objects across a set of shards. For the IOD an object may be a
+shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
+Having the same names with locally correct, globally conflicting definitions
+serves to confuse how the system should work.
+
+\subsection{Transactions and Epochs}
 \label{sec:transactions}
@@
 
-\subsection{IOD Transactions}
+\subsubsection{IOD Transactions}
 To understand how transactions are used in the IOD layer, some terminology and
@@
 
-\subsubsection{Design Philosophy}
+\noindent\textbf{Design Philosophy}
+
 Undocumented, but inherent in the design of these transactions is how faults
@@
 
-\subsection{DAOS Epochs}
+\subsubsection{DAOS Epochs}
 The Epoch mechanism differs from transactions. Instead of focusing on when a
@@
 
-\subsubsection{Design Philosophy}
+\noindent\textbf{Design Philosophy}
+
 While the metadata design is not fully defined, there are a few things that
@@
 
-\section{Broader Design}
-\label{sec:summary}
-
-At a broader level, there are some concerns that were partially clarified
-through conversations with the team.  Consider a shared file system across an
-HPC data center. The current design maintains the metadata in its own
-container. Since copying data from the IOD layer to the DAOS layer requires an
-explicit persist call, how and when synchronizing the metadata across the
-layers and potentially across machines occurs is undefined. Delaying
-synchronization until an explicit persist is called will reduce the update
-frequency, but delays the data visibility on other platforms. Ideally, the
-metadata object would need to be automatically persisted every time a container
-transaction is persisted to the DAOS layer.
-
-The implication of this is that every transaction persist is double operation
-to account for the metadata persist. More importantly, the IOD-layer version of
-the metadata container may contain readable transactions that have not been
-persisted to the DAOS layer. How to handle this inconsistency between the two
-layers still needs to be explored.
-
-A point of confusion rather than a potential design challenge is the change in
-definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
-container is a collection of objects. For the DAOS layer, a container is a
-collection of objects across a set of shards. For the IOD an object may be a
-shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
-Having the same names with locally correct, globally conflicting definitions
-serves to confuse how the system should work.
+\subsection{Function and Analysis Shipping}
+\label{sec:fn-shipping}
+
+In terms of the CN-ION communication model, a client/server architecture
+is implemented: every ION runs an IOFSL (I/O Function Shipping
+Layer) server; the IOFSL client is integrated into the HDF5
+library which runs on each CN. A client can forward requests to any
+number of IONs. Every I/O operation issued by HDF5 is asynchronously
+shipped to the IOFSL server and asynchronously executed.
+
+\subsection{ACG}\label{acg}
+
+This will exemplify how applications make use of the Exascale stack. For
+FF, support for
+GraphLab 
+and GraphBuilder will be prototyped, which are graph-processing
+frameworks. GraphBuilder is a set of MapReduce tasks that extract,
+normalize, partition and serialize a graph out of unstructured data, and
+writes graph-specific formats into HDFS. These files are later consumed
+by GraphLab, a vertex-centric, asynchronous execution engine that runs
+directly on top of HDFS (i.e., non-MapReduce). The following illustrates
+the architecture of both frameworks:
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
+\caption{GraphLab and GraphBuilder stacks}
+\label{fig:graphlab-graphbuilder}
+\end{figure}
+
+In order to make both work on top of the exascale stack, both have to be
+modified. After these modifications are implemented, GraphBuilder will
+be able to write the partitioned graph in (the newly proposed) HDF5
+files which will thus be stored in the IOD nodes (or IONs) in a
+parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
+the library to perform at high speeds by benefiting from the new
+features (see next section). In general both frameworks will be modified
+so that calls to HDFS-based formats are replaced by the proposed HDF5
+ones. This is referred to as the HDF Adaptation Layer or HAL and will
+provide, from the GraphBuilder/GraphLab point of view:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  capability for storing the newly proposed HDF5 format
+\item
+  association of network information to vertices/edges
+\item
+  shipping computation to the IONs
+\item
+  asynchronous vertex updates
+\item
+  efficient data sharing ammong CNs
+\item
+  computation over versioned datasets
+\end{itemize}
 

### 2014-04-30 paper.tex
minor fix to correctly reference IOD VOL image
added some to the related work and fixed up the intro some more. added some global images

@@
 centralized storage will better support the performance and energy requirements
-of extreme scale integrated application workflows. At a more detailed view, the
-various layers of the IO stack each contribute different functionality and
-performance implications.
+of extreme scale integrated application workflows.
 
@@
 
-The architecture incorporates five layers, some of which are potentially
-optional. The top layer is a high level IO library, such as the demonstration
-HDF-5 library~\cite{hdf5}. The intent is to only have access to the storage
-stack through such an API to manage the complexity of working with the lower
-layers and to enable advanced functionality that would require more direct user
-intervention. This layer incorporate the necessary features for ACGs from a
-end-user's perspective.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\columnwidth]{images/arch-mapping}
+\caption{Target Architecture and Component Mapping}
+\label{fig:arch-mapping}
+\end{figure}
+
+At a more detailed view, the various layers of the IO stack each contribute
+different functionality and performance implications.  The
+architecture (Figure~\ref{fig:arch-mapping}) incorporates five layers, some of
+which have potentially optional components.  The top layer is generally a high
+level IO library, such as the demonstration HDF-5 library~\cite{hdf5}. It is in
+dark blue. The intent is to only have access to the storage stack through such
+an API to manage the complexity of working with the lower layers and to enable
+advanced functionality that would require more direct user intervention. This
+layer incorporate the necessary features for ACGs from a end-user's
+perspective.
 
 Below the user API is an IO forwarding layer that redirects IO calls from the
-compute nodes to the IO dispatching layer.  This IO forwarding layer is
-analogous to the function of the IO nodes in a BlueGene machine or the passive
-data staging processes demonstrated
+compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
+is analogous to the function of the IO nodes in a BlueGene machine or the
+passive data staging processes demonstrated
 previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
@@
 storage interface for the IO stack and is the only way to reach the persistent
-storage array in lower layers. Ideally, the IOD layer's functionality can be
-optional based on available hardware. Much of the functionality offered at this
-layer would shift either up or down the stack as discussed in detail below. The
-Distributed Asynchronous Object Storage (DAOS) layer serves as the persistent
-storage interface and is intended to be the traditional file system-like
-foundation on which everything else is built with no dependence on any
-technologies specified above it. For example, the IOD layer with or without
-burst buffers is not required for DAOS to operate properly.  Instead, the DAOS
-layer can handle all of the IO operations from the user API layer, albiet with
-the potential performance penalty of manipulating the shared, persistent
-storage array. At the bottom is the Versioning Object Storage Device (VOSD).
-It serves as the interface for storing objects of all types efficiently. Think
-of this layer as the physical disk interface layer. In terms of Lustre, this
-would replace the file system on individual storage targets with an interface
-friendlier to the containers of objects concept used in the higher layers.
+storage array in lower layers (in green). Ideally, the IOD layer's
+functionality can be optional based on available hardware. Much of the
+functionality offered at this layer would shift either up or down the stack as
+discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
+layer serves as the persistent storage interface and is intended to be the
+traditional file system-like foundation on which everything else is built with
+no dependence on any technologies specified above it (in pink and yellow). For
+example, the IOD layer with or without burst buffers is not required for DAOS
+to operate properly.  Instead, the DAOS layer can handle all of the IO
+operations from the user API layer, albiet with the potential performance
+penalty of manipulating the shared, persistent storage array. At the bottom is
+the Versioning Object Storage Device (VOSD) (in purple).  It serves as the
+interface for storing objects of all types efficiently. Think of this layer as
+the physical disk interface layer. In terms of Lustre, this would replace the
+file system on individual storage targets with an interface friendlier to the
+containers of objects concept used in the higher layers.
 
@@
 information represents information that may or may not have been written down,
-but is the intent of ultimate product. These insights were gained based on
-personal conversations with the team
-members~\cite{JohnBent,QuinceyKoziol,EricBarton} understanding the current
-design as written. These ideas are presented to
-give a fuller picture of where the project is going rather than dwelling on the
+but is the intent of ultimate product.  These ideas are presented to give a
+fuller picture of where the project is going rather than dwelling on any
 limitations of the published documents.
@@
 The major previous work is really limited to full file systems rather than the
-mountain of file system adjustments made over the years. The major players
-related to this work include Ceph, PVFS, Lustre, GPFS, and Panasas.
-
-EXPLAIN A BIT ABOUT HOW EACH CANNOT ADDRESS THE PROBLEMS WITH A CITATION FOR
-EACH. DID I MISS A MAJOR FILE SYSTEM OF NOTE?
+mountain of file system adjustments made over the years.
+
+Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
+both a POSIX and object interface including features typically found in parallel
+file systems. Ceph's unique striping approach uses pseudo-random numbers with a
+known seed eliminating the need for the metadata service to track where each
+stripe in a parallel file is placed.
+
+PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
+such as a single process opening a file and sharing the handle.
+
+Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
+clusters offering scalable performance and fine-grained end-user and programmatic control over how data is placed in the storage system.
+Scalability limitations are part of the motivation for the FFSIO project.
+
+GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good performance for scaling parallel IO tasks and is used extensively by its owner, IBM.
+
+Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive striping system that detects the need for additional stripes for performance and adjusts the file layout as necessary.
+
+Other file systems, like GoogleFS~\cite{ghemawat:googlefs}, address distributed
+rather than parallel computing and cannot be compared directly. The primary
+difference between distributed and parallel file systems is the ability of the
+file system to retrieve data simultaneously from multiple clients, in parallel,
+and treat the resulting collection of pieces as a single object. Distributed
+file systems rely on a single client creating a file, but distributing the set
+of files across a wide array of storage devices. The other, popular distributed
+file system of note is HDFS~\cite{Shvachko:2010:hdfs} that is distributed as
+part of Hadoop. These other file systems are mainly of interest in the context
+of the ACG features of FFSIO and will be discussed more in
+Section~\ref{sec:acg}.
 
@@
   explains the relationship between them. Every file begins with a root
-  group ''/'', which serves as the ''starting-point'' in the object
+  group ``/'', which serves as the ``starting-point'' in the object
   hierarchy.
@@
 
-\section{IOD Layer}
+\section{IO Dispatcher Layer}
 \label{sec:iod}
@@
 
-%Current NAND-based flash devices top out at around 400 MB/sec.
-%The key spec that is missing from this number is that 400 MB/sec is a measure
-%of the fixed number of available IOPS multiplied by the block size. This
-%represents the ideal streaming performance possible. The problem is that it
-%costs an IOP to read 1 byte or 1 block (4KB or 8KB, depending on the device).
-%It costs 1 IOP to write a full block--usually. In some cases, it will cost 2
-%IOPs.  This accounts for the required pre-erase write prior to writing to a
-%reused block. In the worst case, it can be 3 IOPS per write. This would be for
-%a partial block write (read the old block, erase, write the modified block).
-%One-third of 400 MB/sec, about 133 MB/sec, is well below the streaming
-%performance of HDDs.  Granted, there is still rotational and seek latency to
-%deal with for HDDs, but the advantage for SSDs has evaporated and potentially
-%turned into a penalty at a considerable cost premium.  There are faster SSD
-%solutions on the market that incorporate DRAM for caching and using the PCIe
-%bus, for example, but their price precludes them from use in an extreme scale
-%platform.
-
 Given these features, the optionality and even incorporation of burst buffers
@@
 explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
-%The
-%breakdown of the number of participants in each role is shown in
-%Table~\ref{tab:scaling}. For comparison, consider the Number of
-%Sub-Coordinators equivalent to IOD processes. The Processes Per Sub-Coordinator
-%represents the number of clients that use a particular IOD process. For these
-%tests, we maintained a balanced distribution and always used at least two
-%sub-coordinators to slow down the processing.
 
@@
 \end{figure}
-
-%\begin{table}[ht]
-%    \vspace{-0.15in}
-%    \centering
-%    \caption[Scaling Configuration]{Performance Tests Scaling Configuration}
-%    \bigskip
-%    \vspace{-0.15in}
-%
-%    \begin{tabular}{|r|r|r|}
-%\hline
-%Processes & \vtop{\hbox{\strut Number of}\hbox{\strut Sub-Coordinators}} & \vtop{\hbox{\strut Processes Per} \hbox{\strut Sub-Coordinator}}\\
-%\hline
-%%8 & 2 & 4 \\
-%%16 & 2 & 8 \\
-%%32 & 2 & 16 \\
-%%64 & 2 & 32 \\
-%%128 & 2 & 64 \\
-%256 & 2 & 128 \\
-%512 & 2 & 256 \\
-%1024 & 4 & 256 \\
-%2048 & 8 & 256 \\
-%4096 & 16 & 256 \\
-%8192 & 32 & 256 \\
-%16384 & 64 & 256 \\
-%32768 & 128 & 256 \\
-%65536 & 256 & 256 \\
-%\hline
-%    \end{tabular}
-%    \label{tab:scaling}
-%\end{table}
 
@@
 
-%\subsection{Comparison to Other Protocols}
-%Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
-%ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
-%unsuitable for this environment. First, the distributed servers in Paxos
-%systems are all distributed copies of each other that eventually become
-%consistent. Given the scale we wish to address, a single node's memory is
-%unlikely to be able to hold all of the data necessary for many operations at
-%scale. They also do not have a guarantee for when consensus will be achieved
-%without using slower synchronous calls. For the tight timing we wish to
-%support, we need guarantees of when a consistent state has been achieved.
-%Second, these systems also all assume that updates are initiated from a single
-%client process rather than a parallel set of processes as is the standard in
-%HPC environments.
-%
-%\DDT uses a second layer of coordination on the client side that greatly
-%increases the scalability by consolidating messages from clients into unique
-%sets prior to sending to the overall coordinator. A gossip
-%protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
-%purpose, but the delay of eventual consistency is strictly avoided with this
-%protocol to ensure guarantees at particular states in the code. For example, if
-%a failure occurs, the global understanding of the role of all processes is
-%required in order for effective communication to occur for operations like
-%creating sub-transactions or voting. In this case, the protocol can offer
-%stronger statements about consistency than these protocols offer.  These
-%features offer a way to easily scale the transaction protocol given the
-%guarantees we wish to offer. The IOD approach does nothing to address these
-%concerns.
-%
-%Another effort to offer consistency and data integrity for the ZFS file
-%system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
-%focus on the processes all having a notion of completion as a transaction, this
-%work focuses on the integrity of the data movement operations. We view this
-%work as something that should be considered hand-in-hand with a transaction
-%approach to ensure the integrity of the movement of the data in addition to the
-%agreement of processes about the successful completion of a parallel operation.
-
 \subsection{Function and Analysis Shipping}
@@
 
-\subsection{ACG}\label{acg}
+\subsection{ACG}
+\label{sec:acg}
 

### 2014-05-01 paper.tex
filled out DAOS and VOSD sections and added a demonstration section.

@@
 multiple layers, it makes more sense to discuss them independently once an
-understanding of the overall structure has been presented.  The paper is
-concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
+understanding of the overall structure has been presented.  A demonstration
+of the functionality is presented in Section~\ref{sec:evaluation}. This shows
+that the prototype functionality validation prototype can function. The paper
+is concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
 covered in the paper.
@@
 
-The IOD VOL plugin serves as the bridge between HDF5 and the IOD API. It
-is also shown in Figure \ref{vol-arch}. The application calls the HDF5
+The IOD VOL plugin serves as the bridge between HDF5 and the IOD API (Figure \ref{fig:vol-arch}). The application calls the HDF5
 library while running on the system's compute nodes. Using the VOL
@@
 
-This section will briefly describe the limited functionality of the IO
-Forwarding layer and why it was included in the design.  It will not exceed
-probably half of a column.
+The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
+the massive process count on the storage stack. A current trend mixing MPI with
+a threading library like OpenMP is addressing the same issue. In this case,
+the storage stack has far fewer end-points in which to receive requests and
+data from compute processes. By reducing the number of simultaneous requests,
+delays can be reduced. This has been demonstrated for the file open operation
+with Lustre~\cite{lofstead:2009:adaptable} and to some degree for accessing the
+storage devices themselves~\cite{lofstead:2010:io-variability}. The BlueGene
+platform incorporated dedicated hardware to perform this role. The proposed
+functionality for this layer has no special features beyond those previously
+demonstrated.
 
@@
 
-In current proposals, the stack is managed by middleware that sits between the
-application and the parallel file system. Since most of the applications assume
-a POSIX interface to storage, existing middleware either transparently handle
-the I/O operations, appearing as regular POSIX calls to applications; or modify
-the I/O API as little as possible in order to minimize the impact on production
-codes.
-
-As we move towards the exascale goal, this way of interfacing with storage
-becomes more of an obstacle. Many features provided by the stack are hidden by
-the middleware/POSIX layers. If these were visible to the applications, and if
-the applications were able to control how the underlying layers behave, they
-would benefit greatly since the domain knowledge that is available at the
-application level could be used to execute storage operations efficiently. One
-more implication of the 25-year-old POSIX API is in cumbersome middleware,
-since metadata has to be part of the same uni-dimensional byte array being
-stored.
-
-The Fast Forward Storage I/O project is aimed at merging the features of
-existing middleware into a next-generation parallel file system stack. This new
-storage API termed the I/O dispatcher (IOD) is object-based, transactional,
-asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).
+%In current proposals, the stack is managed by middleware that sits between the
+%application and the parallel file system. Since most of the applications assume
+%a POSIX interface to storage, existing middleware either transparently handle
+%the I/O operations, appearing as regular POSIX calls to applications; or modify
+%the I/O API as little as possible in order to minimize the impact on production
+%codes.
+%
+%As we move towards the exascale goal, this way of interfacing with storage
+%becomes more of an obstacle. Many features provided by the stack are hidden by
+%the middleware/POSIX layers. If these were visible to the applications, and if
+%the applications were able to control how the underlying layers behave, they
+%would benefit greatly since the domain knowledge that is available at the
+%application level could be used to execute storage operations efficiently. One
+%more implication of the 25-year-old POSIX API is in cumbersome middleware,
+%since metadata has to be part of the same uni-dimensional byte array being
+%stored.
+%
+%The Fast Forward Storage I/O project is aimed at merging the features of
+%existing middleware into a next-generation parallel file system stack. This new
+%storage API termed the I/O dispatcher (IOD) is object-based, transactional,
+%asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).
 
@@
 
-The prototype stack, from a top-down point of view, has the following
-components:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Arbitrarily Connected Graphs (ACG).
-\item
-  HDF5 extensions.
-\item
-  I/O Dispatcher (IOD).
-\item
-  Distributed Asynchronous Object Storage (DAOS).
-\item
-  Versioning Object Storage Device (VOSD).
-\end{enumerate}
-
-Applications running on the compute nodes will be written in Python scripts or
-in C++ using the HDF5 APIs. These applications will execute analysis on
-graph-based data (ACG). The point is to demonstrate both the HPC and BigData
-use cases of the Exascale architecture. The data structures will be stored in
-HDF5, for which new API extensions will be implemented. These new API calls
-will expose the transactional, asynchronous and function-forwarding semantics
-of the underlying stack.
+%The prototype stack, from a top-down point of view, has the following
+%components:
+%
+%\begin{enumerate}
+%\def\labelenumi{\arabic{enumi}.}
+%\itemsep1pt\parskip0pt\parsep0pt
+%\item
+%  Arbitrarily Connected Graphs (ACG).
+%\item
+%  HDF5 extensions.
+%\item
+%  I/O Dispatcher (IOD).
+%\item
+%  Distributed Asynchronous Object Storage (DAOS).
+%\item
+%  Versioning Object Storage Device (VOSD).
+%\end{enumerate}
+%
+%Applications running on the compute nodes will be written in Python scripts or
+%in C++ using the HDF5 APIs. These applications will execute analysis on
+%graph-based data (ACG). The point is to demonstrate both the HPC and BigData
+%use cases of the Exascale architecture. The data structures will be stored in
+%HDF5, for which new API extensions will be implemented. These new API calls
+%will expose the transactional, asynchronous and function-forwarding semantics
+%of the underlying stack.
 
@@
 nodes.
-
-We give a high-level description of each of layer.
 
@@
 
-One of the bigger concerns is the observation that the original data staging
-proposals all used compute nodes while the newer proposals seek not only to
-make them a fixed portion of the IO stack, but also shared across all machine
-users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
-potential costs and advantages of where to place operators similar to the IOD
-proposed function shipping. There are two key takeaways from PreDatA. First,
-placement matters.  Depending on the communication intensity vs. computation
-intensity, where along the IO path to place the operation can matter
-significantly. Second, and more importantly, the amount of time spent
-processing for the operators was stretched to the point where it consumed
-nearly all of the time between IO operations. The given ratios of compute
-processes to staging process examined is representative for future extreme
-scale platforms. If anything, the ratios offer more staging processes than IOD
-processes would be available.
-
+%One of the bigger concerns is the observation that the original data staging
+%proposals all used compute nodes while the newer proposals seek not only to
+%make them a fixed portion of the IO stack, but also shared across all machine
+%users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
+%potential costs and advantages of where to place operators similar to the IOD
+%proposed function shipping. There are two key takeaways from PreDatA. First,
+%placement matters.  Depending on the communication intensity vs. computation
+%intensity, where along the IO path to place the operation can matter
+%significantly. Second, and more importantly, the amount of time spent
+%processing for the operators was stretched to the point where it consumed
+%nearly all of the time between IO operations. The given ratios of compute
+%processes to staging process examined is representative for future extreme
+%scale platforms. If anything, the ratios offer more staging processes than IOD
+%processes would be available.
+%
 In the case of the written IOD design, it describes a fixed-sized staging area
-that is partitioned on a per-application basis.  This is unlikely to be useful
-because of the limited compute and communication capacity to spare to perform
-these operations at a bottleneck in the IO path.  The use of a separate data
-staging area intentionally separate from the IO path allows using operators on
-limited resources leaving the IO path clear for strictly data movement. A
-nuance of this design is discussed in the Design Philosophy below.
-
-By concentrating the Burst Buffers and function shipping into the storage
-stack, three problems arise.  First, the amount of network bandwidth, IO
-bandwidth, and compute power consumed for example operations from a single
-application is likely to completely monopolize the IOD processes. Second, if
-space and time partitioning is used instead, the functionality risks being too
-small to be useful. Third, the long-term hardware performance advantage for
-SSDs is questionable. Recent studies have shown that the erase-before-write
-and interference between reading and writing with flash-based SSDs can cause
-severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
-of an optional use SSD layer in the new Trinity machine at Los Alamos will
-offer a test bed to determine how likely these observed problems would affect a
-production extreme scale platform.
-
-Given these features, the optionality and even incorporation of burst buffers
-in the current design should be carefully considered. Much of the advanced, key
-functionality proposed as they are currently designed ultimately relies on the
-existence of burst buffers to work. Further thought about how to have an IOD
-layer both with and without a burst buffer is required before they can be
-considered optional. As the design stands today, they are a required part of
-the IOD layer for proper functioning. Unfortunately, it is not clear that they
-can address the performance concerns they are intended to cover.
+that is partitioned on a per-application basis.
+%This is unlikely to be useful
+%because of the limited compute and communication capacity to spare to perform
+%these operations at a bottleneck in the IO path.  The use of a separate data
+%staging area intentionally separate from the IO path allows using operators on
+%limited resources leaving the IO path clear for strictly data movement. A
+%nuance of this design is discussed in the Design Philosophy below.
+%
+%By concentrating the Burst Buffers and function shipping into the storage
+%stack, three problems arise.  First, the amount of network bandwidth, IO
+%bandwidth, and compute power consumed for example operations from a single
+%application is likely to completely monopolize the IOD processes. Second, if
+%space and time partitioning is used instead, the functionality risks being too
+%small to be useful. Third, the long-term hardware performance advantage for
+%SSDs is questionable. Recent studies have shown that the erase-before-write
+%and interference between reading and writing with flash-based SSDs can cause
+%severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
+%of an optional use SSD layer in the new Trinity machine at Los Alamos will
+%offer a test bed to determine how likely these observed problems would affect a
+%production extreme scale platform.
+%
+%Given these features, the optionality and even incorporation of burst buffers
+%in the current design should be carefully considered. Much of the advanced, key
+%functionality proposed as they are currently designed ultimately relies on the
+%existence of burst buffers to work. Further thought about how to have an IOD
+%layer both with and without a burst buffer is required before they can be
+%considered optional. As the design stands today, they are a required part of
+%the IOD layer for proper functioning. Unfortunately, it is not clear that they
+%can address the performance concerns they are intended to cover.
 
@@
 
-\section{DAOS Layer}\label{sec:daos}
+\section{DAOS Layer}
+\label{sec:daos}
+
 The Distributed Asynchronous Object Storage layer serves as the traditional
-file-system interface layer for the storage devices.
+parallel file-system interface layer for the storage devices. This is the
+consistent, global view of the underlying devices represented in this stack
+by the VOSD layer.
+
+This is the layer where the container/object model is translated into the
+physical storage requirements dictated by the physical storage underneath (the
+VOSD layer). The two key design elements of this layer are the handling of
+epochs and the mapping of conatiners and objects to the underlying storage.
+
+Epochs are represented throguh the shard mechanism. Each container shard is
+versioned according to the transactions in the IOD layer that are persisted to
+the DAOS layer as an epoch. The container shard is a piece of the container
+with the contained objects concept. Since the container is largely logical,
+this generally maps to a shard of an object within a container.
+
+Each container is represented by a directory on some storage device containing
+symbolic links to all of the shards it contains and maintains the epoch ID.
+In particular the Highest Committed Epoch (HCE) is an important concept for
+quickly identifying which version of a shard to retrieve and to block writes to
+older epochs since those have been committed.
 
 \section{VOSD Layer}\label{sec:vosd}
-The Versioning Object Storage Device layer operates as the layer for each
-persistent storage device used to support the parallel storage array.
+The Versioning Object Storage Device (OSD) layer operates as the layer for each
+persistent storage device used to support the parallel storage array. In the
+purest form, it uses a local file system to arrange storage of objects that
+represent parts of the higher level objects in containers.
+
+The base level implementation continues the space optimization of only storing
+changes for new versions by using a copy-on-write file system. The prototype
+uses ZFS~\cite{zhang:2010:zfs} for the known stability and integration with
+Lustre. In a production version of the FFSIO stack,
+btrfs~\cite{rodeh:2013:btrfs}, The Linux B-Tree File System, given its
+open-source backing and GPL licensing, is a likely long-term choice.
+
+At a more detailed level, the design for VOSD is an increment beyond the
+current Lustre Object Storage Device design to incorporate the idea of shards
+and the versioning aspects of transactions/epochs. For every DAOS shard, the
+VOSD has information for storing and accessing the currently commited version,
+the Highest Committed Epoch, as well as a staging dataset represeting the
+next version of the object being stored. Both of these are combined in a shard
+root.
+
+For data integrity, an intent log is maiained as part of the underlying file
+system enabling fault recovery.
+
+Beyond the functionality to incorporate and expose the copy-on-write nature
+of the underlying file system and the semantics for storing and processing
+shards and their associated epochs, this is largely an evolution of the
+existing Lustre OSD layer.
 
@@
 
-At a broader level, there are some concerns that were partially clarified
-through conversations with the team.  Consider a shared file system across an
-HPC data center. The current design maintains the metadata in its own
-container. Since copying data from the IOD layer to the DAOS layer requires an
-explicit persist call, how and when synchronizing the metadata across the
-layers and potentially across machines occurs is undefined. Delaying
-synchronization until an explicit persist is called will reduce the update
-frequency, but delays the data visibility on other platforms. Ideally, the
-metadata object would need to be automatically persisted every time a container
-transaction is persisted to the DAOS layer.
-
-The implication of this is that every transaction persist is double operation
-to account for the metadata persist. More importantly, the IOD-layer version of
-the metadata container may contain readable transactions that have not been
-persisted to the DAOS layer. How to handle this inconsistency between the two
-layers still needs to be explored.
-
+%At a broader level, there are some concerns that were partially clarified
+%through conversations with the team.  Consider a shared file system across an
+%HPC data center. The current design maintains the metadata in its own
+%container. Since copying data from the IOD layer to the DAOS layer requires an
+%explicit persist call, how and when synchronizing the metadata across the
+%layers and potentially across machines occurs is undefined. Delaying
+%synchronization until an explicit persist is called will reduce the update
+%frequency, but delays the data visibility on other platforms. Ideally, the
+%metadata object would need to be automatically persisted every time a container
+%transaction is persisted to the DAOS layer.
+%
+%The implication of this is that every transaction persist is double operation
+%to account for the metadata persist. More importantly, the IOD-layer version of
+%the metadata container may contain readable transactions that have not been
+%persisted to the DAOS layer. How to handle this inconsistency between the two
+%layers still needs to be explored.
+%
 A point of confusion rather than a potential design challenge is the change in
@@
 these offers different functionality, but are connected as is explained below.
-How these differ from the \DDT approach is also explored.  While IOD's and
-\DDTns's transactions are seemingly very different, they use a similar
-high-level design, but very different implementation, to solve the same
-problem.
+%How these differ from the \DDT approach is also explored.  While IOD's and
+%\DDTns's transactions are seemingly very different, they use a similar
+%high-level design, but very different implementation, to solve the same
+%problem.
 
@@
 
-Based on the lessons from the \DDT metadata
-service~\cite{lofstead:2012:txn-metadata} construction and the prior
-experiments with LWFS, having a completely separate metadata service is
-workable. Rather than making it a bottleneck in the IO path, it is another
-service that users must interact with if they need those services.  Users can
-manage everything by maintaining the metadata including the list of objects
-themselves. However, there are drawbacks to this 100\% client-side approach.
-
-With a client-side only approach, there is a serious risk of the metadata
-service and the object store getting out of sync.  While having a metadata-less
-object storage service is desirable, the different semantics from traditional
-file systems requires some considerations. In this case, should these services
-get out of sync, three particular risks are introduced.  First, a client could
-create a dangling entry in the metadata service that does not correspond to any
-objects in the object store. Second, a client could create orphaned objects
-that have no associated metadata entries. Third, updates to the metadata or
-object store service should be an atomic operation, but due to a lack of
-coordination, a window where the system is inconsistent appears.
-
-Ultimately, the consistency semantics required must be determined. If a
-metadata service is required and it must be in sync with the object storage
-service, then additional work must be performed. In traditional file systems,
-the metadata and object storage updates are atomic. With decoupling metadata
-from object storage, should this atomicity still be desired, it requires both
-the ability for the services to participate in a task that is part of a larger
-atomic operation and a higher-level mechanism to manage the atomic operation.
-
-Overall, while additional work is required to maintain a client-side only
-metadata service, it eliminates any potential bottlenecks related to updating
-metadata related to the object storage. The burden of tracking striping and
-other metadata that has traditionally been part of the metadata associated with
-the file system will have to be maintained by the object storage service. The
-lack of a centralized, serialized bottleneck to store that information improves
-concurrency.
-
-\subsection{Comparison to \DDTns}
-The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
-approach for handling ACID-style transactions in an environment with parallel
-clients and multiple servers (doubly distributed). Rather than being aimed
-solely at data movement operations, \DDT seeks to address the general problem
-of managing any operation with multiple clients and servers.  Consider the
-management of the analysis/visualization area, potentially similar to the IOD
-concept. The transaction protocol is used to help manage resizing of the
-resource allocation to the various analysis and visualization components.  For
-the purposes of this discussion, \DDT could also be used to manage changing how
-IOD processes and/or nodes are used without exposing these changes to the
-client processes prematurely.  This has been described and analyzed
-previously~\cite{dayal:2013:io-containers}.
-
-The example metadata and data storage services created as part of the \DDT
-project have no dependencies between transactions that prevent visibility
-should an older version be incomplete. This additional, intentional requirement
-by IOD offers different functionality than \DDTns's example services. In the
-case of \DDTns, the functionality is more minimal, but also avoids some of the
-concerns outlined below.
-
-The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
-scalability issues and demonstrated a scalable client-side coordination model
-with excellent performance. The performance measured for a complex transaction
-with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
-explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
-
-\begin{figure}[ht]
+%Based on the lessons from the \DDT metadata
+%service~\cite{lofstead:2012:txn-metadata} construction and the prior
+%experiments with LWFS, having a completely separate metadata service is
+%workable. Rather than making it a bottleneck in the IO path, it is another
+%service that users must interact with if they need those services.  Users can
+%manage everything by maintaining the metadata including the list of objects
+%themselves. However, there are drawbacks to this 100\% client-side approach.
+%
+%With a client-side only approach, there is a serious risk of the metadata
+%service and the object store getting out of sync.  While having a metadata-less
+%object storage service is desirable, the different semantics from traditional
+%file systems requires some considerations. In this case, should these services
+%get out of sync, three particular risks are introduced.  First, a client could
+%create a dangling entry in the metadata service that does not correspond to any
+%objects in the object store. Second, a client could create orphaned objects
+%that have no associated metadata entries. Third, updates to the metadata or
+%object store service should be an atomic operation, but due to a lack of
+%coordination, a window where the system is inconsistent appears.
+%
+%Ultimately, the consistency semantics required must be determined. If a
+%metadata service is required and it must be in sync with the object storage
+%service, then additional work must be performed. In traditional file systems,
+%the metadata and object storage updates are atomic. With decoupling metadata
+%from object storage, should this atomicity still be desired, it requires both
+%the ability for the services to participate in a task that is part of a larger
+%atomic operation and a higher-level mechanism to manage the atomic operation.
+%
+%Overall, while additional work is required to maintain a client-side only
+%metadata service, it eliminates any potential bottlenecks related to updating
+%metadata related to the object storage. The burden of tracking striping and
+%other metadata that has traditionally been part of the metadata associated with
+%the file system will have to be maintained by the object storage service. The
+%lack of a centralized, serialized bottleneck to store that information improves
+%concurrency.
+
+%\subsection{Comparison to \DDTns}
+%The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
+%approach for handling ACID-style transactions in an environment with parallel
+%clients and multiple servers (doubly distributed). Rather than being aimed
+%solely at data movement operations, \DDT seeks to address the general problem
+%of managing any operation with multiple clients and servers.  Consider the
+%management of the analysis/visualization area, potentially similar to the IOD
+%concept. The transaction protocol is used to help manage resizing of the
+%resource allocation to the various analysis and visualization components.  For
+%the purposes of this discussion, \DDT could also be used to manage changing how
+%IOD processes and/or nodes are used without exposing these changes to the
+%client processes prematurely.  This has been described and analyzed
+%previously~\cite{dayal:2013:io-containers}.
+%
+%The example metadata and data storage services created as part of the \DDT
+%project have no dependencies between transactions that prevent visibility
+%should an older version be incomplete. This additional, intentional requirement
+%by IOD offers different functionality than \DDTns's example services. In the
+%case of \DDTns, the functionality is more minimal, but also avoids some of the
+%concerns outlined below.
+%
+%The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
+%scalability issues and demonstrated a scalable client-side coordination model
+%with excellent performance. The performance measured for a complex transaction
+%with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
+%explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
+%
+%\begin{figure}[ht]
+%%\vspace{-0.15in}
+%\centering
+%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
 %\vspace{-0.15in}
-\centering
-\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
-\vspace{-0.15in}
-\caption{Total Transaction Overhead}
-\label{fig:performance}
+%\caption{Total Transaction Overhead}
+%\label{fig:performance}
+%%\vspace{-0.15in}
+%\end{figure}
+%
+%At a high level, both \DDT and the IOD transactions have the same design. In
+%both cases, a hierarchical model is employed. In the case of \DDTns, it is a
+%purely client-side tree using semi-synchronous messaging. The messaging itself,
+%in the current implementation, uses asynchronous MPI messages. The synchronous
+%component comes from the timeout mechanism used to detect faults.  It forces a
+%level of coordination and synchronization for the protocol. For IOD, it is a
+%server-side tree and fully asynchronous relying on the existing Lustre fault
+%detection mechanism for failure detection. In both cases, there is a master in
+%charge of managing the transaction and a collection of workers that aggregate
+%into the master through second-level leaders. Beyond that, there are some
+%significant differences. Some of the different choices made by IOD raise some
+%possible concern.
+%
+%The multi-leader model introduces the possibility of forcing a rollback of an
+%entire transaction when a partial retry might be sufficient for success. Since
+%the transactions are managed at a high level rather than the individual
+%tasks, a failure in a limited distributed task can cause the entire transaction
+%to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
+%10 have an additional shared task to complete. If the task shared by the 3
+%processes fails on any of the three, the entire transaction would roll back
+%because it is a coarse-grained success/failure. If a concept like
+%sub-transactions at a task granularity were used, then it would be possible for
+%the one process that failed to report just that failure. Then the transaction
+%manager could reassign the resources for these 3 processes and try just that
+%operation again. If it now succeeds, then the overall transaction can be marked
+%successful only redoing the minimum amount of work required.
+%
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
 %\vspace{-0.15in}
-\end{figure}
-
-At a high level, both \DDT and the IOD transactions have the same design. In
-both cases, a hierarchical model is employed. In the case of \DDTns, it is a
-purely client-side tree using semi-synchronous messaging. The messaging itself,
-in the current implementation, uses asynchronous MPI messages. The synchronous
-component comes from the timeout mechanism used to detect faults.  It forces a
-level of coordination and synchronization for the protocol. For IOD, it is a
-server-side tree and fully asynchronous relying on the existing Lustre fault
-detection mechanism for failure detection. In both cases, there is a master in
-charge of managing the transaction and a collection of workers that aggregate
-into the master through second-level leaders. Beyond that, there are some
-significant differences. Some of the different choices made by IOD raise some
-possible concern.
-
-The multi-leader model introduces the possibility of forcing a rollback of an
-entire transaction when a partial retry might be sufficient for success. Since
-the transactions are managed at a high level rather than the individual
-tasks, a failure in a limited distributed task can cause the entire transaction
-to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
-10 have an additional shared task to complete. If the task shared by the 3
-processes fails on any of the three, the entire transaction would roll back
-because it is a coarse-grained success/failure. If a concept like
-sub-transactions at a task granularity were used, then it would be possible for
-the one process that failed to report just that failure. Then the transaction
-manager could reassign the resources for these 3 processes and try just that
-operation again. If it now succeeds, then the overall transaction can be marked
-successful only redoing the minimum amount of work required.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
-\vspace{-0.15in}
-\caption{Optimized Protocol}
-\label{fig:optimized-protocol}
-%\vspace{-0.25in}
-\end{figure}
-
-\DDT has addressed theses issues in a couple of ways. First, the
-sub-coordinators each have a list of processes from which they expect messages.
-Should a message be missed, it is noticed and corrective action can be taken.
-Second, \DDT has the concept of sub-transactions. The messaging requirements
-are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
-represent finer grained operations than the entire output, \DDT can manage
-multiple writes per client by using a sub-transaction to represent the output
-for any item to the file (container). Because of how the sub-transactions are
-managed, the singleton sub-transactions, ones in which only a single process
-participates, must be declared before the transaction begins so that its
-existence can be broadcast as part of the begin transaction message. This
-ensures there is global knowledge that the sub-transaction is expected.  That
-way if the coordinator (transaction leader) fails, whichever process takes over
-that role knows to expect a completion message for that sub-transaction or the
-overall transaction cannot complete. Global sub-transactions can be defined at
-any time since they are a global, synchronized operation broadcasting their
-existence. While this additional layer does introduce messaging, the overhead
-is quite small.
-
-The advantages of eliminating these messages is not performance as demonstrated
-by the performance of \DDTns. Instead, it offers a much less synchronous model
-that matches with different programming models, such as Charm++ or other
-task-based approaches. Since it can work for the bulk-synchronous model also,
-it is a more broadly applicable approach. This assumes that the observed
-potential issues can be addressed successfully.
+%\caption{Optimized Protocol}
+%\label{fig:optimized-protocol}
+%%\vspace{-0.25in}
+%\end{figure}
+%
+%\DDT has addressed theses issues in a couple of ways. First, the
+%sub-coordinators each have a list of processes from which they expect messages.
+%Should a message be missed, it is noticed and corrective action can be taken.
+%Second, \DDT has the concept of sub-transactions. The messaging requirements
+%are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
+%represent finer grained operations than the entire output, \DDT can manage
+%multiple writes per client by using a sub-transaction to represent the output
+%for any item to the file (container). Because of how the sub-transactions are
+%managed, the singleton sub-transactions, ones in which only a single process
+%participates, must be declared before the transaction begins so that its
+%existence can be broadcast as part of the begin transaction message. This
+%ensures there is global knowledge that the sub-transaction is expected.  That
+%way if the coordinator (transaction leader) fails, whichever process takes over
+%that role knows to expect a completion message for that sub-transaction or the
+%overall transaction cannot complete. Global sub-transactions can be defined at
+%any time since they are a global, synchronized operation broadcasting their
+%existence. While this additional layer does introduce messaging, the overhead
+%is quite small.
+%
+%The advantages of eliminating these messages is not performance as demonstrated
+%by the performance of \DDTns. Instead, it offers a much less synchronous model
+%that matches with different programming models, such as Charm++ or other
+%task-based approaches. Since it can work for the bulk-synchronous model also,
+%it is a more broadly applicable approach. This assumes that the observed
+%potential issues can be addressed successfully.
 
@@
 
+\section{Demonstration}
+This stack has a prototype implementation intended to test concepts rather
+than performance and scalability. It has focused on examining the interaction
+of the different APIs for each layer to flesh out any detailed requirements
+or concerns that may have been missed in the conceptualization of this IO stack.
+
+We run this demonstration using two different modes. A standard HDF-5 test
+application is used with the native file system layer writing directly to
+storage is measured. Then, the same test is run against the requirements
+prototype to demonstrate that it can function for this example as well. The
+performance of both of these tests are reported to give a very rough idea of
+the overhead that might be involved. Rather than a true overhead, this should
+be considered the maximum overhead that should be expected once an optimized,
+fully functional IO stack is deployed without relying on translating to an
+underlying parallel file system. Figure~\ref{fig:evaluation} shows the
+relative performance for both test runs.
+
+The testing evironment for this demonstration is ....
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
+\caption{Functionality Demonstration Validation}
+\label{fig:evaluation}
+\end{figure}
+
 \section{Conclusions}

### 2014-05-02 paper.tex
Lots of editing

@@
 storage array in lower layers (in green). Ideally, the IOD layer's
-functionality can be optional based on available hardware. Much of the
-functionality offered at this layer would shift either up or down the stack as
-discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
-layer serves as the persistent storage interface and is intended to be the
+functionality can be optional based on available hardware and provided on the
+IO Nodes (IONs). Much of the functionality offered at this layer would shift
+either up or down the stack as discussed in detail below. The Distributed
+Asynchronous Object Storage (DAOS) layer serves as the persistent storage
+interface and translation layer between the user-visible object model and the
+requirements of the underlying storage infrastructure. It is intended to be the
 traditional file system-like foundation on which everything else is built with
@@
 the Versioning Object Storage Device (VOSD) (in purple).  It serves as the
-interface for storing objects of all types efficiently. Think of this layer as
-the physical disk interface layer. In terms of Lustre, this would replace the
-file system on individual storage targets with an interface friendlier to the
-containers of objects concept used in the higher layers.
+interface for storing objects of all types efficiently for each storage device
+in the parallel storage array. Think of this layer as the physical disk
+interface layer. In terms of Lustre, this would replace the API on individual
+storage devices with an interface friendlier to the containers of objects and
+transactions/epochs concepts used in the higher layers.
 
@@
 fuller picture of where the project is going rather than dwelling on any
-limitations of the published documents.
+limitations of the published documents. This is most important to illustrate
+how different concepts will work across layers since that information is spread
+across multiple documents and may lack a cohesive overall view.
 
@@
 The major previous work is really limited to full file systems rather than the
-mountain of file system adjustments made over the years.
+mountain of file system refinements made over the years.
 
@@
 Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
-clusters offering scalable performance and fine-grained end-user and programmatic control over how data is placed in the storage system.
-Scalability limitations are part of the motivation for the FFSIO project.
-
-GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good performance for scaling parallel IO tasks and is used extensively by its owner, IBM.
-
-Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive striping system that detects the need for additional stripes for performance and adjusts the file layout as necessary.
+clusters offering scalable performance and fine-grained end-user and
+programmatic control over how data is placed in the storage system. Scalability
+limitations are part of the motivation for the FFSIO project.
+
+GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
+performance for scaling parallel IO tasks and is used extensively by its owner,
+IBM.
+
+Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
+striping system that detects the need for additional stripes for performance
+and adjusts the file layout as necessary.
 
@@
 difference between distributed and parallel file systems is the ability of the
-file system to retrieve data simultaneously from multiple clients, in parallel,
-and treat the resulting collection of pieces as a single object. Distributed
-file systems rely on a single client creating a file, but distributing the set
-of files across a wide array of storage devices. The other, popular distributed
-file system of note is HDFS~\cite{Shvachko:2010:hdfs} that is distributed as
-part of Hadoop. These other file systems are mainly of interest in the context
-of the ACG features of FFSIO and will be discussed more in
-Section~\ref{sec:acg}.
+file system to store and retrieve data simultaneously from multiple clients, in
+parallel, and treat the resulting collection of pieces as a single object.
+Distributed file systems rely on a single client creating a file, but
+distributing the set of files across a wide array of storage devices. The
+other, popular distributed file system of note is
+HDFS~\cite{Shvachko:2010:hdfs} that is distributed as part of Hadoop. These
+other file systems are mainly of interest in the context of the ACG features of
+FFSIO and will be discussed more in Section~\ref{sec:acg}.
 
@@
 \label{sec:end-user}
+
 Since the proposal specifies a high-level IO API will be the primary end-user
 interface for programmatically interacting with the FFSIO stack, the team used
-the HDF-5 API and leverated the Virtual Object Layer (VOL) for the initial
-design and implementation demonstration. The VOL provides a way to intercept
-HDF-5 API calls just below the API layer to implement alternative
-functionality. For this project, rather than the default writing to disk in the
-HDF-5 format, the VOL layer is used to interact with the IOD layer and the new
-concepts it offers, such as transactions, completely hidden from the user.
-This both offers a straightforward way for existing applications to test
-against the FFSIO stack and to ensure proper semantics are maintained at the
-lower levels. Some additional API functions have been introduced to support
-new functionality, such as the ACG support for Hadoop-style applications.  The
-additions comprise (Figure~\ref{fig:vol-arch}):
+the HDF-5 API and leveraged the Virtual Object Layer (VOL) for the initial
+design and implementation demonstration. This also serves as a good test
+determining what are strictly necessary extensions to an existing IO API to
+support the new functionality.  The additional functionality, such as
+transactions, can be ignored for legacy implementations, but these applications
+will not be able to take advantage of the asynchronous IO support inherent to
+the new API.  The additions comprise (Figure~\ref{fig:vol-arch}):
 
@@
 \item
-  Object-storage API based on HDF5 to support high-level data models.
-  This exposes asynchronous, transactional semantics to the application,
-  as well as end-to-end data integrity. It will also allows the usage of
-  pointer data types, passing of hints down to the storage and support
-  for asynchronous index building, maintenance and querying.
-
-\item
-  Virtual Object Layer (VOL) plugin that translates HDF5 API requests
-  from applications to IOD API calls.
+  API extensions to support new functionality provided by the FFSIO project.
+  This includes calls for managing asynchronous request lists, performing
+  asynchronous operations, creating and managing transactions, end-to-end data
+  integrity, and data type and functions to support the ACG functionality more
+  efficiently than the current API.
+
+%\item
+%  Virtual Object Layer (VOL) offers the ability to plug in new implementations
+%  for HDF-5 semantics. In this case, the HDF-5 calls are implemented using the
+%  IOD or DAOS API rather than MPI-IO or POSIX.
 
@@
   Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
-  the application developer with the capacity of sending computation down to the
-  IONs and get back results.
+  the application developer with the capability of sending computation down to
+  the IONs and get back results and perform other operations such as indexing
+  and data reorganization for more efficient retrieval.
 
@@
 
-HDF5~\cite{hdf5} is a versatile data model containing complex data objects
+HDF5~\cite{hdf5} has a versatile data model offering complex data objects
 and metadata. Its information set is a collection of datasets, groups,
@@
 \centering
-\includegraphics[scale=0.30]{images/hdf5.png}
+\includegraphics[width=\columnwidth]{images/hdf5.png}
 \caption{Sample file illustrating the hierarchical structure of HDF5 files}
@@
 \item
+  \textbf{Group}: A group is an explicit association between HDF5
+  objects. It is synonymous with directories in a file system. A group
+  could contain multiple other groups, datasets, datatypes or attributes within
+  it.
+
+\item
   \textbf{Dataset}: HDF5 datasets are objects that represent actual data
-  or content. Datasets are arrays which can have multiple dimensions. A
+  or content. Datasets are arrays with potentially multiple dimensions. A
   dataset is characterized by a dataspace and a datatype. The dataspace
@@
   elements.
-
-\item
-  \textbf{Group}: A group is an explicit association between HDF5
-  objects. It is synonymous with directories in a file system. A group
-  could contain multiple other groups, datasets or datatypes within it.
 
@@
 which is used internally by HDF5 to index its objects.
-
-The Fast Forward project is working on the adaptation of the HDF5 libraries in
-order to expose the storage stack to applications.
 
@@
 
+For this project, rather than the default writing to disk in the HDF-5 format,
+the VOL layer is used to interact with the IOD layer and the different concepts
+it offers without requiring all of the functionality be exposed to users. For
+example, the containers and objects concept is transparently mapped to the
+files and datasets existing HDF5 users are familiar with.  This reduces the
+difficulty in porting applications to use the new IO stack.
+
 \begin{figure}[htbp]
 \centering
-\includegraphics[scale=0.20]{images/vol-arch.png}
+\includegraphics[width=\columnwidth]{images/vol-arch.png}
 \caption{Architectural view of the VOL abstraction mechanism}
@@
 
-\subsection{IOD VOL plugin}
-\label{iod-vol-plugin}
-
-The IOD VOL plugin serves as the bridge between HDF5 and the IOD API (Figure \ref{fig:vol-arch}). The application calls the HDF5
-library while running on the system's compute nodes. Using the VOL
-architecture, the IOD VOL plugin uses a function shipper (RPC library)
-to forward the VOL calls to a server component running on the I/O nodes
-(IONs). At that point, the VOL calls are translated into I/O Dispatcher
-(IOD) API calls and executed at the IONs.
-
-The mapping between HDF5 objects and IOD objects is the following:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  file = container
-\item
-  dataset = array
-\item
-  group = key-value store
-\item
-  attribute = key-value store
-\end{itemize}
+%\subsection{IOD VOL plugin}
+%\label{iod-vol-plugin}
+
+The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
+(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
+running on the system's compute nodes. Using the VOL architecture, the IOD VOL
+plugin uses a function shipper (RPC library) to forward the VOL calls to a
+server component running on the IO nodes (IONs). This function shipping is the
+IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
+arrive at the IO nodes, they are translated into I/O Dispatcher (IOD) API calls
+and executed at the IONs.
+
+\subsection{HDF5 to FFSIO Mapping}
+\label{sec:hdf-to-ffsio}
+
+Since HDF5 has traditionally offered an interface focused on files and the
+internal data types, such as datasets, these concepts must be mapped onto the
+proposed FFSIO data storage concepts. This mapping is shown in Table~\ref{tab:mapping}
+
+\begin{table}[ht]
+    %\vspace{-0.15in}
+    \centering
+    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
+    \bigskip
+    %\vspace{-0.15in}
+
+    \begin{tabular}{|r|r|}
+\hline
+HDF5 & FFSIO\\
+\hline
+file & container \\
+dataset & array \\
+group & key-value store \\
+attribute(s) & key-value store \\
+\hline
+    \end{tabular}
+    \label{tab:mapping}
+\end{table}
+
+In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.
 
@@
 platform incorporated dedicated hardware to perform this role. The proposed
-functionality for this layer has no special features beyond those previously
-demonstrated.
+functionality for this layer, beyond managing the number of connections to
+the IOD layer, is to implement function shipping from the compute nodes to the
+IO nodes.
+
+For the basic HDF5 calls, this will work the same as how the Nessie
+staging~\cite{lofstead:2011:nessie-staging} shifted the collective IO data
+rearrangement calls to a reduced number of processes. The prototype
+implementation will only support accessing functionality already deployed to
+the IO nodes through an RPC mechanism. This initial implementation will use
+Mercury~\cite{Soumagne:2013:mercury} to access the remote functionality. For
+dynamically defined functions, a different system will be required leveraging
+something like C-on-demand~\cite{abbasi:2011:c-on-demand} or some other dynaimc
+deployment and compilation or an interpreter system.
 
@@
 \label{sec:iod}
+
+Strictly speaking, the IO Dispatcher layer and included functionality, such as
+burst buffers, is an optional component of the proposed IO stack. All of the
+functionality, such as function and analysis shipping, transaction management,
+and managing asynchronous data movement can be handled by other portions of the
+stack. For an extreme scale platform, the IOD layer will be an essential
+pressure relief valve for the underlying persistent storage layer. By making it
+optional, the proposed stack can be deployed more easily on smaller clusters or
+for those with more constrainted budgets. For simplicity, the rest of this
+section will describe a full stack including all of the proposed IOD
+components.
 
@@
 offloading the communication load from the compute nodes. IOD has three main
-purposes. First, if the optional burst buffer is available, it works as a fast
-cache absorbing write operations for the slower trickle out to the central
-storage array. It can also be used to retrieve objects from the central storage
-array for more efficient read operations and offers data filtering to make
-client reads more efficient.  Second, it offers the transaction mechanism for
-controlling data set visibility and to manage faults that would prevent a data
-set from being used. Third, data processing operations can be placed in the
-IOD. These operations are intended to offer data rearrangement, filtering, and
-similar operations prior to data reaching the central storage array.
+purposes. First, the burst buffers works as a fast cache absorbing write
+operations for the slower trickle out to the central storage array. It can also
+be used to retrieve objects from the central storage array for more efficient
+read operations and offers data filtering to make client reads more efficient.
+Second, it offers the transaction mechanism for controlling data set visibility
+and to manage faults that would prevent a data set from being used. These
+transactions are local to the IOD layer until persisted to the DAOS layer
+eliminating the need for burdening the persistent storage with transient data.
+Third, data processing operations can be placed in the IOD. These operations
+are intended to offer data rearrangement, filtering, and similar operations
+prior to data reaching the central storage array.
 
@@
 
-\subsection{I/O Dispatcher API}
-\label{sec:fast-forwards-io-dispatcher-api}
-
-The Fast Forward I/O and Storage project is aimed at integrating many of
-the services existing currently in middleware libraries in a
-next-generation storage stack, exposed through a non-POSIX,
-transactional, asynchronous, object-based API.
-
-From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
-interface allows the user to:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  transactions
-\item
-  asynchrony
-\item
-  objects
-\item
-  layout
-\item
-  placement
-\item
-  formats
-\end{itemize}
-
-Type of objects:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  container
-\item
-  key-value store
-\item
-  blobs
-\item
-  multi-dimensional arrays
-\end{itemize}
-
-As mentioned previously, IOD allows to store distinct types of objects.
-From here on we will limit our discussion to multidimensional arrays
-(HDF5 datasets) and leave the treatment of other object types as future
-work. A multidimensional array can be sharded among the distinct I/O
-nodes in the staging area. The IOD API supports the following
-strategies:
+%\subsection{I/O Dispatcher API}
+%\label{sec:fast-forwards-io-dispatcher-api}
+%
+%The Fast Forward I/O and Storage project is aimed at integrating many of
+%the services existing currently in middleware libraries in a
+%next-generation storage stack, exposed through a non-POSIX,
+%transactional, asynchronous, object-based API.
+%
+%From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
+%interface allows the user to:
+%
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%\item
+%  transactions
+%\item
+%  asynchrony
+%\item
+%  objects
+%\item
+%  layout
+%\item
+%  placement
+%\item
+%  formats
+%\end{itemize}
+%
+%Type of objects:
+
+\subsection{FFSIO Data Model Types}
+\label{sec:data-model}
+
+With the shift from a directories and stream-of-bytes files model to the
+container and object model, some description is required to better understand
+how these concepts are being used as well as the raw benefits.
+
+\subsubsection{Container}
+As mentioned above, the concept of a container is similar to that of a file in
+a traditional file system. However, rather than being in a directory structure,
+each container essentially is stored in a hash-space allowing direct access to
+any container without regard to the current organizational context of the
+file system. For example, there is no need to navigate a directory hierarchy
+to name a particular container.
+
+Functionally, a container plays the same role as a file in that it holds a
+collection of presumably related data intended to be accessed and manipulated
+as a unit. Since this is extended from HDF5 files, the container could also be
+viewed as a directory tree of objects where each directory entry specifies
+either a sub-directory (group) or some data or attribute.
+
+\subsubsection{Key-Value Store}
+This is the base type for the container. Since the container represents
+something akin to HDF5 files, everything is stored within a hierarchical
+namespace. The root namespace is represented by the base key-value store and
+contains a list of all of the objects for this portion of the namespace as well
+as additional key-value objects representing sub-groups for the hierarchy. Each
+of those key-value store objects works identically.
+
+Attributes are stored in a key-value store object, but use the
+multi-dimnensional array and blob objects to store the values for the
+attributes.
+
+\subsubsection{Multi-Dimensional Array}
+By treating arrays as a special case separate from blobs, additional
+opportunities are enabled. For example, by knowing that an object is an array,
+proper slicing of that array onto IO nodes can be done without involving higher
+levels of the IO stack.
+
+\subsubsection{Blob}
+All other data is stored as a stream-of-bytes without regard to the actual
+data type.
+
+\subsection{Multi-Dimensional Array Data Distribution}
+
+For both IO performance and to aid in analysis and other data processing, the
+multi-dimensional array object can be split across multiple IO nodes. Each
+piece of this array is called a {\em shard}. The IOD API supports the following
+sharding strategies:
 
@@
 the storage cluster into the I/O nodes or read them directly from the
-storage cluster. At the semantic level (HDF5), indexes can be created
-for datasets, which results in being able to read through an index
-instead of directly from the base array.
-
-All these distinct alternatives result in having many different ways of
-executing the same analysis task. In the future, we plan to extend our
-design to incorporate an optimization phase that will prepare a task by
-taking into account the distinct possible ways of executing and select
-the most promising one. In our subsequent discussions, we consider only
+storage cluster. At the semantic level (HDF5), indices can be created
+for datasets resulting in being able to read through an index instead of
+directly from the base array.
+
+All of these distinct alternatives result in having many different ways for
+executing the same analysis task.
+%In the future, we plan to extend our
+%design to incorporate an optimization phase that will prepare a task by
+%taking into account the distinct possible ways of executing and select
+%the most promising one.
+In the subsequent discussions, we consider only
 data-movement optimization, i.e., sending the analysis code as close as
@@
 \subsection{I/O Nodes}
-The ``I/O nodes'' handle requests forwarded by the scientific applications,
-potentially integrate a tier of solid-state devices to absorb the burst of
-random operations, and organize/re-format the data so that transfers to/from
-the staging area from/to the traditional parallel file system can be done more
-efficiently. It also has the capacity to executing analysis on the fresh data
-that has just been generated by simulation applications running at the compute
-nodes. As the data arrives, re-organization and data preparation can be applied
-in order to anticipate the execution of analytical tasks.
+IOD processes are hosted on the IO nodes that interface a general compute area
+with the storage array. The IO noes handle requests forwarded by the scientific
+applications, potentially integrate a tier of solid-state devices to absorb the
+burst of random operations, and organize/re-format the data so that transfers
+to/from the staging area from/to the traditional parallel file system can be
+done more efficiently. It also has the capacity to executing analysis on the
+fresh data that has just been generated by simulation applications running at
+the compute nodes. As the data arrives, re-organization and data preparation
+can be applied in order to anticipate the execution of analytical tasks.
 
@@
 \centering
-\includegraphics[scale=0.30]{images/exa-arch.png}
+\includegraphics[width=\columnwidth]{images/exa-arch.png}
 \caption{Exascale Architecture}
@@
 A common cluster configuration for this type of deployment is shown in
-Figure~\ref{fig:exa-arch}. The designated I/O nodes (IONs) are connected to the
+Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
 compute nodes (CNs) through the same fast fabric (e.g., InfiniBand), while the
 connection to the external storage cluster is through a secondary slower
-channel (e.g., 10Gb Ethernet). I/O staging thus introduces a new layer in the
-application-to-storage stack, as shown in Figure~\ref{fig:exa-stack} (left).
+channel (e.g., 10Gb Ethernet). By providing additional storage, such as SSDs on
+the IO nodes, these nodes are capable of better regulating the IO pressure on
+the underlying storage array better than simple forwaring gateways. For this
+project, using something like SSDs on the IO nodes is termed a {\em Burst
+Buffer} and is disucssed below.
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.30]{images/exa-stack.png}
-\caption{Top-to-bottom view of the component stack}
-\label{fig:exa-stack}
-\end{figure}
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics[width=\columnwidth]{images/exa-stack.png}
+%\caption{Top-to-bottom view of the component stack}
+%\label{fig:exa-stack}
+%\end{figure}
 
@@
 functionality might be incorporated and the potential benefits.  Later, these
-concepts were proposed to be incorporated as part of the IO
-stack~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
-The current Fast Forward IOD design recommends incorporating SSDs, but
-specifically lists these devices as optional. Unfortunately, not incorporating
-burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
-the IOD design currently is written assuming burst buffers. This means that the
-bulk of IO operations will only hit the IOD layer and proposed functionality,
-such as the function shipping, do not discuss the impact on the DAOS layer or
-the operators themselves should a burst buffer not be available.  Consider the
-important functionality of data rearrangement and doing things like changing
-the fast array dimension on shared, spinning media.  The design of DAOS assumes
-that it will only be involved when persisting a completed transaction and only
-for a fraction of the total transactions created.  Transaction frequency at the
-IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
-is the process of transferring the set of copy-on-write changes from
-transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
-mirrors the copy-on-write functionality, but stores the differences between
-epochs rather than individual transactions. The primary difference is that at
-the DAOS layer, data is stored in large, logically contiguous chunks to manage
-the metadata load and hopefully improve read performance.
+concepts were proposed to be incorporated into the existing IO stack
+architecture~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
+%The current Fast Forward IOD design recommends incorporating SSDs, but
+%specifically lists these devices as optional. Unfortunately, not incorporating
+%burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
+%the IOD design currently is written assuming burst buffers. This means that the
+%bulk of IO operations will only hit the IOD layer and proposed functionality,
+%such as the function shipping, do not discuss the impact on the DAOS layer or
+%the operators themselves should a burst buffer not be available.  Consider the
+%important functionality of data rearrangement and doing things like changing
+%the fast array dimension on shared, spinning media.  The design of DAOS assumes
+%that it will only be involved when persisting a completed transaction and only
+%for a fraction of the total transactions created.  Transaction frequency at the
+%IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
+%is the process of transferring the set of copy-on-write changes from
+%transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
+%mirrors the copy-on-write functionality, but stores the differences between
+%epochs rather than individual transactions. The primary difference is that at
+%the DAOS layer, data is stored in large, logically contiguous chunks to manage
+%the metadata load and hopefully improve read performance.
 
@@
 %
+
 In the case of the written IOD design, it describes a fixed-sized staging area
-that is partitioned on a per-application basis.
+that is partitioned on a per-application basis. As part of an application
+being deployed into the platform, each application will be allocated a fixed
+number of IO nodes for exclusive use during the application run. This provides
+guarantees about how much burst buffer space and processing capability will be
+available for the applications.
+
 %This is unlikely to be useful
@@
 %nuance of this design is discussed in the Design Philosophy below.
-%
+
+Future work will generalize this model to potentially support dynamic IO node
+allocation and potentially examine the possibility of oversubscription. It will
+be strictly necessary to consider shared IO nodes for cases where the number
+of deployed applications exceeds the number of IO nodes. This is also under
+consideration for future work. This first pass is focused on extreme scale
+application runs that use the vast majority of a platform rather than a
+capacity cluster where end-to-end performance is a lesser concern.
+
 %By concentrating the Burst Buffers and function shipping into the storage
@@
 compute nodes, new options for buffering data prior to being moved to
-centralized storage become available and addresses some of the concerns about
+centralized storage become available and addresses potential concerns about
 SSD performance. For example, including a small amount of Phase Change memory
@@
 
-Another concern that is acknowledged, but no thought has been applied to, is
-the requirement that a single IOD process of the set assigned to an application
-is the master for any operation.  Should the number of concurrent applications
-exceed the available nodes, sharing an IOD process will be required. The
-requirements both in terms of scheduling and resource management were cut from
-the project due to funding limitations. Since partitioning of the IOD processes
-for exclusive use by particular applications is the assumed operating mode,
-should insufficient IOD resources be available, either a job could be delayed
-or IOD resources could be reallocated from a different process could be
-redeployed for use by the new job. Handling resilience concerns for the IOD
-processes must also be address.  These sorts of considerations still need to be
-made for a full production system.
+%Another concern that is acknowledged, but no thought has been applied to, is
+%the requirement that a single IOD process of the set assigned to an application
+%is the master for any operation.  Should the number of concurrent applications
+%exceed the available nodes, sharing an IOD process will be required. The
+%requirements both in terms of scheduling and resource management were cut from
+%the project due to funding limitations. Since partitioning of the IOD processes
+%for exclusive use by particular applications is the assumed operating mode,
+%should insufficient IOD resources be available, either a job could be delayed
+%or IOD resources could be reallocated from a different process could be
+%redeployed for use by the new job. Handling resilience concerns for the IOD
+%processes must also be address.  These sorts of considerations still need to be
+%made for a full production system.
 
@@
 
-Epochs are represented throguh the shard mechanism. Each container shard is
-versioned according to the transactions in the IOD layer that are persisted to
-the DAOS layer as an epoch. The container shard is a piece of the container
-with the contained objects concept. Since the container is largely logical,
-this generally maps to a shard of an object within a container.
+There is a bit of a terminology shift between the IOD layer and the DAOS
+layer. For the IOD layer, a shard represents a portion of an object that is
+spread across potentially multiple IO nodes. For the DAOS layer, a shard
+represents the portion of a container that is spread across potentially
+multiple physical storage devices. The physical storage devices are represented
+by the VOSD layer described in Section~\ref{sec:vosd}.
+
+While transactions at the HDF5 and IOD layer use the same term, at the DAOS
+layer the terminology shifts. Instead of transactions, the term {\em epochs}
+is used instead. Rather than attempting to introduce confusion, this is intended
+to help clarify how these concepts are used at different layers of the FFSIO
+stack. In the HDF5 and IOD layer, every operation has a transaction that may
+or may not ultimately be persisted to the DAOS layer. When a transaction is
+persisted to DAOS, it is termed an epoch to reflect that this is a persistent
+version of the container. For simplicity the epoch ID is the same as the
+transaction ID that was persisted.
+
+The process of persisting a transaction to an epoch involves a process called
+``flattening''. Since this stack uses a copy-on-write approach to reduce the
+space requirement for new versions of existing files, when a transaction is
+persisted to a DAOS epoch, all of the changes between the last epoch and the
+current epoch must be combined into a single entry. The process of combining
+the changes from a series of transactions into a single update for an epoch
+is termed flattening.
+
+%Epochs are represented through the shard mechanism. Each container shard is
+%versioned according to the transactions in the IOD layer that are persisted to
+%the DAOS layer as an epoch. The container shard is a piece of the container
+%with the contained objects concept. Since the container is largely logical,
+%this generally maps to a shard of an object within a container.
+
+The current implementation has the DAOS layer map the container/object data
+model onto a directory/file data model used for most existing file systems.
+Should a fully object-based file system be deployed at the VOSD layer, this
+mapping would be unnecessary. The current projections suggest that a standard
+POSIX-like file system will likey be used at the lowest level on each storage
+device requiring the mapping at some level. To perform this mapping, DAOS
+considers the following.
 
@@
 older epochs since those have been committed.
+
+Overall, the DAOS layer serves as the shared persistent storage interface for
+the IO stack. In the case of a data center-wide storage array, the DAOS layer
+would be shared across all of the platforms with the upper layers being local
+to each individual platform.
+
+To address consistency issues between platforms, containers at the DAOS layer
+must know of every transaction. To address this, a container is updated every
+time a new transaction is created for it and closed or aborted. This ensures
+that if multiple platforms are writing to the same container sequentially that
+they will not have conflicts in the highest transaction number. The FFSIO stack
+does not support multiple applications from the same or different platforms
+using a shared DAOS layer to write to the same container at the same time.
 
@@
 %
-A point of confusion rather than a potential design challenge is the change in
-definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
-container is a collection of objects. For the DAOS layer, a container is a
-collection of objects across a set of shards. For the IOD an object may be a
-shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
-Having the same names with locally correct, globally conflicting definitions
-serves to confuse how the system should work.
+%A point of confusion rather than a potential design challenge is the change in
+%definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
+%container is a collection of objects. For the DAOS layer, a container is a
+%collection of objects across a set of shards. For the IOD an object may be a
+%shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
+%Having the same names with locally correct, globally conflicting definitions
+%serves to confuse how the system should work.
 
@@
 
-The transaction mechanism manifests in two forms. At the IOD layer, they are
-called transactions and are used to judge whether or not a set of distributed,
-asynchronous modifications across a set of related objects is complete or not.
-It is also used to control access by treating the transaction ID of committed
-transaction as a version identifier.  At the DAOS layer, they are called epochs
-and represent persisted (durable) transactions from the IOD layer. Each of
-these offers different functionality, but are connected as is explained below.
+As mentioned above, the transaction mechanism manifests in two forms. At the
+IOD layer, they are called transactions and are used to judge whether or not a
+set of distributed, asynchronous modifications across a set of related objects
+is complete or not.  It is also used to control access by treating the
+transaction ID of committed transaction as a version identifier.  At the DAOS
+layer, they are called epochs and represent persisted (durable) transactions
+from the IOD layer. Each of these offers different functionality, but are
+connected as is explained below.
 %How these differ from the \DDT approach is also explored.  While IOD's and
@@
 
-This will exemplify how applications make use of the Exascale stack. For
-FF, support for
-GraphLab 
-and GraphBuilder will be prototyped, which are graph-processing
-frameworks. GraphBuilder is a set of MapReduce tasks that extract,
-normalize, partition and serialize a graph out of unstructured data, and
-writes graph-specific formats into HDFS. These files are later consumed
-by GraphLab, a vertex-centric, asynchronous execution engine that runs
-directly on top of HDFS (i.e., non-MapReduce). The following illustrates
-the architecture of both frameworks:
+This will exemplify how applications make use of the Exascale stack. For FF,
+support for GraphLab and GraphBuilder will be prototyped, which are
+graph-processing frameworks. GraphBuilder is a set of MapReduce tasks that
+extract, normalize, partition and serialize a graph out of unstructured data,
+and writes graph-specific formats into HDFS. These files are later consumed by
+GraphLab, a vertex-centric, asynchronous execution engine that runs directly on
+top of HDFS (i.e., non-MapReduce). The following illustrates the architecture
+of both frameworks:
 
@@
 \centering
-\includegraphics[scale=0.15]{images/graphlab-and-graphbuilder.png}
+\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
 \caption{GraphLab and GraphBuilder stacks}

### 2014-05-07 paper.tex
more changes

@@
 At a more detailed view, the various layers of the IO stack each contribute
-different functionality and performance implications.  The
-architecture (Figure~\ref{fig:arch-mapping}) incorporates five layers, some of
-which have potentially optional components.  The top layer is generally a high
-level IO library, such as the demonstration HDF-5 library~\cite{hdf5}. It is in
-dark blue. The intent is to only have access to the storage stack through such
-an API to manage the complexity of working with the lower layers and to enable
-advanced functionality that would require more direct user intervention. This
-layer incorporate the necessary features for ACGs from a end-user's
-perspective.
+different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
+incorporates five layers, some of which have potentially optional components.
+The top layer is generally a high level IO library, such as the demonstration
+HDF-5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
+access to the storage stack through such an API to manage the complexity of
+working with the lower layers and to enable advanced functionality that would
+require more direct user intervention. This layer incorporates the necessary
+features for ACGs from a end-user's perspective.
 
@@
 have considerable functionality. The IO dispatcher (IOD) serves as the primary
-storage interface for the IO stack and is the only way to reach the persistent
-storage array in lower layers (in green). Ideally, the IOD layer's
-functionality can be optional based on available hardware and provided on the
-IO Nodes (IONs). Much of the functionality offered at this layer would shift
-either up or down the stack as discussed in detail below. The Distributed
-Asynchronous Object Storage (DAOS) layer serves as the persistent storage
-interface and translation layer between the user-visible object model and the
-requirements of the underlying storage infrastructure. It is intended to be the
-traditional file system-like foundation on which everything else is built with
-no dependence on any technologies specified above it (in pink and yellow). For
-example, the IOD layer with or without burst buffers is not required for DAOS
-to operate properly.  Instead, the DAOS layer can handle all of the IO
-operations from the user API layer, albiet with the potential performance
-penalty of manipulating the shared, persistent storage array. At the bottom is
-the Versioning Object Storage Device (VOSD) (in purple).  It serves as the
-interface for storing objects of all types efficiently for each storage device
-in the parallel storage array. Think of this layer as the physical disk
-interface layer. In terms of Lustre, this would replace the API on individual
-storage devices with an interface friendlier to the containers of objects and
-transactions/epochs concepts used in the higher layers.
+storage interface for the IO stack (in green) and offers features like Burst
+Buffers to insulate the persistent storage array from bursty IO workloads.
+Ideally, the IOD layer's functionality can be optional based on available
+hardware and compute power provided on the IO Nodes (IONs). Much of the
+functionality offered at this layer would shift either up or down the stack as
+discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
+layer serves as the persistent storage interface and translation layer between
+the user-visible object model and the requirements of the underlying storage
+infrastructure. It is intended to be the traditional file system-like
+foundation on which everything else is built with no dependence on any
+technologies specified above it (in dark pink and yellow). For example, the IOD
+layer with or without burst buffers is not required for DAOS to operate
+properly.  Instead, the DAOS layer can handle all of the IO operations from the
+user API layer, albiet with the potential performance penalty of manipulating
+the shared, persistent storage array. At the bottom is the Versioning Object
+Storage Device (VOSD) (in purple).  It serves as the interface for storing
+objects of all types efficiently for each storage device in the parallel
+storage array. Think of this layer as the physical disk interface layer. In
+terms of Lustre, this would replace the API on individual storage devices with
+an interface friendlier to the containers of objects and transactions/epochs
+concepts used in the higher layers.
 
@@
 The major previous work is really limited to full file systems rather than the
-mountain of file system refinements made over the years.
+mountain of file system refinements made over the years. A selection of these
+other file systems and some features that make it relatively unique are
+described below.
 
@@
 PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
-such as a single process opening a file and sharing the handle.
+such as a single process opening a file and sharing the handle. It has been
+commercialized in recent years as OrangeFS.
 
@@
 clusters offering scalable performance and fine-grained end-user and
-programmatic control over how data is placed in the storage system. Scalability
-limitations are part of the motivation for the FFSIO project.
+programmatic control over how data is placed in the storage system.
 
@@
 
-The idea of sharding is to store a portion of a data set on a single storage
-target. This is similiar in concept to the HDF-5 hyperslab. The FFSIO stack
-supports sharding the data in the default or some other structured way as well
-as ``re-sharding'' based on application needs. For example, reordering the
-data so that a different dimension is the ``fast'' dimension may greatly
-improve the performance of a subsequent data analytics task. A common scenario
-where this is common is a Fortran code (column-major) writes data for a C code
-(row-major) to analyze.
+The idea of sharding is to store a logically complete portion of a data set on
+a single storage target. This is similiar in concept to the HDF-5 hyperslab.
+The FFSIO stack supports sharding the data in the default or some other
+structured way as well as ``re-sharding'' based on application needs. For
+example, reordering the data so that a different dimension is the ``fast''
+dimension may greatly improve the performance of a subsequent data analytics
+task. A common scenario where this is common is a Fortran code (column-major)
+writes data for a C code (row-major) to analyze.
+
+Additionally, sharding at the IOD layer is done in a couple of different ways.
+When writing from the compute nodes, the default layout will reflect what the
+processes write. This can be reorganized using function shipping. At the DAOS
+layer, this data organization may change based on the type of storage devices
+beneath it.
 
@@
 
-\section{VOSD Layer}\label{sec:vosd}
+\section{VOSD Layer}
+\label{sec:vosd}
+
 The Versioning Object Storage Device (OSD) layer operates as the layer for each
@@
 \label{sec:broader}
+
+Several concepts crosscut many of these layers and are best described in a
+single location. For example, transactions and epochs are visible from the
+user API level down into the DAOS layer. While each layer affects the concept,
+it is best to look at it across all of the layers.
+
+In the subsections that follow, we examine transactions and epochs, metadata
+management, function and analysis shipping, and the aribtrary connected graphs
+support.
 
@@
 
-This will exemplify how applications make use of the Exascale stack. For FF,
-support for GraphLab and GraphBuilder will be prototyped, which are
-graph-processing frameworks. GraphBuilder is a set of MapReduce tasks that
-extract, normalize, partition and serialize a graph out of unstructured data,
-and writes graph-specific formats into HDFS. These files are later consumed by
-GraphLab, a vertex-centric, asynchronous execution engine that runs directly on
-top of HDFS (i.e., non-MapReduce). The following illustrates the architecture
-of both frameworks:
+What people popularly consider Big Data applications fall into two broad
+categories. First, data processing tasks that fit into the MapReduce model
+where data is tagged and sorted to discover relationships. These sorts of
+applications only require scale out rather than scale up. Scaling out requires
+replicas to process data simultaneously, but do not need to coordinate for that
+data processing. Scaling up, what scientific simulations do, requires sometimes
+serious coordination between the data processing processes for any of them to
+succeed. The second category is more similar to scientific simulations in that
+it requires information from other processes to succedd. These graph-style
+applications can be done using Hadoop, but it does not fit as well.
+
+To address these graph applications, GraphLab and GraphBuilder are evaluated
+to determine how to extend the IO stack to better support these applications.
+GraphBuilder is a set of MapReduce tasks that extract, normalize, partition and
+serialize a graph out of unstructured data, and writes graph-specific formats
+into HDFS. These files are later consumed by GraphLab, a vertex-centric,
+asynchronous execution engine that runs directly on top of HDFS (i.e.,
+non-MapReduce). The following illustrates the architecture of both frameworks:
 
@@
 modified. After these modifications are implemented, GraphBuilder will
-be able to write the partitioned graph in (the newly proposed) HDF5
+be able to write the partitioned graph in the newly proposed HDF5
 files which will thus be stored in the IOD nodes (or IONs) in a
@@
 the library to perform at high speeds by benefiting from the new
-features (see next section). In general both frameworks will be modified
-so that calls to HDFS-based formats are replaced by the proposed HDF5
+features, such as the function shipping. In general both frameworks will be
+modified so that calls to HDFS-based formats are replaced by the proposed HDF5
 ones. This is referred to as the HDF Adaptation Layer or HAL and will
@@
 
+The initial phase of this project has determined the necessary changes in the
+HDF-5 format to support these features. These identified features will be
+proven during phase 2 with a demonstration of GraphBuilder and GraphLab.
+
 \section{Demonstration}
+\label{sec:evaluation}
+
 This stack has a prototype implementation intended to test concepts rather
@@
 The Fast Forward Storage and IO Stack project has designed a good first pass at
-addressing the requirements for an extreme scale data storage mechanism. The
-split between the IOD layer and the DAOS layer offers a fast place for
-intermediate data without requiring the overhead of writing to persistent
-storage. The envisioned transaction mechanism, while not perfect in the current
-form, is another good attempt to address both failures and prevent access to
-incomplete or incorrect data by downstream data consumers. Integrated with the
-IOD functionality, this concept represents the consensus approach for what will
-be required.
-
-The partial metadata management incorporated into the IOD layer and the lack of
-consideration for how to handle and recover from failures are oversights in the
-current documents. It is our understanding that these will be addressed in the
-next phase and we hope to help inform that effort with our experiences.
-
-We hope that the efforts made in the \DDTns, Lightweight File Systems, and
-other efforts to explore the requirements for this space, along with the
-analysis presented in this paper will prove useful for the next phase of the
-Fast Forward project.
+addressing the requirements for an extreme scale data storage mechanism.  By
+prefering a high level user API like HDF5 rather than using the POSIX
+interface, more advanced functionality can be incorporated with less end-user
+impact. The introduction of the IOD layer with buffering will absorb the
+difference between the compute node IO demands and the available bandwidth in
+the storage array. With DAOS supporting translating the container and object
+model to the underlying storage options, different storage technologies can be
+deployed over time.
+
+With the overall stack design complete with a prototype implementation,
+refinements, such as fault detection and recovery, can be designed and tested.
+These and other activities for phase 2 will ultimately generate what is likely
+to be the next generation storage stack for extreme scale platforms.
 

### 2014-05-12 paper.tex
near final edit and incorporated the performance graphs

@@
 %\usepackage{subfig}
-%\usepackage{subfigure}
+\usepackage{subfigure}
 \usepackage[pdftex]{}
@@
 
-\addtolength{\parskip}{-0.06in}
+\addtolength{\parskip}{-0.08in}
 
@@
 excellent opportunity to explore the complete design and how it will address
-the needs of extreme scale platforms.  This paper not only provides a timely
-summary of important aspects of the design specifications but also capture the
-underlying reasoning that is not available elsewhere. With this information, we
-encourate the broader community to understand the design, intent, and future
-directions to foster discussion guiding phase 2 and the ultimate production
-storage stack based on this work.
-
-While many of the design elements are best practices from current research,
-there are many new contributions as well as different integration approaches
-offering different performance and quality of service guarantees. The specifics
-of how the different layers interact offer a new approach for managing data for
-extreme scale platforms and workloads.
-
-For the current demonstration system, the HDF5 user API gives a familiar
-way to interact with all of the new functionality while largely insulating 
-end-users from the intracisies of dealing with the advanced functionality
-offered by the stack.  The various layers below the user API include optional
-components that may aid in scalability and simplify implementation across a
-wide range of platforms.  The data storage layer offers support both
-for local machine and data center wide storage arrays as part of the design.
-
-In addition to these features supporting HPC platforms, the stack also
-addresses the need of big data applications as demonstrated through the
-Arbitrary Connected Graphs tool.
+the needs of extreme scale platforms.  This paper examimes the overall design.
+Each layer of the proposed stack is discussed in some detail along with
+cross-cutting topics that impact multiple layers, such as transactions and
+metadata management. We also describe the proposed features to better support
+some classes of big data applications.
+
+This paper not only provides a timely summary of important aspects of the
+design specifications but also captures the underlying reasoning that is not
+available elsewhere. With this information, we encourate the broader community
+to understand the design, intent, and future directions to foster discussion
+guiding phase 2 and the ultimate production storage stack based on this work.
+An initial performance evaluation of the early prototype implementation is also
+provided to validate the presented design.
 
@@
 Current production HPC IO stack design is unlikely to offer sufficient features
-and performance to adequately serve the needs of an extreme scale platform.
-Adding to the complexity of the problem is the variety of Big Data problems
-that are taking advantage of these platforms. Trying to address both of these
-domains is a daunting problem.
+and performance to adequately serve extreme scale science platform
+requirements.  Adding to the problem complexity is the variety of Big Data
+problems users want to address using these platforms. Trying to address both of
+these domains is a daunting problem.
 
@@
 Energy's Office of Advanced Simulation and Computing and Advanced Scientific
-Computing Research commissioned an effort to develop a design and prototype for
-an IO stack suitable for the extreme scale environment called the Fast Forward
-Storage and IO (FFSIO) project. This is a joint effort led by Lawrence
-Livermore National Laboratory, with the DOE Data Management Nexus leads Rob
-Ross and Gary Grider as coordinators and contract lead Mark Gary. The
+Computing Research commissioned a project to develop a design and prototype for
+an IO stack suitable for the extreme scale environment. It will be referred to
+as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
+by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
+leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
 participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
@@
 be the perfect target architecture, the reality of budgets has tempered many of
-the decisions. For exmaple, extensive availability of NVRAN or SSDs on all of
+the decisions. For example, extensive availability of NVRAN or SSDs on all of
 the compute nodes is currently not economically feasible limiting some of the
@@
 multi-version concurrency control. Through the use of a copy-on-write style
-mechanism reducing space and time requirements, multiple versions of an object
-can be stored efficiently. By assuming the client interface will be through an
-IO library, a more complicated interface that offers richer functionality can
-be incorporated while requiring only minimal end-user code changes.  Managing
-most data access in a platform-local layer rather than requiring writing to
-centralized storage will better support the performance and energy requirements
-of extreme scale integrated application workflows.
+mechanism reducing storage space and writing time requirements, multiple
+versions of an object can be stored efficiently. By assuming the client
+interface will be through an IO library, a more complicated interface offering
+richer functionality can be incorporated while requiring only minimal end-user
+code changes.  Managing most data access in a platform-local layer rather than
+requiring writing to centralized storage will better support the performance
+and energy requirements of extreme scale application compositions.
 
@@
 \begin{figure}[htbp]
+\vspace{-0.10in}
 \centering
 \includegraphics[width=\columnwidth]{images/arch-mapping}
+\vspace{-0.15in}
 \caption{Target Architecture and Component Mapping}
 \label{fig:arch-mapping}
+\vspace{-0.15in}
 \end{figure}
@@
 The top layer is generally a high level IO library, such as the demonstration
-HDF-5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
+HDF5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
 access to the storage stack through such an API to manage the complexity of
@@
 
-Along with the analysis of the published design documents, a discussion of the
+Along with analysis of the published design documents, a discussion of the
 design philosophy representing the overall intent is presented. This
@@
 discusses the programmatic interface end users will see when interacting with
-the storage array. This will be discussed in the contect of the HDF-5 based
+the storage array. This will be discussed in the contect of the HDF5 based
 example library used for the functionality demonstration. Section~\ref{sec:iof}
@@
 potentially optional and mention the cross-cutting features discussed in a
-later, cross-cutting section. Section~\ref{sec:daos} discusses how the
-persistent storage array interface itself functions. As with the IOD layer, the
-disucssion of cross-cutting features will be mentioned, but disucssed more
-fully in the cross-cutting section. The VOSD layer is disucssed in
-Section~\ref{sec:vosd}. In particular, the mapping between the DAOS and VOSD
-layers are explored as it pertains to the physical storage. Next is an
-exploration of cross-cutting features like transactions and metadata management
-in Section~\ref{sec:broader}. Since these and other features are spread across
-multiple layers, it makes more sense to discuss them independently once an
-understanding of the overall structure has been presented.  A demonstration
-of the functionality is presented in Section~\ref{sec:evaluation}. This shows
-that the prototype functionality validation prototype can function. The paper
-is concluded in Section~\ref{sec:conclusion} with a summary of the broad issues
-covered in the paper.
+later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
+layer functions. As with the IOD layer, the cross-cutting features will be
+mentioned, but disucssed more fully in the cross-cutting section. The VOSD
+layer is disucssed in Section~\ref{sec:vosd}. In particular, the mapping
+between the DAOS and VOSD layers are explored as it pertains to the physical
+storage. Next is an exploration of cross-cutting features like transactions and
+metadata management in Section~\ref{sec:broader}. Since these and other
+features are spread across multiple layers, it makes more sense to discuss them
+independently once an understanding of the overall structure has been
+presented.  A demonstration of the functionality is presented in
+Section~\ref{sec:evaluation}. This shows that the prototype system based on the
+proposed design can function. The paper is concluded in
+Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
+paper.
 
@@
 Many projects over the last couple of decades have sought to address some
-challenging aspect of parallel file system design. The recent rise of the ``Big
+challenging aspect of parallel file system design. The recent rise of ``Big
 Data'' applications with different characteristic IO patterns have somewhat
@@
 interface for programmatically interacting with the FFSIO stack, the team used
-the HDF-5 API and leveraged the Virtual Object Layer (VOL) for the initial
+the HDF5 API and leveraged the Virtual Object Layer (VOL) for the initial
 design and implementation demonstration. This also serves as a good test
@@
 %  Virtual Object Layer (VOL) offers the ability to plug in new implementations
-%  for HDF-5 semantics. In this case, the HDF-5 calls are implemented using the
+%  for HDF5 semantics. In this case, the HDF5 calls are implemented using the
 %  IOD or DAOS API rather than MPI-IO or POSIX.
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=\columnwidth]{images/hdf5.png}
-\caption{Sample file illustrating the hierarchical structure of HDF5 files}
-\label{fig:hdf5-sample}
-\end{figure}
+%\begin{figure}[htbp]
+%\vspace{-0.10in}
+%\centering
+%\includegraphics[width=\columnwidth]{images/hdf5.png}
+%\vspace{-0.20in}
+%\caption{Sample file illustrating the hierarchical structure of HDF5 files}
+%\label{fig:hdf5-sample}
+%\vspace{-0.25in}
+%\end{figure}
 
@@
 \item
-  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
-  is represented by a file. It is a collection of objects that also
-  explains the relationship between them. Every file begins with a root
-  group ``/'', which serves as the ``starting-point'' in the object
-  hierarchy.
+  \textbf{File}: In the HDF5 data model, the collection of data items stored
+  together is represented by a file. It is an object collection that also
+  describes the relationship between them. Every file begins with a root
+  group ``/'' serving as the ``starting-point'' in the object hierarchy.
 
 \item
-  \textbf{Group}: A group is an explicit association between HDF5
+  \textbf{Group}: A group is an object allowing association between HDF5
   objects. It is synonymous with directories in a file system. A group
@@
 \item
-  \textbf{Dataset}: HDF5 datasets are objects that represent actual data
+  \textbf{Dataset}: HDF5 datasets are objects representing actual data
   or content. Datasets are arrays with potentially multiple dimensions. A
   dataset is characterized by a dataspace and a datatype. The dataspace
-  captures the rank (number of dimensions), and the current and maximum
+  captures the rank (number of dimensions) and the current and maximum
   extent in each dimension. The datatype describes the type of its data
@@
 \item
-  \textbf{Attribute}: Attributes are used for annotating datasets,
-  groups, and datatype objects. They are datasets themselves, and are
-  attached to existing objects they annotate.
+  \textbf{Attribute}: Attributes are used for annotating HDF5 objects. They are
+  datasets themselves and are attached to existing objects.
 
@@
 
-For example, as shown in Figure \ref{fig:hdf5-sample}, the file Sample.h5
-contains the root group which itself contains a group G1 and two datasets, D1
-and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1.
-The objects and the relationships between them can be represented as a B-tree,
-which is used internally by HDF5 to index its objects.
+%For example, as shown in Figure \ref{fig:hdf5-sample}, the file Sample.h5
+%contains the root group which itself contains a group G1 and two datasets, D1
+%and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1.
+%The objects and the relationships between them can be represented as a B-tree,
+%which is used internally by HDF5 to index its objects.
 
@@
 just below the public API. The VOL exports an interface that allows writing
-plugins for HDF5, thereby enabling developers to handle the data in ways other
-than necessarily writing to storage in an HDF5 format.  Plugin writers provide
-an implementation for a set of functions that are trusted to provide the
-proper semantics represented by the function for the new environment. For
-example, data staging could be implemented in the VOL layer by replacing
-writing to disk in the HDF5 format to sending data to a data staging area using
-some messaging mechanism.
-
-For this project, rather than the default writing to disk in the HDF-5 format,
-the VOL layer is used to interact with the IOD layer and the different concepts
-it offers without requiring all of the functionality be exposed to users. For
+plugins for HDF5 enabling developers to handle data in ways other than writing
+to storage in an HDF5 format.  Plugin writers provide an implementation for a
+set of functions and are trusted to provide the proper semantics for the new
+environment. For example, data staging could be implemented in the VOL layer by
+replacing writing to disk in the HDF5 format to sending data to a data staging
+area using some messaging mechanism.
+
+For this project, rather than the default writing to disk in the HDF5 format,
+the VOL is used to interact with the IOD layer and the different concepts it
+offers without requiring all of the functionality be exposed to users. For
 example, the containers and objects concept is transparently mapped to the
 files and datasets existing HDF5 users are familiar with.  This reduces the
-difficulty in porting applications to use the new IO stack.
+difficulty porting applications to the new IO stack.
 
 \begin{figure}[htbp]
+\vspace{-0.10in}
 \centering
 \includegraphics[width=\columnwidth]{images/vol-arch.png}
+\vspace{-0.10in}
 \caption{Architectural view of the VOL abstraction mechanism}
 \label{fig:vol-arch}
+\vspace{-0.10in}
 \end{figure}
@@
 IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
-arrive at the IO nodes, they are translated into I/O Dispatcher (IOD) API calls
+arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
 and executed at the IONs.
@@
 \begin{table}[ht]
-    %\vspace{-0.15in}
+    \vspace{-0.10in}
     \centering
@@
     \bigskip
-    %\vspace{-0.15in}
+    \vspace{-0.15in}
 
@@
 the massive process count on the storage stack. A current trend mixing MPI with
-a threading library like OpenMP is addressing the same issue. In this case,
-the storage stack has far fewer end-points in which to receive requests and
-data from compute processes. By reducing the number of simultaneous requests,
-delays can be reduced. This has been demonstrated for the file open operation
-with Lustre~\cite{lofstead:2009:adaptable} and to some degree for accessing the
-storage devices themselves~\cite{lofstead:2010:io-variability}. The BlueGene
-platform incorporated dedicated hardware to perform this role. The proposed
-functionality for this layer, beyond managing the number of connections to
-the IOD layer, is to implement function shipping from the compute nodes to the
-IO nodes.
+a threading library like OpenMP is addressing the same issue, but limited to
+handling the parallelism on a single node rather than multiple nodes. Projected
+extreme scale platforms will have far fewer storage stack end-points per
+compute process in which to receive requests and data. By reducing the number
+of simultaneous requests, delays can be reduced. This has been demonstrated for
+the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
+degree for accessing the storage devices
+themselves~\cite{lofstead:2010:io-variability}.  The BlueGene platform
+incorporated dedicated hardware to perform this role. The proposed
+functionality for this layer, beyond managing the number of connections to the
+IOD layer, is to implement function shipping from the compute nodes to the IO
+nodes.
 
@@
 Strictly speaking, the IO Dispatcher layer and included functionality, such as
-burst buffers, is an optional component of the proposed IO stack. All of the
-functionality, such as function and analysis shipping, transaction management,
-and managing asynchronous data movement can be handled by other portions of the
-stack. For an extreme scale platform, the IOD layer will be an essential
-pressure relief valve for the underlying persistent storage layer. By making it
-optional, the proposed stack can be deployed more easily on smaller clusters or
-for those with more constrainted budgets. For simplicity, the rest of this
-section will describe a full stack including all of the proposed IOD
-components.
-
-The core idea for IOD is to provide a way to manage the IO load that is
+burst buffers, is optional. All of the functionality, such as function and
+analysis shipping, transaction management, and managing asynchronous data
+movement can be handled by other portions of the stack. For an extreme scale
+platform, the IOD layer will be an essential pressure relief valve for the
+underlying persistent storage layer. By making it optional, the proposed stack
+can be deployed more easily on smaller clusters or for those with more
+constrainted budgets. For simplicity, the rest of this section will describe a
+full stack including all of the proposed IOD components.
+
+The core idea for IOD is to provide a way to manage IO load that is
 separate from the compute nodes and the storage array. Communication intensive
 activities, such as data rearrangement, can be moved to the IOD layer
-offloading the communication load from the compute nodes. IOD has three main
-purposes. First, the burst buffers works as a fast cache absorbing write
-operations for the slower trickle out to the central storage array. It can also
+reducing the number of participants and message count. IOD has three main
+purposes. First, the burst buffers work as a fast cache absorbing write
+operations that then trickles out to the central storage array. It can also
 be used to retrieve objects from the central storage array for more efficient
@@
 Second, it offers the transaction mechanism for controlling data set visibility
-and to manage faults that would prevent a data set from being used. These
-transactions are local to the IOD layer until persisted to the DAOS layer
-eliminating the need for burdening the persistent storage with transient data.
-Third, data processing operations can be placed in the IOD. These operations
-are intended to offer data rearrangement, filtering, and similar operations
-prior to data reaching the central storage array.
+and to manage faults that could expose an incomplete or corrupt data set to
+users. These transactions are local to the IOD layer until persisted to the
+DAOS layer eliminating the need for burdening the persistent storage with
+transient data.  Third, data processing operations can be placed in the IOD.
+These operations are intended to offer functionality like data rearrangement
+and filtering prior to data reaching the central storage array.
 
@@
 
-%\subsection{I/O Dispatcher API}
+%\subsection{IO Dispatcher API}
 %\label{sec:fast-forwards-io-dispatcher-api}
 %
-%The Fast Forward I/O and Storage project is aimed at integrating many of
+%The Fast Forward IO and Storage project is aimed at integrating many of
 %the services existing currently in middleware libraries in a
@@
 %
-%From a high-level point of view, the Fast Forward's I/O Dispatcher (IOD)
+%From a high-level point of view, the Fast Forward's IO Dispatcher (IOD)
 %interface allows the user to:
@@
 multi-dimensional array object can be split across multiple IO nodes. Each
-piece of this array is called a {\em shard}. The IOD API supports the following
-sharding strategies:
+piece of this array is called a {\em shard}.  The idea of sharding is to store
+a logically complete portion of a data set on a single storage target. This is
+similiar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
+the data in the default or some other structured way as well as ``re-sharding''
+based on application needs. For example, reordering the data so that a
+different dimension is the ``fast'' dimension may greatly improve the
+performance of a subsequent data analytics task. A common scenario where this
+is common is a Fortran code (column-major) writes data for a C code (row-major)
+to analyze.  The IOD API supports the following sharding strategies:
 
@@
   \textbf{contiguous}. fixed chunking, distributed in a round-robin
-  fashion accross the I/O nodes.
+  fashion accross the IO nodes.
 \item
@@
 \item
-  \textbf{user-defined}. either contiguous or chunked but user specifies
+  \textbf{user-defined}. either contiguous or chunked, but user specifies
   where to place each individual shard.
@@
 multiple formats if desired. Also, the user can pre-fetch objects from
-the storage cluster into the I/O nodes or read them directly from the
+the storage cluster into the IO nodes or read them directly from the
 storage cluster. At the semantic level (HDF5), indices can be created
@@
 
-\subsubsection{Sharding}
-\label{sec:sharding}
-
-The idea of sharding is to store a logically complete portion of a data set on
-a single storage target. This is similiar in concept to the HDF-5 hyperslab.
-The FFSIO stack supports sharding the data in the default or some other
-structured way as well as ``re-sharding'' based on application needs. For
-example, reordering the data so that a different dimension is the ``fast''
-dimension may greatly improve the performance of a subsequent data analytics
-task. A common scenario where this is common is a Fortran code (column-major)
-writes data for a C code (row-major) to analyze.
-
-Additionally, sharding at the IOD layer is done in a couple of different ways.
-When writing from the compute nodes, the default layout will reflect what the
-processes write. This can be reorganized using function shipping. At the DAOS
-layer, this data organization may change based on the type of storage devices
-beneath it.
-
-\subsection{I/O Nodes}
+%At the DAOS layer, the sharding can be changed based on the type of storage
+%devices beneath it.
+
+\subsection{IO Nodes}
 IOD processes are hosted on the IO nodes that interface a general compute area
-with the storage array. The IO noes handle requests forwarded by the scientific
-applications, potentially integrate a tier of solid-state devices to absorb the
-burst of random operations, and organize/re-format the data so that transfers
-to/from the staging area from/to the traditional parallel file system can be
-done more efficiently. It also has the capacity to executing analysis on the
-fresh data that has just been generated by simulation applications running at
-the compute nodes. As the data arrives, re-organization and data preparation
-can be applied in order to anticipate the execution of analytical tasks.
+with the storage array. The IO nodes handle requests forwarded by the
+scientific applications, potentially integrate a tier of solid-state devices to
+absorb the burst of random or high volume operations, and organize/re-format
+the data so that transfers to/from the staging area from/to the traditional
+parallel file system can be done more efficiently. It also has the capacity to
+execute analysis on data recently generated by simulation applications running
+at the compute nodes, but not persisted to the storage array. As the data
+arrives, re-organization and data preparation can be applied in order to
+anticipate the execution of analytical tasks.
 
@@
 \centering
+\vspace{-0.10in}
 \includegraphics[width=\columnwidth]{images/exa-arch.png}
+\vspace{-0.20in}
 \caption{Exascale Architecture}
 \label{fig:exa-arch}
+\vspace{-0.10in}
 \end{figure}
 
-A common cluster configuration for this type of deployment is shown in
+A common configuration for this type of deployment is shown in
 Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
-compute nodes (CNs) through the same fast fabric (e.g., InfiniBand), while the
-connection to the external storage cluster is through a secondary slower
-channel (e.g., 10Gb Ethernet). By providing additional storage, such as SSDs on
-the IO nodes, these nodes are capable of better regulating the IO pressure on
+compute nodes (CNs) through the same fast fabric (e.g., InfiniBand) while the
+connection to the external storage cluster is through a secondary, slower
+channel (e.g., 10Gb Ethernet). By providing additional storage on the IO nodes,
+such as SSDs, these nodes are capable of better regulating the IO pressure on
 the underlying storage array better than simple forwaring gateways. For this
@@
 %a POSIX interface to storage, existing middleware either transparently handle
-%the I/O operations, appearing as regular POSIX calls to applications; or modify
-%the I/O API as little as possible in order to minimize the impact on production
+%the IO operations, appearing as regular POSIX calls to applications; or modify
+%the IO API as little as possible in order to minimize the impact on production
 %codes.
@@
 %
-%The Fast Forward Storage I/O project is aimed at merging the features of
+%The Fast Forward Storage IO project is aimed at merging the features of
 %existing middleware into a next-generation parallel file system stack. This new
-%storage API termed the I/O dispatcher (IOD) is object-based, transactional,
-%asynchronous and I/O-staging-aware (Figure~\ref{fig:exa-stack} (right)).
+%storage API termed the IO dispatcher (IOD) is object-based, transactional,
+%asynchronous and IO-staging-aware (Figure~\ref{fig:exa-stack} (right)).
 
@@
 %\item
-%  I/O Dispatcher (IOD).
+%  IO Dispatcher (IOD).
 %\item
@@
 
-As mentioned before, the stack provides \emph{non-POSIX, object-based,
-transactional and asynchronous active-storage}, meaning that POSIX is
-supplanted by new object interfaces that reach up to the HDF5 layer
-(Application I/O in the stack diagram); transactional semantics are present,
-from HDF down to the VOSD layer (Storage layer); clients don't need to wait for
-any blocking operation; and analysis can be shipped and executed on I/O or DAOS
-nodes.
+%As mentioned before, the stack provides \emph{non-POSIX, object-based,
+%transactional and asynchronous active-storage}, meaning that POSIX is
+%supplanted by new object interfaces that reach up to the HDF5 layer
+%(Application IO in the stack diagram); transactional semantics are present,
+%from HDF down to the VOSD layer (Storage layer); clients don't need to wait for
+%any blocking operation; and analysis can be shipped and executed on IO or DAOS
+%nodes.
 
@@
 Future work will generalize this model to potentially support dynamic IO node
-allocation and potentially examine the possibility of oversubscription. It will
-be strictly necessary to consider shared IO nodes for cases where the number
-of deployed applications exceeds the number of IO nodes. This is also under
-consideration for future work. This first pass is focused on extreme scale
-application runs that use the vast majority of a platform rather than a
-capacity cluster where end-to-end performance is a lesser concern.
+allocation and examine the possibility of oversubscription. It will be strictly
+necessary to consider shared IO nodes for cases where the number of deployed
+applications exceeds the number of IO nodes.  This first phase focuses on
+extreme scale application runs that use the vast majority of a platform rather
+than a capacity cluster where end-to-end performance is a lesser concern.
 
@@
 The burst buffers design, as presented in the IOD documents, limits the
-placement of the function operators and SSD buffers to the IO nodes. The team
-does acknowledge the limitations of this design and intend to ultimately focus
-on spreading the IOD layer from the IO nodes into the compute area as well.
-This is intended both to help address the limitations of the IO bandwidth and
-compute capability of these few nodes for data processing, but also to take
-advantage of new layers in the storage hierarchy. By incorporating NVRAM into
-compute nodes, new options for buffering data prior to being moved to
-centralized storage become available and addresses potential concerns about
-SSD performance. For example, including a small amount of Phase Change memory
-into many or most compute nodes offers a way to move data outside of both the
+placement of the function operators and SSD buffers to the IO nodes. The
+limitations of this design are acknowledged and the intent is to ultimately
+spread the IOD layer from the IO nodes into the compute area as well.  This is
+intended both to help address the limitations of the IO bandwidth and compute
+capability of these few nodes for data processing, but also to take advantage
+of new layers in the storage hierarchy. By incorporating NVRAM into compute
+nodes, new options for buffering data prior to being moved to centralized
+storage become available and addresses potential concerns about SSD
+performance. For example, including a small amount of Phase Change memory into
+many or most compute nodes offers a way to move data outside of both the
 compute and IO path for data and communication intensive operations. Other
-projects~\cite{zheng:2010:predata} have suggested this will have value, but the
+projects~\cite{zheng:2010:predata} have shown this will have value, but the
 cost will have to be considered as part of the overall platform budget. This
@@
 
-Burst buffers being optional is a high level goal, but not one considered at a
-detailed level within the design. For example, if there is no burst buffer, all
-of the advanced functionality proposed for the IOD layer would have to work
-against the DAOS layer instead. For example, function shipping assumes it will
-operate on fast, local data within the IOD layer rather than against the
-globally shared DAOS layer. With the additional desire to support using compute
-node resources for these operations, serious work will be required to make a
-fully functional end-to-end IOD layer implementation for a production system.
+Burst buffers being optional is a high level goal, but not considered in detail
+within the phase one design. If there is no burst buffer, all of the advanced
+functionality proposed for the IOD layer would have to work against the DAOS
+layer instead. For example, function shipping assumes it will operate on fast,
+local data within the IOD layer rather than against the globally shared DAOS
+layer. With the additional desire to support using compute node resources for
+these operations, serious work will be required to make a fully functional
+end-to-end IOD layer implementation for a production system.
 
@@
 The Distributed Asynchronous Object Storage layer serves as the traditional
-parallel file-system interface layer for the storage devices. This is the
+parallel file system interface layer for the storage devices. This is the
 consistent, global view of the underlying devices represented in this stack
@@
 represents the portion of a container that is spread across potentially
-multiple physical storage devices. The physical storage devices are represented
-by the VOSD layer described in Section~\ref{sec:vosd}.
+multiple physical storage devices.  The physical storage devices are
+represented by the VOSD layer described in Section~\ref{sec:vosd}.
 
@@
 
-%Epochs are represented through the shard mechanism. Each container shard is
-%versioned according to the transactions in the IOD layer that are persisted to
-%the DAOS layer as an epoch. The container shard is a piece of the container
-%with the contained objects concept. Since the container is largely logical,
-%this generally maps to a shard of an object within a container.
-
 The current implementation has the DAOS layer map the container/object data
@@
 
+%Epochs are represented through the shard mechanism. Each container shard is
+%versioned according to the transactions in the IOD layer that are persisted to
+%the DAOS layer as an epoch. The container shard is a piece of the container
+%with the contained objects concept. Since the container is largely logical,
+%this generally maps to a shard of an object within a container.
+
 Overall, the DAOS layer serves as the shared persistent storage interface for
@@
 
-The Versioning Object Storage Device (OSD) layer operates as the layer for each
-persistent storage device used to support the parallel storage array. In the
-purest form, it uses a local file system to arrange storage of objects that
+The Versioning Object Storage Device (OSD) layer operates as the interface for
+each persistent storage device used to support the parallel storage array. In
+the purest form, it uses a local file system to arrange storage of objects that
 represent parts of the higher level objects in containers.
@@
 single location. For example, transactions and epochs are visible from the
-user API level down into the DAOS layer. While each layer affects the concept,
+user API level down into the VOSD layer. While each layer affects the concept,
 it is best to look at it across all of the layers.
@@
 
-As mentioned above, the transaction mechanism manifests in two forms. At the
-IOD layer, they are called transactions and are used to judge whether or not a
-set of distributed, asynchronous modifications across a set of related objects
-is complete or not.  It is also used to control access by treating the
-transaction ID of committed transaction as a version identifier.  At the DAOS
-layer, they are called epochs and represent persisted (durable) transactions
-from the IOD layer. Each of these offers different functionality, but are
-connected as is explained below.
+As mentioned above, the transaction mechanism manifests in two forms. From the
+user level down through the IOD layer, they are called transactions and are
+used to judge whether or not a set of distributed, asynchronous modifications
+across a set of related objects is complete or not.  It is also used to control
+access by treating the transaction ID of committed transaction as a version
+identifier.  At the DAOS layer and below, they are called epochs and represent
+persisted (durable) transactions from the IOD layer. Each of these offers
+different functionality, but are connected as is explained below.
 %How these differ from the \DDT approach is also explored.  While IOD's and
@@
 
-\subsubsection{IOD Transactions}
+\subsubsection{Transactions}
 To understand how transactions are used in the IOD layer, some terminology and
@@
 multi-dimensional arrays, and blobs.  The easiest way to understand these types
-is to evaluate these from the perspective of an HDF-5 file, the initial user
+is to evaluate these from the perspective of an HDF5 file, the initial user
 interface layer. The key-value store represents a collection of attributes or
@@
 
-\subsubsection{DAOS Epochs}
+\subsubsection{Epochs}
 The Epoch mechanism differs from transactions. Instead of focusing on when a
-particular output is complete, an epoch represents incremental persisted copies
-of a container. To simplify the mapping between an IOD transaction and the DAOS
-epochs, when an IOD transaction is persisted to DAOS, the IOD transaction ID is
-the used as the epoch ID. The key difference is that at the DAOS layer, some
-transaction (epoch) IDs will not be represented since not all IOD transactions
-are necessarily persisted.
+particular output is complete, an epoch represents incremental persisted
+container copies.  To simplify the mapping between an IOD transaction and the
+DAOS epochs, when an IOD transaction is persisted to DAOS, the IOD transaction
+ID is the used as the epoch ID. The key difference is that at the DAOS layer,
+some transaction (epoch) IDs will not be represented with data since not all
+IOD transactions are necessarily persisted. Maintaining this ID continuity is
+critial for multiplatform use. Since the shared point is the DAOS layer, any
+user adding a new version to a file must be able to determine the most recent
+transaction ID no matter from where the container was updated last.
 
@@
 While the metadata design is not fully defined, there are a few things that
-are intended. For example, there is a standard, well-known container that is
-the system metadata. This includes the list of all other containers. This
+are intended. For example, there will be a standard, well-known container that
+is the system metadata. This includes the list of all other containers. This
 container is treated like any other data in the system and striped as
@@
 difficult. Serious thought on how to do this effectively outside the data path
-should be considered for phase two.
+will be considered for phase two.
 
@@
 
-In terms of the CN-ION communication model, a client/server architecture
-is implemented: every ION runs an IOFSL (I/O Function Shipping
-Layer) server; the IOFSL client is integrated into the HDF5
-library which runs on each CN. A client can forward requests to any
-number of IONs. Every I/O operation issued by HDF5 is asynchronously
-shipped to the IOFSL server and asynchronously executed.
+A client/server architecture is implemented for the Compute Node-IO Node
+communication model.  Every ION runs an IOFSL (IO Function Shipping Layer)
+server. The IOFSL client is integrated into the HDF5 library which runs on each
+CN. A client can forward requests to any number of IONs. Every IO operation
+issued by HDF5 is asynchronously shipped to the IOFSL server and asynchronously
+executed. As it is currently implemented, the only functionality that can be
+``shipped'' already exists on the IONs and is activated using RPC calls. This
+will be re-evaluated for phase two to provide more dynamic functionality.
 
@@
 succeed. The second category is more similar to scientific simulations in that
-it requires information from other processes to succedd. These graph-style
+it requires information from other processes to succeed. These graph-style
 applications can be done using Hadoop, but it does not fit as well.
@@
 \centering
+\vspace{-0.10in}
 \includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
+\vspace{-0.30in}
 \caption{GraphLab and GraphBuilder stacks}
 \label{fig:graphlab-graphbuilder}
+\vspace{-0.10in}
 \end{figure}
@@
 The initial phase of this project has determined the necessary changes in the
-HDF-5 format to support these features. These identified features will be
+HDF5 format to support these features. These identified features will be
 proven during phase 2 with a demonstration of GraphBuilder and GraphLab.
@@
 
-This stack has a prototype implementation intended to test concepts rather
-than performance and scalability. It has focused on examining the interaction
-of the different APIs for each layer to flesh out any detailed requirements
-or concerns that may have been missed in the conceptualization of this IO stack.
-
-We run this demonstration using two different modes. A standard HDF-5 test
-application is used with the native file system layer writing directly to
-storage is measured. Then, the same test is run against the requirements
-prototype to demonstrate that it can function for this example as well. The
-performance of both of these tests are reported to give a very rough idea of
-the overhead that might be involved. Rather than a true overhead, this should
-be considered the maximum overhead that should be expected once an optimized,
-fully functional IO stack is deployed without relying on translating to an
-underlying parallel file system. Figure~\ref{fig:evaluation} shows the
-relative performance for both test runs.
-
-The testing evironment for this demonstration is ....
-
-\begin{figure}[htbp]
+This stack has an early prototype implementation intended to test concepts
+rather than performance and scalability. It has focused on examining the
+interaction of the different APIs for each layer to flesh out any detailed
+requirements or concerns that may have been missed in the conceptualization of
+this IO stack. To demonstrate the viability of the IO stack described in this
+paper, we show some very early performance results from the untuned prototype.
+
+All of the tests are performed on the Buffy Cray test cluster at Los Alamos
+National Laboratory. It consists of ......
+
+We run two different sets of tests. The first set in
+Figure~\ref{fig:eval-hosts} show the reading and writing performance for
+different number of hosts. Each read or write is 4 GB against the x-axis number
+of hosts. The second set in Figure~\ref{fig:eval-size} show the performance of
+reading and writing different sizes for 56 clients, the smallest client count
+when performance stabilizes in the number of hosts tests.  The performance of
+both of these tests are reported to give a very rough idea of the overhead that
+might be involved. Rather than a true overhead, this should be considered the
+maximum overhead that should be expected once an optimized, fully functional IO
+stack is deployed without relying on translating to an underlying parallel file
+system.
+
+\begin{figure*}[htbp!]
 \centering
-\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
-\caption{Functionality Demonstration Validation}
-\label{fig:evaluation}
-\end{figure}
+\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+\vspace{-0.05in}
+\end{figure*}
 
@@
 
-With the overall stack design complete with a prototype implementation,
-refinements, such as fault detection and recovery, can be designed and tested.
-These and other activities for phase 2 will ultimately generate what is likely
-to be the next generation storage stack for extreme scale platforms.
+With the overall stack design a prototype implementation complete, refinements,
+such as fault detection and recovery, can be designed and tested.  These and
+other activities for phase 2 will ultimately generate what is likely to be the
+next generation storage stack for extreme scale platforms.
 
 \section{Acknowledgements}
-The authors would like to thank Eric Barton, John Bent, Gary Grider, and
-Quincey Koziol for their early review comments and discussions that clarified
-the details of the design, the intent of the design, and the future plans.
 

### 2014-05-12 paper.tex
added GraphBuilder and GraphLab refs, beefed up the ACG section.

@@
 
-\subsection{ACG}
+\subsection{Arbitrarily Connected Graphs}
 \label{sec:acg}
@@
 What people popularly consider Big Data applications fall into two broad
-categories. First, data processing tasks that fit into the MapReduce model
+categories. First, data processing tasks that can fit into the MapReduce model
 where data is tagged and sorted to discover relationships. These sorts of
@@
 data processing. Scaling up, what scientific simulations do, requires sometimes
-serious coordination between the data processing processes for any of them to
-succeed. The second category is more similar to scientific simulations in that
-it requires information from other processes to succeed. These graph-style
-applications can be done using Hadoop, but it does not fit as well.
-
-To address these graph applications, GraphLab and GraphBuilder are evaluated
-to determine how to extend the IO stack to better support these applications.
-GraphBuilder is a set of MapReduce tasks that extract, normalize, partition and
-serialize a graph out of unstructured data, and writes graph-specific formats
-into HDFS. These files are later consumed by GraphLab, a vertex-centric,
-asynchronous execution engine that runs directly on top of HDFS (i.e.,
-non-MapReduce). The following illustrates the architecture of both frameworks:
+serious coordination between the processes for any of them to succeed. In the
+middle are graph applications that, with some replicated data, can be made to
+fit reasonably well into the MapReduce model. The challenge is having access to
+the edge and vertex lists effectively to build partitions for independent
+processing. GraphBuilder~\cite{Jain:2013:GraphBuilder} is a tool to generate
+effective graph paritions reducing the load for using MapReduce to process
+graph data sets. GraphLab~\cite{Low:2012:GraphLab} offers a way to process
+these graphs efficiently for parallel platforms using a minimum of
+communication. Using these tools as motivators, changes to the HDF5 interface
+and the underlying storage infrastructure is proposed.  The following
+illustrates the architecture of both frameworks:
 
@@
 
-In order to make both work on top of the exascale stack, both have to be
-modified. After these modifications are implemented, GraphBuilder will
-be able to write the partitioned graph in the newly proposed HDF5
-files which will thus be stored in the IOD nodes (or IONs) in a
+In order to make both of these tools work on top of the exascale stack, they
+both have to be modified. After these modifications are implemented,
+GraphBuilder will be able to write the partitioned graph in the newly proposed
+HDF5 files which will thus be stored in the IOD nodes (or IONs) in a
 parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
@@
 modified so that calls to HDFS-based formats are replaced by the proposed HDF5
-ones. This is referred to as the HDF Adaptation Layer or HAL and will
+equivalents. This is referred to as the HDF Adaptation Layer or HAL and will
 provide, from the GraphBuilder/GraphLab point of view:
@@
 HDF5 format to support these features. These identified features will be
-proven during phase 2 with a demonstration of GraphBuilder and GraphLab.
+proven during phase two with a demonstration of GraphBuilder and GraphLab.
 
@@
 All of the tests are performed on the Buffy Cray test cluster at Los Alamos
-National Laboratory. It consists of ......
+National Laboratory. It consists of 84 nodes each with dual, 8 core Intel
+Xeon ES-2670 CPUs at 2.6 GHz and 32 GB RAM. The inteconnect is Cray Gemini.
+The storage array are three sets of SSD-based lustre file systems.
 

### 2014-05-12 paper.tex
removed the large commented out sections
spell check completed. Initial submission version.

@@
 excellent opportunity to explore the complete design and how it will address
-the needs of extreme scale platforms.  This paper examimes the overall design.
+the needs of extreme scale platforms.  This paper examines the overall design.
 Each layer of the proposed stack is discussed in some detail along with
@@
 design specifications but also captures the underlying reasoning that is not
-available elsewhere. With this information, we encourate the broader community
+available elsewhere. With this information, we encourage the broader community
 to understand the design, intent, and future directions to foster discussion
@@
 properly.  Instead, the DAOS layer can handle all of the IO operations from the
-user API layer, albiet with the potential performance penalty of manipulating
+user API layer, albeit with the potential performance penalty of manipulating
 the shared, persistent storage array. At the bottom is the Versioning Object
@@
 discusses the programmatic interface end users will see when interacting with
-the storage array. This will be discussed in the contect of the HDF5 based
+the storage array. This will be discussed in the context of the HDF5 based
 example library used for the functionality demonstration. Section~\ref{sec:iof}
@@
 layer functions. As with the IOD layer, the cross-cutting features will be
-mentioned, but disucssed more fully in the cross-cutting section. The VOSD
-layer is disucssed in Section~\ref{sec:vosd}. In particular, the mapping
+mentioned, but discussed more fully in the cross-cutting section. The VOSD
+layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
 between the DAOS and VOSD layers are explored as it pertains to the physical
@@
 
-Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
+Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
 clusters offering scalable performance and fine-grained end-user and
@@
 
-%\item
-%  Virtual Object Layer (VOL) offers the ability to plug in new implementations
-%  for HDF5 semantics. In this case, the HDF5 calls are implemented using the
-%  IOD or DAOS API rather than MPI-IO or POSIX.
-
 \item\label{item:fs}
@@
 components of HDF5 are described below.
-
-%\begin{figure}[htbp]
-%\vspace{-0.10in}
-%\centering
-%\includegraphics[width=\columnwidth]{images/hdf5.png}
-%\vspace{-0.20in}
-%\caption{Sample file illustrating the hierarchical structure of HDF5 files}
-%\label{fig:hdf5-sample}
-%\vspace{-0.25in}
-%\end{figure}
 
@@
 \end{itemize}
-
-%For example, as shown in Figure \ref{fig:hdf5-sample}, the file Sample.h5
-%contains the root group which itself contains a group G1 and two datasets, D1
-%and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1.
-%The objects and the relationships between them can be represented as a B-tree,
-%which is used internally by HDF5 to index its objects.
 
@@
 \end{figure}
-
-%\subsection{IOD VOL plugin}
-%\label{iod-vol-plugin}
 
@@
 dynamically defined functions, a different system will be required leveraging
-something like C-on-demand~\cite{abbasi:2011:c-on-demand} or some other dynaimc
+something like C-on-demand~\cite{abbasi:2011:c-on-demand} or some other dynamic
 deployment and compilation or an interpreter system.
@@
 can be deployed more easily on smaller clusters or for those with more
-constrainted budgets. For simplicity, the rest of this section will describe a
+constrained budgets. For simplicity, the rest of this section will describe a
 full stack including all of the proposed IOD components.
@@
 
-%\subsection{IO Dispatcher API}
-%\label{sec:fast-forwards-io-dispatcher-api}
-%
-%The Fast Forward IO and Storage project is aimed at integrating many of
-%the services existing currently in middleware libraries in a
-%next-generation storage stack, exposed through a non-POSIX,
-%transactional, asynchronous, object-based API.
-%
-%From a high-level point of view, the Fast Forward's IO Dispatcher (IOD)
-%interface allows the user to:
-%
-%\begin{itemize}
-%\itemsep1pt\parskip0pt\parsep0pt
-%\item
-%  transactions
-%\item
-%  asynchrony
-%\item
-%  objects
-%\item
-%  layout
-%\item
-%  placement
-%\item
-%  formats
-%\end{itemize}
-%
-%Type of objects:
-
 \subsection{FFSIO Data Model Types}
@@
 Attributes are stored in a key-value store object, but use the
-multi-dimnensional array and blob objects to store the values for the
+multi-dimensional array and blob objects to store the values for the
 attributes.
@@
 a logically complete portion of a data set on a single storage target. This is
-similiar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
+similar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
 the data in the default or some other structured way as well as ``re-sharding''
@@
   \textbf{contiguous}. fixed chunking, distributed in a round-robin
-  fashion accross the IO nodes.
+  fashion across the IO nodes.
 \item
@@
 All of these distinct alternatives result in having many different ways for
-executing the same analysis task.
-%In the future, we plan to extend our
-%design to incorporate an optimization phase that will prepare a task by
-%taking into account the distinct possible ways of executing and select
-%the most promising one.
-In the subsequent discussions, we consider only
-data-movement optimization, i.e., sending the analysis code as close as
-possible to the data. In practice, this means we focus on identifying
-sharding of datasets and execute code accordingly over the appropriate
-shards.
-
-%At the DAOS layer, the sharding can be changed based on the type of storage
-%devices beneath it.
+executing the same analysis task.  In the subsequent discussions, we consider
+only data-movement optimization, i.e., sending the analysis code as close as
+possible to the data. In practice, this means we focus on identifying sharding
+of datasets and execute code accordingly over the appropriate shards.
 
@@
 \vspace{-0.20in}
-\caption{Exascale Architecture}
+\caption{Extreme Scale Architecture}
 \label{fig:exa-arch}
@@
 such as SSDs, these nodes are capable of better regulating the IO pressure on
-the underlying storage array better than simple forwaring gateways. For this
+the underlying storage array better than simple forwarding gateways. For this
 project, using something like SSDs on the IO nodes is termed a {\em Burst
-Buffer} and is disucssed below.
-
-%In current proposals, the stack is managed by middleware that sits between the
-%application and the parallel file system. Since most of the applications assume
-%a POSIX interface to storage, existing middleware either transparently handle
-%the IO operations, appearing as regular POSIX calls to applications; or modify
-%the IO API as little as possible in order to minimize the impact on production
-%codes.
-%
-%As we move towards the exascale goal, this way of interfacing with storage
-%becomes more of an obstacle. Many features provided by the stack are hidden by
-%the middleware/POSIX layers. If these were visible to the applications, and if
-%the applications were able to control how the underlying layers behave, they
-%would benefit greatly since the domain knowledge that is available at the
-%application level could be used to execute storage operations efficiently. One
-%more implication of the 25-year-old POSIX API is in cumbersome middleware,
-%since metadata has to be part of the same uni-dimensional byte array being
-%stored.
-%
-%The Fast Forward Storage IO project is aimed at merging the features of
-%existing middleware into a next-generation parallel file system stack. This new
-%storage API termed the IO dispatcher (IOD) is object-based, transactional,
-%asynchronous and IO-staging-aware (Figure~\ref{fig:exa-stack} (right)).
-
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=\columnwidth]{images/exa-stack.png}
-%\caption{Top-to-bottom view of the component stack}
-%\label{fig:exa-stack}
-%\end{figure}
-
-%The prototype stack, from a top-down point of view, has the following
-%components:
-%
-%\begin{enumerate}
-%\def\labelenumi{\arabic{enumi}.}
-%\itemsep1pt\parskip0pt\parsep0pt
-%\item
-%  Arbitrarily Connected Graphs (ACG).
-%\item
-%  HDF5 extensions.
-%\item
-%  IO Dispatcher (IOD).
-%\item
-%  Distributed Asynchronous Object Storage (DAOS).
-%\item
-%  Versioning Object Storage Device (VOSD).
-%\end{enumerate}
-%
-%Applications running on the compute nodes will be written in Python scripts or
-%in C++ using the HDF5 APIs. These applications will execute analysis on
-%graph-based data (ACG). The point is to demonstrate both the HPC and BigData
-%use cases of the Exascale architecture. The data structures will be stored in
-%HDF5, for which new API extensions will be implemented. These new API calls
-%will expose the transactional, asynchronous and function-forwarding semantics
-%of the underlying stack.
-
-%As mentioned before, the stack provides \emph{non-POSIX, object-based,
-%transactional and asynchronous active-storage}, meaning that POSIX is
-%supplanted by new object interfaces that reach up to the HDF5 layer
-%(Application IO in the stack diagram); transactional semantics are present,
-%from HDF down to the VOSD layer (Storage layer); clients don't need to wait for
-%any blocking operation; and analysis can be shipped and executed on IO or DAOS
-%nodes.
+Buffer} and is discussed below.
 
@@
 architecture~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
-%The current Fast Forward IOD design recommends incorporating SSDs, but
-%specifically lists these devices as optional. Unfortunately, not incorporating
-%burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
-%the IOD design currently is written assuming burst buffers. This means that the
-%bulk of IO operations will only hit the IOD layer and proposed functionality,
-%such as the function shipping, do not discuss the impact on the DAOS layer or
-%the operators themselves should a burst buffer not be available.  Consider the
-%important functionality of data rearrangement and doing things like changing
-%the fast array dimension on shared, spinning media.  The design of DAOS assumes
-%that it will only be involved when persisting a completed transaction and only
-%for a fraction of the total transactions created.  Transaction frequency at the
-%IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
-%is the process of transferring the set of copy-on-write changes from
-%transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
-%mirrors the copy-on-write functionality, but stores the differences between
-%epochs rather than individual transactions. The primary difference is that at
-%the DAOS layer, data is stored in large, logically contiguous chunks to manage
-%the metadata load and hopefully improve read performance.
-
-%One of the bigger concerns is the observation that the original data staging
-%proposals all used compute nodes while the newer proposals seek not only to
-%make them a fixed portion of the IO stack, but also shared across all machine
-%users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
-%potential costs and advantages of where to place operators similar to the IOD
-%proposed function shipping. There are two key takeaways from PreDatA. First,
-%placement matters.  Depending on the communication intensity vs. computation
-%intensity, where along the IO path to place the operation can matter
-%significantly. Second, and more importantly, the amount of time spent
-%processing for the operators was stretched to the point where it consumed
-%nearly all of the time between IO operations. The given ratios of compute
-%processes to staging process examined is representative for future extreme
-%scale platforms. If anything, the ratios offer more staging processes than IOD
-%processes would be available.
-%
 
@@
 
-%This is unlikely to be useful
-%because of the limited compute and communication capacity to spare to perform
-%these operations at a bottleneck in the IO path.  The use of a separate data
-%staging area intentionally separate from the IO path allows using operators on
-%limited resources leaving the IO path clear for strictly data movement. A
-%nuance of this design is discussed in the Design Philosophy below.
-
 Future work will generalize this model to potentially support dynamic IO node
@@
 than a capacity cluster where end-to-end performance is a lesser concern.
-
-%By concentrating the Burst Buffers and function shipping into the storage
-%stack, three problems arise.  First, the amount of network bandwidth, IO
-%bandwidth, and compute power consumed for example operations from a single
-%application is likely to completely monopolize the IOD processes. Second, if
-%space and time partitioning is used instead, the functionality risks being too
-%small to be useful. Third, the long-term hardware performance advantage for
-%SSDs is questionable. Recent studies have shown that the erase-before-write
-%and interference between reading and writing with flash-based SSDs can cause
-%severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
-%of an optional use SSD layer in the new Trinity machine at Los Alamos will
-%offer a test bed to determine how likely these observed problems would affect a
-%production extreme scale platform.
-%
-%Given these features, the optionality and even incorporation of burst buffers
-%in the current design should be carefully considered. Much of the advanced, key
-%functionality proposed as they are currently designed ultimately relies on the
-%existence of burst buffers to work. Further thought about how to have an IOD
-%layer both with and without a burst buffer is required before they can be
-%considered optional. As the design stands today, they are a required part of
-%the IOD layer for proper functioning. Unfortunately, it is not clear that they
-%can address the performance concerns they are intended to cover.
 
@@
 
-%Another concern that is acknowledged, but no thought has been applied to, is
-%the requirement that a single IOD process of the set assigned to an application
-%is the master for any operation.  Should the number of concurrent applications
-%exceed the available nodes, sharing an IOD process will be required. The
-%requirements both in terms of scheduling and resource management were cut from
-%the project due to funding limitations. Since partitioning of the IOD processes
-%for exclusive use by particular applications is the assumed operating mode,
-%should insufficient IOD resources be available, either a job could be delayed
-%or IOD resources could be reallocated from a different process could be
-%redeployed for use by the new job. Handling resilience concerns for the IOD
-%processes must also be address.  These sorts of considerations still need to be
-%made for a full production system.
-
 \section{DAOS Layer}
@@
 VOSD layer). The two key design elements of this layer are the handling of
-epochs and the mapping of conatiners and objects to the underlying storage.
+epochs and the mapping of containers and objects to the underlying storage.
 
@@
 mapping would be unnecessary. The current projections suggest that a standard
-POSIX-like file system will likey be used at the lowest level on each storage
+POSIX-like file system will likely be used at the lowest level on each storage
 device requiring the mapping at some level. To perform this mapping, DAOS
@@
 older epochs since those have been committed.
-
-%Epochs are represented through the shard mechanism. Each container shard is
-%versioned according to the transactions in the IOD layer that are persisted to
-%the DAOS layer as an epoch. The container shard is a piece of the container
-%with the contained objects concept. Since the container is largely logical,
-%this generally maps to a shard of an object within a container.
 
@@
 and the versioning aspects of transactions/epochs. For every DAOS shard, the
-VOSD has information for storing and accessing the currently commited version,
-the Highest Committed Epoch, as well as a staging dataset represeting the
+VOSD has information for storing and accessing the currently committed version,
+the Highest Committed Epoch, as well as a staging dataset representing the
 next version of the object being stored. Both of these are combined in a shard
@@
 
-For data integrity, an intent log is maiained as part of the underlying file
+For data integrity, an intent log is maintained as part of the underlying file
 system enabling fault recovery.
@@
 In the subsections that follow, we examine transactions and epochs, metadata
-management, function and analysis shipping, and the aribtrary connected graphs
+management, function and analysis shipping, and the arbitrary connected graphs
 support.
-
-%At a broader level, there are some concerns that were partially clarified
-%through conversations with the team.  Consider a shared file system across an
-%HPC data center. The current design maintains the metadata in its own
-%container. Since copying data from the IOD layer to the DAOS layer requires an
-%explicit persist call, how and when synchronizing the metadata across the
-%layers and potentially across machines occurs is undefined. Delaying
-%synchronization until an explicit persist is called will reduce the update
-%frequency, but delays the data visibility on other platforms. Ideally, the
-%metadata object would need to be automatically persisted every time a container
-%transaction is persisted to the DAOS layer.
-%
-%The implication of this is that every transaction persist is double operation
-%to account for the metadata persist. More importantly, the IOD-layer version of
-%the metadata container may contain readable transactions that have not been
-%persisted to the DAOS layer. How to handle this inconsistency between the two
-%layers still needs to be explored.
-%
-%A point of confusion rather than a potential design challenge is the change in
-%definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
-%container is a collection of objects. For the DAOS layer, a container is a
-%collection of objects across a set of shards. For the IOD an object may be a
-%shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
-%Having the same names with locally correct, globally conflicting definitions
-%serves to confuse how the system should work.
 
@@
 different functionality, but are connected as is explained below.
-%How these differ from the \DDT approach is also explored.  While IOD's and
-%\DDTns's transactions are seemingly very different, they use a similar
-%high-level design, but very different implementation, to solve the same
-%problem.
 
@@
 IOD transactions are necessarily persisted. Maintaining this ID continuity is
-critial for multiplatform use. Since the shared point is the DAOS layer, any
+critical for multiplatform use. Since the shared point is the DAOS layer, any
 user adding a new version to a file must be able to determine the most recent
@@
 treating it just as data is a central design goal of the Fast Forward project.
-This is a hybrid approach to metadata management that is half-way between
+This is a hybrid approach to metadata management that is halfway between
 providing no inherent metadata support and having a fully integrated, but
@@
 
-%Based on the lessons from the \DDT metadata
-%service~\cite{lofstead:2012:txn-metadata} construction and the prior
-%experiments with LWFS, having a completely separate metadata service is
-%workable. Rather than making it a bottleneck in the IO path, it is another
-%service that users must interact with if they need those services.  Users can
-%manage everything by maintaining the metadata including the list of objects
-%themselves. However, there are drawbacks to this 100\% client-side approach.
-%
-%With a client-side only approach, there is a serious risk of the metadata
-%service and the object store getting out of sync.  While having a metadata-less
-%object storage service is desirable, the different semantics from traditional
-%file systems requires some considerations. In this case, should these services
-%get out of sync, three particular risks are introduced.  First, a client could
-%create a dangling entry in the metadata service that does not correspond to any
-%objects in the object store. Second, a client could create orphaned objects
-%that have no associated metadata entries. Third, updates to the metadata or
-%object store service should be an atomic operation, but due to a lack of
-%coordination, a window where the system is inconsistent appears.
-%
-%Ultimately, the consistency semantics required must be determined. If a
-%metadata service is required and it must be in sync with the object storage
-%service, then additional work must be performed. In traditional file systems,
-%the metadata and object storage updates are atomic. With decoupling metadata
-%from object storage, should this atomicity still be desired, it requires both
-%the ability for the services to participate in a task that is part of a larger
-%atomic operation and a higher-level mechanism to manage the atomic operation.
-%
-%Overall, while additional work is required to maintain a client-side only
-%metadata service, it eliminates any potential bottlenecks related to updating
-%metadata related to the object storage. The burden of tracking striping and
-%other metadata that has traditionally been part of the metadata associated with
-%the file system will have to be maintained by the object storage service. The
-%lack of a centralized, serialized bottleneck to store that information improves
-%concurrency.
-
-%\subsection{Comparison to \DDTns}
-%The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
-%approach for handling ACID-style transactions in an environment with parallel
-%clients and multiple servers (doubly distributed). Rather than being aimed
-%solely at data movement operations, \DDT seeks to address the general problem
-%of managing any operation with multiple clients and servers.  Consider the
-%management of the analysis/visualization area, potentially similar to the IOD
-%concept. The transaction protocol is used to help manage resizing of the
-%resource allocation to the various analysis and visualization components.  For
-%the purposes of this discussion, \DDT could also be used to manage changing how
-%IOD processes and/or nodes are used without exposing these changes to the
-%client processes prematurely.  This has been described and analyzed
-%previously~\cite{dayal:2013:io-containers}.
-%
-%The example metadata and data storage services created as part of the \DDT
-%project have no dependencies between transactions that prevent visibility
-%should an older version be incomplete. This additional, intentional requirement
-%by IOD offers different functionality than \DDTns's example services. In the
-%case of \DDTns, the functionality is more minimal, but also avoids some of the
-%concerns outlined below.
-%
-%The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
-%scalability issues and demonstrated a scalable client-side coordination model
-%with excellent performance. The performance measured for a complex transaction
-%with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
-%explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
-%
-%\begin{figure}[ht]
-%%\vspace{-0.15in}
-%\centering
-%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
-%\vspace{-0.15in}
-%\caption{Total Transaction Overhead}
-%\label{fig:performance}
-%%\vspace{-0.15in}
-%\end{figure}
-%
-%At a high level, both \DDT and the IOD transactions have the same design. In
-%both cases, a hierarchical model is employed. In the case of \DDTns, it is a
-%purely client-side tree using semi-synchronous messaging. The messaging itself,
-%in the current implementation, uses asynchronous MPI messages. The synchronous
-%component comes from the timeout mechanism used to detect faults.  It forces a
-%level of coordination and synchronization for the protocol. For IOD, it is a
-%server-side tree and fully asynchronous relying on the existing Lustre fault
-%detection mechanism for failure detection. In both cases, there is a master in
-%charge of managing the transaction and a collection of workers that aggregate
-%into the master through second-level leaders. Beyond that, there are some
-%significant differences. Some of the different choices made by IOD raise some
-%possible concern.
-%
-%The multi-leader model introduces the possibility of forcing a rollback of an
-%entire transaction when a partial retry might be sufficient for success. Since
-%the transactions are managed at a high level rather than the individual
-%tasks, a failure in a limited distributed task can cause the entire transaction
-%to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
-%10 have an additional shared task to complete. If the task shared by the 3
-%processes fails on any of the three, the entire transaction would roll back
-%because it is a coarse-grained success/failure. If a concept like
-%sub-transactions at a task granularity were used, then it would be possible for
-%the one process that failed to report just that failure. Then the transaction
-%manager could reassign the resources for these 3 processes and try just that
-%operation again. If it now succeeds, then the overall transaction can be marked
-%successful only redoing the minimum amount of work required.
-%
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
-%\vspace{-0.15in}
-%\caption{Optimized Protocol}
-%\label{fig:optimized-protocol}
-%%\vspace{-0.25in}
-%\end{figure}
-%
-%\DDT has addressed theses issues in a couple of ways. First, the
-%sub-coordinators each have a list of processes from which they expect messages.
-%Should a message be missed, it is noticed and corrective action can be taken.
-%Second, \DDT has the concept of sub-transactions. The messaging requirements
-%are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
-%represent finer grained operations than the entire output, \DDT can manage
-%multiple writes per client by using a sub-transaction to represent the output
-%for any item to the file (container). Because of how the sub-transactions are
-%managed, the singleton sub-transactions, ones in which only a single process
-%participates, must be declared before the transaction begins so that its
-%existence can be broadcast as part of the begin transaction message. This
-%ensures there is global knowledge that the sub-transaction is expected.  That
-%way if the coordinator (transaction leader) fails, whichever process takes over
-%that role knows to expect a completion message for that sub-transaction or the
-%overall transaction cannot complete. Global sub-transactions can be defined at
-%any time since they are a global, synchronized operation broadcasting their
-%existence. While this additional layer does introduce messaging, the overhead
-%is quite small.
-%
-%The advantages of eliminating these messages is not performance as demonstrated
-%by the performance of \DDTns. Instead, it offers a much less synchronous model
-%that matches with different programming models, such as Charm++ or other
-%task-based approaches. Since it can work for the bulk-synchronous model also,
-%it is a more broadly applicable approach. This assumes that the observed
-%potential issues can be addressed successfully.
-
 \subsection{Function and Analysis Shipping}
@@
 communication model.  Every ION runs an IOFSL (IO Function Shipping Layer)
-server. The IOFSL client is integrated into the HDF5 library which runs on each
+server. The IOFSL client is integrated into the HDF5 library running on each
 CN. A client can forward requests to any number of IONs. Every IO operation
@@
 processing. GraphBuilder~\cite{Jain:2013:GraphBuilder} is a tool to generate
-effective graph paritions reducing the load for using MapReduce to process
+effective graph partitions reducing the load for using MapReduce to process
 graph data sets. GraphLab~\cite{Low:2012:GraphLab} offers a way to process
@@
 
-In order to make both of these tools work on top of the exascale stack, they
-both have to be modified. After these modifications are implemented,
+In order to make both of these tools work on top of the extreme scale stack,
+they both have to be modified. After these modifications are implemented,
 GraphBuilder will be able to write the partitioned graph in the newly proposed
@@
 \item
-  efficient data sharing ammong CNs
+  efficient data sharing among CNs
 \item
@@
 \label{sec:evaluation}
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+\vspace{-0.05in}
+\end{figure*}
 
@@
 National Laboratory. It consists of 84 nodes each with dual, 8 core Intel
-Xeon ES-2670 CPUs at 2.6 GHz and 32 GB RAM. The inteconnect is Cray Gemini.
-The storage array are three sets of SSD-based lustre file systems.
+Xeon ES-2670 CPUs at 2.6 GHz and 32 GB RAM. The interconnect is Cray Gemini.
+The storage array are three sets of SSD-based Lustre file systems.
 
@@
 
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
-\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Number of Hosts}
-\label{fig:eval-hosts}
-\vspace{-0.05in}
-\end{figure*}
-
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
-\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Data Sizes}
-\label{fig:eval-size}
-\vspace{-0.05in}
-\end{figure*}
-
 \section{Conclusions}
@@
 addressing the requirements for an extreme scale data storage mechanism.  By
-prefering a high level user API like HDF5 rather than using the POSIX
+preferring a high level user API like HDF5 rather than using the POSIX
 interface, more advanced functionality can be incorporated with less end-user

### 2014-05-12 paper.tex
new Buffy machine specs.
finished adding the machine specs

@@
 \includegraphics[width=\columnwidth]{images/vol-arch.png}
-\vspace{-0.10in}
+\vspace{-0.20in}
 \caption{Architectural view of the VOL abstraction mechanism}
@@
 internal data types, such as datasets, these concepts must be mapped onto the
-proposed FFSIO data storage concepts. This mapping is shown in Table~\ref{tab:mapping}
-
-\begin{table}[ht]
-    \vspace{-0.10in}
-    \centering
-    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
-    \bigskip
-    \vspace{-0.15in}
-
-    \begin{tabular}{|r|r|}
-\hline
-HDF5 & FFSIO\\
-\hline
-file & container \\
-dataset & array \\
-group & key-value store \\
-attribute(s) & key-value store \\
-\hline
-    \end{tabular}
-    \label{tab:mapping}
-\end{table}
-
+proposed FFSIO data storage concepts. At a high level, think of the HDF5 types
+mapping to FFSIO types as follows: file $\rightarrow$ container, dataset $\rightarrow$
+array, and groups and attributes $\rightarrow$ key-value store.
+%This mapping is shown in Table~\ref{tab:mapping}
+
+%\begin{table}[ht]
+%    \vspace{-0.10in}
+%    \centering
+%    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
+%    \bigskip
+%    \vspace{-0.15in}
+%
+%    \begin{tabular}{|r|r|}
+%\hline
+%HDF5 & FFSIO\\
+%\hline
+%file & container \\
+%dataset & array \\
+%group & key-value store \\
+%attribute(s) & key-value store \\
+%\hline
+%    \end{tabular}
+%    \label{tab:mapping}
+%\end{table}
+%
 In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.
@@
 
-All of the tests are performed on the Buffy Cray test cluster at Los Alamos
-National Laboratory. It consists of 84 nodes each with dual, 8 core Intel
-Xeon ES-2670 CPUs at 2.6 GHz and 32 GB RAM. The interconnect is Cray Gemini.
-The storage array are three sets of SSD-based Lustre file systems.
+All of the tests are performed on the Buffy Cray XC30-AC at LANL.  It consists
+of 64 compute nodes each with dual, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz
+and 64 GB RAM.  The interconnect is Cray Aries. There are 14 IO nodes
+consisting of single socket, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz with 32
+GB of RAM.  There are also a metadata, login, and 2 boot nodes.  The storage
+array has disk and SSD paritions. One is a DDN Lustre system with 192 TB disk
+usable with a minimum of 5 GB/sec performance. The rest are SSDs consisting of
+an EMC flash array connected via FDR InfiniBand with 22 TB and 48 GB/sec write
+performance.
 

### 2014-05-13 paper.tex
added NFS reference and fixed attribute notion error.

@@
 
-Other file systems, like GoogleFS~\cite{ghemawat:googlefs}, address distributed
-rather than parallel computing and cannot be compared directly. The primary
-difference between distributed and parallel file systems is the ability of the
-file system to store and retrieve data simultaneously from multiple clients, in
-parallel, and treat the resulting collection of pieces as a single object.
-Distributed file systems rely on a single client creating a file, but
-distributing the set of files across a wide array of storage devices. The
-other, popular distributed file system of note is
-HDFS~\cite{Shvachko:2010:hdfs} that is distributed as part of Hadoop. These
-other file systems are mainly of interest in the context of the ACG features of
-FFSIO and will be discussed more in Section~\ref{sec:acg}.
+Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
+HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
+computing and cannot be compared directly. The primary difference between
+distributed and parallel file systems is the ability of the file system to
+store and retrieve data simultaneously from multiple clients, in parallel, and
+treat the resulting collection of pieces as a single object.  Distributed file
+systems rely on a single client creating a file, but distributing the set of
+files across a wide array of storage devices. The other, popular distributed
+file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
+decades for enterprise file systems.  These other file systems are mainly of
+interest in the context of the ACG features of FFSIO and will be discussed more
+in Section~\ref{sec:acg}.
 
@@
 viewed as a directory tree of objects where each directory entry specifies
-either a sub-directory (group) or some data or attribute.
+either a sub-directory (group) or some data or attribute. For the IOD layer, it
+is a collection of objects. For HDF5, interpreting the objects builds the
+structure.
 
@@
 as additional key-value objects representing sub-groups for the hierarchy. Each
-of those key-value store objects works identically.
-
-Attributes are stored in a key-value store object, but use the
-multi-dimensional array and blob objects to store the values for the
-attributes.
+of those key-value store objects works identically.  Attributes are stored in a
+key-value store object.
+%, but use the
+%multi-dimensional array and blob objects to store the values for the
+%attributes.
 

### 2015-09-15 paper.tex
pared down to the PDSW limit

@@
-\documentclass[conference]{IEEEtran}
+\documentclass{sig-alt-gov2}
 \pdfpagewidth=8.5in
@@
 
-%\conferenceinfo{Cluster'14,} {September 17, 2014, Madrid, Spain}
-%\CopyrightYear{2014}
+%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
+%\CopyrightYear{2015}
 %\crdata{978-1-4503-1103-8/11/11}
@@
 
+\numberofauthors{1}
 \author{
-\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
-\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
+\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2], Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
+\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
+\footnotemark[2]~~University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu\\
+\footnotemark[3]~~HDF Group koziol@hdfgroup.org\\
+\footnotemark[4]~~EMC john.bent@emc.com\\
+\footnotemark[5]~~Intel eric.barton@intel.com
 }
@@
 The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
-project is going to have significant impact on storage systems design within
-and beyond the HPC community. With phase 1 of the project complete, it is an
-excellent opportunity to explore the complete design and how it will address
-the needs of extreme scale platforms.  This paper examines the overall design.
-Each layer of the proposed stack is discussed in some detail along with
-cross-cutting topics that impact multiple layers, such as transactions and
-metadata management. We also describe the proposed features to better support
-some classes of big data applications.
-
-This paper not only provides a timely summary of important aspects of the
-design specifications but also captures the underlying reasoning that is not
-available elsewhere. With this information, we encourage the broader community
-to understand the design, intent, and future directions to foster discussion
-guiding phase 2 and the ultimate production storage stack based on this work.
-An initial performance evaluation of the early prototype implementation is also
-provided to validate the presented design.
+project will significantly impact storage systems design within and beyond the
+HPC community. With phase 2 starting, it is an excellent opportunity to explore
+the complete design and how it will address the needs of extreme scale
+platforms.
+%Each layer of the proposed stack is discussed in some detail along
+%with cross-cutting topics that impact multiple layers, such as transactions and
+%metadata management.
+%We also describe the proposed features to better support
+%some classes of big data applications.
+Key underlying reasoning driving design
+decisions not available elsewhere are presented.  We encourage the broader
+community to understand the design, intent, and future directions to foster
+discussion guiding phase 2 and the ultimate production storage stack based on
+this work.  An initial performance evaluation of the prototype implementation
+validates the presented design.
 
@@
 
-Current production HPC IO stack design is unlikely to offer sufficient features
-and performance to adequately serve extreme scale science platform
-requirements.  Adding to the problem complexity is the variety of Big Data
-problems users want to address using these platforms. Trying to address both of
-these domains is a daunting problem.
-
-To address these challenges, a joint effort between the US Department of
+%Current production HPC IO stack design is unlikely to offer sufficient features
+%and performance to adequately serve extreme scale science platform
+%requirements.  Adding to the problem complexity is the variety of Big Data
+%problems users want to address using these platforms. Trying to address both of
+%these domains is a daunting problem.
+
+%To address these challenges,
+A joint effort between the US Department of
 Energy's Office of Advanced Simulation and Computing and Advanced Scientific
@@
 set~\cite{fastforward:2014:docs} for a future IO stack to address the
-identified challenges. The first phase recently completed with a second phase
-currently getting underway as of early 2014. The core focus of the first phase
+identified challenges. The core focus of the first phase
 was basic functionality and design. While an idealized potential system would
 be the perfect target architecture, the reality of budgets has tempered many of
-the decisions. For example, extensive availability of NVRAN or SSDs on all of
-the compute nodes is currently not economically feasible limiting some of the
-potential design choices.  With this in mind, the second phase will refine this
-design incorporating fault recovery and other features missing from the first
-phase.
+the decisions.
+%For example, extensive availability of NVRAM or SSDs on all of
+%the compute nodes is currently not economically feasible limiting some of the
+%potential design choices.  With this in mind, the second phase will refine this
+%design incorporating fault recovery and other features missing from the first
+%phase.
 
@@
 
-Overall, the architecture shifts from the idea of files and directories to
-containers of objects. This shift avoids the bottlenecks related to the POSIX
-files and directories structure such as the serialization of file creation, the
-limitation and impact of the number of files in a directory, and the limited
-semantics of a byte stream. Instead, the new interface focuses on high-level
-data models and their properties and relationships. This concept permeates the
-entire IO stack.
-
-In addition to addressing the traditional scientific workload, this project
-seeks to expand functionality to better support Big Data type applications. The
-key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
-Map-Reduce systems. Key system features are introduced to efficiently support
-these computing models in addition to the typically bursty IO loads of more
-traditional HPC applications.
+%Overall, the architecture shifts from the idea of files and directories to
+%containers of objects. This shift avoids the bottlenecks related to the POSIX
+%files and directories structure such as file creation serialization, the
+%file count in a directory limitation and impact, and the limited
+%semantics of a byte stream. Instead, the new interface focuses on high-level
+%data models and their properties and relationships. This concept permeates the
+%entire IO stack.
+
+%In addition to addressing the traditional scientific workload, this project
+%seeks to expand functionality to better support Big Data type applications. The
+%key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
+%Map-Reduce systems. Key system features are introduced to efficiently support
+%these computing models in addition to the typically bursty IO loads of more
+%traditional HPC applications. These features are not discussed for space
+%reasons.
 
@@
 \includegraphics[width=\columnwidth]{images/arch-mapping}
-\vspace{-0.15in}
-\caption{Target Architecture and Component Mapping}
+\vspace{-0.20in}
+\caption{Architecture and Component Mapping}
 \label{fig:arch-mapping}
@@
 working with the lower layers and to enable advanced functionality that would
-require more direct user intervention. This layer incorporates the necessary
-features for ACGs from a end-user's perspective.
+require more direct user intervention.
+%This layer incorporates the necessary
+%features for ACGs from a end-user's perspective.
 
@@
 Buffers to insulate the persistent storage array from bursty IO workloads.
-Ideally, the IOD layer's functionality can be optional based on available
-hardware and compute power provided on the IO Nodes (IONs). Much of the
-functionality offered at this layer would shift either up or down the stack as
-discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
+%Ideally, the IOD layer's functionality can be optional based on available
+%hardware and compute power provided on the IO Nodes (IONs). Much of the
+%functionality offered at this layer would shift either up or down the stack as
+%discussed in detail below.
+The Distributed Asynchronous Object Storage (DAOS)
 layer serves as the persistent storage interface and translation layer between
@@
 foundation on which everything else is built with no dependence on any
-technologies specified above it (in dark pink and yellow). For example, the IOD
-layer with or without burst buffers is not required for DAOS to operate
-properly.  Instead, the DAOS layer can handle all of the IO operations from the
-user API layer, albeit with the potential performance penalty of manipulating
-the shared, persistent storage array. At the bottom is the Versioning Object
+technologies specified above it (in dark pink and yellow).
+%For example, the IOD
+%layer with or without burst buffers is not required for DAOS to operate
+%properly.  Instead, the DAOS layer can handle all of the IO operations from the
+%user API layer, albeit with the potential performance penalty of manipulating
+%the shared, persistent storage array.
+At the bottom is the Versioning Object
 Storage Device (VOSD) (in purple).  It serves as the interface for storing
 objects of all types efficiently for each storage device in the parallel
-storage array. Think of this layer as the physical disk interface layer. In
-terms of Lustre, this would replace the API on individual storage devices with
-an interface friendlier to the containers of objects and transactions/epochs
-concepts used in the higher layers.
-
-Along with analysis of the published design documents, a discussion of the
-design philosophy representing the overall intent is presented. This
-information represents information that may or may not have been written down,
-but is the intent of ultimate product.  These ideas are presented to give a
-fuller picture of where the project is going rather than dwelling on any
-limitations of the published documents. This is most important to illustrate
-how different concepts will work across layers since that information is spread
-across multiple documents and may lack a cohesive overall view.
+storage array.
+%Think of this layer as the physical disk interface layer. In
+%terms of Lustre, this would replace the API on individual storage devices with
+%an interface friendlier to the containers of objects and transactions/epochs
+%concepts used in the higher layers.
+
+%Along with analysis of the published design documents, a discussion of the
+%design philosophy representing the overall intent is presented. This
+%information represents information that may or may not have been written down,
+%but is the intent of ultimate product.  These ideas are presented to give a
+%fuller picture of where the project is going rather than dwelling on any
+%limitations of the published documents. This is most important to illustrate
+%how different concepts will work across layers since that information is spread
+%across multiple documents and may lack a cohesive overall view.
 
@@
 discusses the programmatic interface end users will see when interacting with
-the storage array. This will be discussed in the context of the HDF5 based
-example library used for the functionality demonstration. Section~\ref{sec:iof}
+the storage array.
+%This will be discussed in the context of the HDF5 based
+%example library used for the functionality demonstration.
+Section~\ref{sec:iof}
 briefly discusses the motivation and proposal for the IO forwarding layer.
 Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
-functionality it offers. This will detail the pieces of the layer that are
-potentially optional and mention the cross-cutting features discussed in a
-later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
-layer functions. As with the IOD layer, the cross-cutting features will be
-mentioned, but discussed more fully in the cross-cutting section. The VOSD
-layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
-between the DAOS and VOSD layers are explored as it pertains to the physical
-storage. Next is an exploration of cross-cutting features like transactions and
-metadata management in Section~\ref{sec:broader}. Since these and other
-features are spread across multiple layers, it makes more sense to discuss them
-independently once an understanding of the overall structure has been
-presented.  A demonstration of the functionality is presented in
-Section~\ref{sec:evaluation}. This shows that the prototype system based on the
-proposed design can function. The paper is concluded in
+functionality it offers.
+%This will detail the pieces of the layer that are
+%potentially optional and mention the cross-cutting features discussed in a
+%later, cross-cutting section.
+Section~\ref{sec:daos} discusses how the DAOS
+layer functions.
+%As with the IOD layer, the cross-cutting features will be
+%mentioned, but discussed more fully in the cross-cutting section.
+The VOSD layer is discussed in Section~\ref{sec:vosd}.
+%In particular, the mapping
+%between the DAOS and VOSD layers are explored as it pertains to the physical
+%storage.
+Next is an exploration of cross-cutting features like transactions and
+metadata management in Section~\ref{sec:broader}.
+%Since these and other
+%features are spread across multiple layers, it makes more sense to discuss them
+%independently once an understanding of the overall structure has been
+%presented.
+A demonstration of the functionality is presented in
+Section~\ref{sec:evaluation}.
+%This shows that the prototype system based on the
+%proposed design can function.
+The paper is concluded in
 Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
@@
 
-Many projects over the last couple of decades have sought to address some
-challenging aspect of parallel file system design. The recent rise of ``Big
-Data'' applications with different characteristic IO patterns have somewhat
-complicated the picture. Extreme scale machines will be expected to handle both
-the traditional simulation-related workloads as well as applications more
-squarely in the Big Data arena. This will require some adjustments to the
-underlying system for good performance for both scenarios.
-
-The major previous work is really limited to full file systems rather than the
-mountain of file system refinements made over the years. A selection of these
-other file systems and some features that make it relatively unique are
-described below.
+%Many projects over the last couple of decades have sought to address some
+%challenging aspect of parallel file system design. The recent rise of ``Big
+%Data'' applications with different characteristic IO patterns have somewhat
+%complicated the picture. Extreme scale machines will be expected to handle both
+%the traditional simulation-related workloads as well as applications more
+%squarely in the Big Data arena. This will require some adjustments to the
+%underlying system for good performance for both scenarios.
+%
+%The major previous work is really limited to full file systems rather than the
+%mountain of file system refinements made over the years. A selection of these
+%other file systems and some features that make it relatively unique are
+%described below.
 
@@
 both a POSIX and object interface including features typically found in parallel
-file systems. Ceph's unique striping approach uses pseudo-random numbers with a
-known seed eliminating the need for the metadata service to track where each
-stripe in a parallel file is placed.
-
+file systems.
+%Ceph's unique striping approach uses pseudo-random numbers with a
+%known seed eliminating the need for the metadata service to track where each
+%stripe in a parallel file is placed.
 PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
-such as a single process opening a file and sharing the handle. It has been
-commercialized in recent years as OrangeFS.
-
+such as a single process opening a file and sharing the handle.
+%It has been
+%commercialized in recent years as OrangeFS.
 Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
@@
 programmatic control over how data is placed in the storage system.
-
 GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
@@
 IBM.
-
 Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
@@
 HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
-computing and cannot be compared directly. The primary difference between
-distributed and parallel file systems is the ability of the file system to
-store and retrieve data simultaneously from multiple clients, in parallel, and
-treat the resulting collection of pieces as a single object.  Distributed file
-systems rely on a single client creating a file, but distributing the set of
-files across a wide array of storage devices. The other, popular distributed
+computing and cannot be compared directly.
+%The primary difference between
+%distributed and parallel file systems is the ability of the file system to
+%store and retrieve data simultaneously from multiple clients, in parallel, and
+%treat the resulting collection of pieces as a single object.  Distributed file
+%systems rely on a single client creating a file, but distributing the set of
+%files across a wide array of storage devices.
+The other, popular distributed
 file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
-decades for enterprise file systems.  These other file systems are mainly of
-interest in the context of the ACG features of FFSIO and will be discussed more
-in Section~\ref{sec:acg}.
+decades for enterprise file systems.
+%These other file systems are mainly of
+%interest in the context of the ACG features of FFSIO and will be discussed more
+%in Section~\ref{sec:acg}.
 
@@
 determining what are strictly necessary extensions to an existing IO API to
-support the new functionality.  The additional functionality, such as
-transactions, can be ignored for legacy implementations, but these applications
-will not be able to take advantage of the asynchronous IO support inherent to
-the new API.  The additions comprise (Figure~\ref{fig:vol-arch}):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-
-\item
-  API extensions to support new functionality provided by the FFSIO project.
+support the new functionality.
+%The additional functionality, such as
+%transactions, can be ignored for legacy implementations, but these applications
+%will not be able to take advantage of the asynchronous IO support inherent to
+%the new API.
+%The additions comprise (Figure~\ref{fig:vol-arch}):
+%
+%\begin{enumerate}
+%\def\labelenumi{\arabic{enumi}.}
+%\itemsep1pt\parskip0pt\parsep0pt
+%
+%\item
+\textbf{API extensions} to support new functionality provided by the FFSIO project.
   This includes calls for managing asynchronous request lists, performing
   asynchronous operations, creating and managing transactions, end-to-end data
-  integrity, and data type and functions to support the ACG functionality more
+  integrity, and data type and functions to support the big data oriented functionality more
   efficiently than the current API.
-
-\item\label{item:fs}
-  Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
+%
+%\item
+\label{item:fs}
+  \textbf{Function shipping} from Compute Nodes (CN) to IO Nodes (ION). This provides
   the application developer with the capability of sending computation down to
@@
   and data reorganization for more efficient retrieval.
-
-\item
-  Analysis Shipping from CN to IONs or DAOS nodes. This is similar
+%
+%\item
+  \textbf{Analysis Shipping} from CN to IONs or DAOS nodes. This is similar
   to~\ref{item:fs} but instead of returning the result over the network, it
@@
 
-\end{enumerate}
-
-Function and Analysis Shipping are part of the cross-cutting features and are
-discussed in Section~\ref{sec:fn-shipping}.
-
-HDF5~\cite{hdf5} has a versatile data model offering complex data objects
-and metadata. Its information set is a collection of datasets, groups,
-datatypes and metadata objects. The data model defines mechanisms for
-creating associations between various information items. The main
-components of HDF5 are described below.
-
-\begin{itemize}
-
-\item
-  \textbf{File}: In the HDF5 data model, the collection of data items stored
-  together is represented by a file. It is an object collection that also
-  describes the relationship between them. Every file begins with a root
-  group ``/'' serving as the ``starting-point'' in the object hierarchy.
-
-\item
-  \textbf{Group}: A group is an object allowing association between HDF5
-  objects. It is synonymous with directories in a file system. A group
-  could contain multiple other groups, datasets, datatypes or attributes within
-  it.
-
-\item
-  \textbf{Dataset}: HDF5 datasets are objects representing actual data
-  or content. Datasets are arrays with potentially multiple dimensions. A
-  dataset is characterized by a dataspace and a datatype. The dataspace
-  captures the rank (number of dimensions) and the current and maximum
-  extent in each dimension. The datatype describes the type of its data
-  elements.
-
-\item
-  \textbf{Attribute}: Attributes are used for annotating HDF5 objects. They are
-  datasets themselves and are attached to existing objects.
-
-\end{itemize}
-
-\subsection{Virtual Object Layer}
-\label{virtual-object-layer}
-
-The Virtual Object Layer is an abstraction mechanism internal to the HDF5
-library~\cite{hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
-just below the public API. The VOL exports an interface that allows writing
-plugins for HDF5 enabling developers to handle data in ways other than writing
-to storage in an HDF5 format.  Plugin writers provide an implementation for a
-set of functions and are trusted to provide the proper semantics for the new
-environment. For example, data staging could be implemented in the VOL layer by
-replacing writing to disk in the HDF5 format to sending data to a data staging
-area using some messaging mechanism.
-
-For this project, rather than the default writing to disk in the HDF5 format,
-the VOL is used to interact with the IOD layer and the different concepts it
-offers without requiring all of the functionality be exposed to users. For
-example, the containers and objects concept is transparently mapped to the
-files and datasets existing HDF5 users are familiar with.  This reduces the
-difficulty porting applications to the new IO stack.
-
-\begin{figure}[htbp]
-\vspace{-0.10in}
-\centering
-\includegraphics[width=\columnwidth]{images/vol-arch.png}
-\vspace{-0.20in}
-\caption{Architectural view of the VOL abstraction mechanism}
-\label{fig:vol-arch}
-\vspace{-0.10in}
-\end{figure}
-
-The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
-(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
-running on the system's compute nodes. Using the VOL architecture, the IOD VOL
-plugin uses a function shipper (RPC library) to forward the VOL calls to a
-server component running on the IO nodes (IONs). This function shipping is the
-IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
-arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
-and executed at the IONs.
+%\end{enumerate}
+
+%Function and Analysis Shipping are part of the cross-cutting features and are
+%discussed in Section~\ref{sec:fn-shipping}.
+
+%HDF5~\cite{hdf5} has a versatile data model offering complex data objects
+%and metadata. Its information set is a collection of datasets, groups,
+%datatypes and metadata objects. The data model defines mechanisms for
+%creating associations between various information items. The main
+%components of HDF5 are described below.
+%
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%
+%\item
+%  \textbf{File}: In the HDF5 data model, the collection of data items stored
+%  together is represented by a file. It is an object collection that also
+%  describes the relationship between them. Every file begins with a root
+%  group ``/'' serving as the ``starting-point'' in the object hierarchy.
+%
+%\item
+%  \textbf{Group}: A group is an object allowing association between HDF5
+%  objects. It is synonymous with directories in a file system. A group
+%  could contain multiple other groups, datasets, datatypes or attributes within
+%  it.
+%
+%\item
+%  \textbf{Dataset}: HDF5 datasets are objects representing actual data
+%  or content. Datasets are arrays with potentially multiple dimensions. A
+%  dataset is characterized by a dataspace and a datatype. The dataspace
+%  captures the rank (number of dimensions) and the current and maximum
+%  extent in each dimension. The datatype describes the type of its data
+%  elements.
+%
+%\item
+%  \textbf{Attribute}: Attributes are used for annotating HDF5 objects. They are
+%  datasets themselves and are attached to existing objects.
+%
+%\end{itemize}
+
+%\subsection{Virtual Object Layer}
+%\label{virtual-object-layer}
+
+%As shown in Figure~\ref{fig:vol-arch} it is implemented
+%just below the public API. The VOL exports an interface that allows writing
+%plugins for HDF5 enabling developers to handle data in ways other than writing
+%to storage in an HDF5 format.  Plugin writers provide an implementation for a
+%set of functions and are trusted to provide the proper semantics for the new
+%environment. For example, data staging could be implemented in the VOL layer by
+%replacing writing to disk in the HDF5 format to sending data to a data staging
+%area using some messaging mechanism.
+%
+%For this project, rather than the default writing to disk in the HDF5 format,
+%the VOL is used to interact with the IOD layer and the different concepts it
+%offers without requiring all of the functionality be exposed to users. For
+%example, the containers and objects concept is transparently mapped to the
+%files and datasets existing HDF5 users are familiar with.  This reduces the
+%difficulty porting applications to the new IO stack.
+%
+%\begin{figure}[htbp]
+%\vspace{-0.10in}
+%\centering
+%\includegraphics[width=\columnwidth]{images/vol-arch.png}
+%\vspace{-0.20in}
+%\caption{VOL Abstration Architecture}
+%\label{fig:vol-arch}
+%\vspace{-0.10in}
+%\end{figure}
+%
+%The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
+%(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
+%running on the system's compute nodes. Using the VOL architecture, the IOD VOL
+%plugin uses a function shipper (RPC library) to forward the VOL calls to a
+%server component running on the IO nodes (IONs). This function shipping is the
+%IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
+%arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
+%and executed at the IONs.
 
@@
 
-Since HDF5 has traditionally offered an interface focused on files and the
+Since HDF5~\cite{hdf5} has traditionally offered an interface focused on files and the
 internal data types, such as datasets, these concepts must be mapped onto the
@@
 array, and groups and attributes $\rightarrow$ key-value store.
+Shifting from file-oriented to the FFSIO structure happens in a Virtual Object Layer (VOL) plug-in. The VOL is an abstraction mechanism internal to the HDF5
+library that allows replacing the file IO operations with other calls.
 %This mapping is shown in Table~\ref{tab:mapping}
@@
 %
-In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.
+FFSIO types are discussed in Section~\ref{sec:iod}.
 
@@
 The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
-the massive process count on the storage stack. A current trend mixing MPI with
-a threading library like OpenMP is addressing the same issue, but limited to
-handling the parallelism on a single node rather than multiple nodes. Projected
-extreme scale platforms will have far fewer storage stack end-points per
-compute process in which to receive requests and data. By reducing the number
-of simultaneous requests, delays can be reduced. This has been demonstrated for
-the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
-degree for accessing the storage devices
-themselves~\cite{lofstead:2010:io-variability}.  The BlueGene platform
+the massive process count on the storage stack.
+%A current trend mixing MPI with
+%a threading library like OpenMP is addressing the same issue, but limited to
+%handling the parallelism on a single node rather than multiple nodes. Projected
+%extreme scale platforms will have far fewer storage stack end-points per
+%compute process in which to receive requests and data. By reducing the number
+%of simultaneous requests, delays can be reduced. This has been demonstrated for
+%the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
+%degree for accessing the storage devices
+%themselves~\cite{lofstead:2010:io-variability}.
+The BlueGene platform
 incorporated dedicated hardware to perform this role. The proposed
@@
 purposes. First, the burst buffers work as a fast cache absorbing write
-operations that then trickles out to the central storage array. It can also
-be used to retrieve objects from the central storage array for more efficient
-read operations and offers data filtering to make client reads more efficient.
+operations or pre-staging read operations.
+%that then trickles out to the central storage array. It can also
+%be used to retrieve objects from the central storage array for more efficient
+%read operations and offers data filtering to make client reads more efficient.
 Second, it offers the transaction mechanism for controlling data set visibility
 and to manage faults that could expose an incomplete or corrupt data set to
-users. These transactions are local to the IOD layer until persisted to the
-DAOS layer eliminating the need for burdening the persistent storage with
-transient data.  Third, data processing operations can be placed in the IOD.
+users.
+%These transactions are local to the IOD layer until persisted to the
+%DAOS layer eliminating the need for burdening the persistent storage with
+%transient data.
+Third, data processing operations can be placed in the IOD.
 These operations are intended to offer functionality like data rearrangement
@@
 
-While these ideas are not necessarily new, they are new twists on best of class
-efforts for these technologies. For example, offloading the collective
-two-phase data sieving from the compute nodes to reorganize data has proven
-effective at reducing the total time for writing data due to fewer participants
-involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
-Beyond these broad items, there are many important details some of which are
-examined in more detail below.
+%While these ideas are not necessarily new, they are new twists on best of class
+%efforts for these technologies. For example, offloading the collective
+%two-phase data sieving from the compute nodes to reorganize data has proven
+%effective at reducing the total time for writing data due to fewer participants
+%involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
+%Beyond these broad items, there are many important details some of which are
+%examined in more detail below.
 
@@
 
-\subsubsection{Container}
-As mentioned above, the concept of a container is similar to that of a file in
-a traditional file system. However, rather than being in a directory structure,
-each container essentially is stored in a hash-space allowing direct access to
+\textbf{Container}
+%As mentioned above, the concept of
+A container is similar to that of a file in a traditional file system. However,
+rather than being in a directory structure, each container essentially is
+stored in a hash-space allowing direct access to
 any container without regard to the current organizational context of the
-file system. For example, there is no need to navigate a directory hierarchy
-to name a particular container.
-
-Functionally, a container plays the same role as a file in that it holds a
-collection of presumably related data intended to be accessed and manipulated
-as a unit. Since this is extended from HDF5 files, the container could also be
+file system.
+%For example, there is no need to navigate a directory hierarchy
+%to name a particular container.
+%
+%Functionally, a container plays the same role as a file in that it holds a
+%collection of presumably related data intended to be accessed and manipulated
+%as a unit.
+Since this is extended from HDF5 files, the container could also be
 viewed as a directory tree of objects where each directory entry specifies
@@
 
-\subsubsection{Key-Value Store}
+\textbf{Key-Value Store}
 This is the base type for the container. Since the container represents
@@
 
-\subsubsection{Multi-Dimensional Array}
+\textbf{Multi-Dimensional Array}
 By treating arrays as a special case separate from blobs, additional
@@
 
-\subsubsection{Blob}
+\textbf{Blob}
 All other data is stored as a stream-of-bytes without regard to the actual
@@
 
-\subsection{Multi-Dimensional Array Data Distribution}
+\subsection{Multi-Dimensional Arrays}
 
@@
 is common is a Fortran code (column-major) writes data for a C code (row-major)
-to analyze.  The IOD API supports the following sharding strategies:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{contiguous}. fixed chunking, distributed in a round-robin
+to analyze.  The IOD API supports the following sharding strategies.
+%
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%\item
+  \textbf{(1)contiguous}. fixed chunking, distributed in a round-robin
   fashion across the IO nodes.
-\item
-  \textbf{chunked}. same as above but with irregular (sparse) chunking.
-\item
-  \textbf{user-defined}. either contiguous or chunked, but user specifies
+%\item
+  \textbf{(2)chunked}. same as above but with irregular (sparse) chunking.
+%\item
+  \textbf{(3)user-defined}. either contiguous or chunked, but user specifies
   where to place each individual shard.
-\end{itemize}
-
-It is possible to request the transformation of an object's physical
-layout to other formats, having multiple copies of the same objects in
-multiple formats if desired. Also, the user can pre-fetch objects from
-the storage cluster into the IO nodes or read them directly from the
-storage cluster. At the semantic level (HDF5), indices can be created
-for datasets resulting in being able to read through an index instead of
-directly from the base array.
-
-All of these distinct alternatives result in having many different ways for
-executing the same analysis task.  In the subsequent discussions, we consider
-only data-movement optimization, i.e., sending the analysis code as close as
-possible to the data. In practice, this means we focus on identifying sharding
-of datasets and execute code accordingly over the appropriate shards.
+%\end{itemize}
+
+%It is possible to request the transformation of an object's physical
+%layout to other formats, having multiple copies of the same objects in
+%multiple formats if desired. Also, the user can pre-fetch objects from
+%the storage cluster into the IO nodes or read them directly from the
+%storage cluster. At the semantic level (HDF5), indices can be created
+%for datasets resulting in being able to read through an index instead of
+%directly from the base array.
+%
+%All of these distinct alternatives result in having many different ways for
+%executing the same analysis task.  In the subsequent discussions, we consider
+%only data-movement optimization, i.e., sending the analysis code as close as
+%possible to the data. In practice, this means we focus on identifying sharding
+%of datasets and execute code accordingly over the appropriate shards.
 
@@
 anticipate the execution of analytical tasks.
-
-\begin{figure}[htbp]
-\centering
-\vspace{-0.10in}
-\includegraphics[width=\columnwidth]{images/exa-arch.png}
-\vspace{-0.20in}
-\caption{Extreme Scale Architecture}
-\label{fig:exa-arch}
-\vspace{-0.10in}
-\end{figure}
-
-A common configuration for this type of deployment is shown in
-Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
-compute nodes (CNs) through the same fast fabric (e.g., InfiniBand) while the
-connection to the external storage cluster is through a secondary, slower
-channel (e.g., 10Gb Ethernet). By providing additional storage on the IO nodes,
-such as SSDs, these nodes are capable of better regulating the IO pressure on
-the underlying storage array better than simple forwarding gateways. For this
+%
+%\begin{figure}[htbp]
+%\centering
+%\vspace{-0.10in}
+%\includegraphics[width=\columnwidth]{images/exa-arch.png}
+%\vspace{-0.20in}
+%\caption{Extreme Scale Architecture}
+%\label{fig:exa-arch}
+%\vspace{-0.10in}
+%\end{figure}
+%
+%A common configuration for this type of deployment is shown in
+%Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
+%compute nodes (CNs) through the same fast fabric (e.g., InfiniBand) while the
+%connection to the external storage cluster is through a secondary, slower
+%channel (e.g., 10Gb Ethernet). By providing additional storage on the IO nodes,
+%such as SSDs, these nodes are capable of better regulating the IO pressure on
+%the underlying storage array better than simple forwarding gateways.
+For this
 project, using something like SSDs on the IO nodes is termed a {\em Burst
@@
 available for the applications.
-
-Future work will generalize this model to potentially support dynamic IO node
-allocation and examine the possibility of oversubscription. It will be strictly
-necessary to consider shared IO nodes for cases where the number of deployed
-applications exceeds the number of IO nodes.  This first phase focuses on
+%
+%Future work will generalize this model to potentially support dynamic IO node
+%allocation and examine the possibility of oversubscription. It will be strictly
+%necessary to consider shared IO nodes for cases where the number of deployed
+%applications exceeds the number of IO nodes.
+This first phase focuses on
 extreme scale application runs that use the vast majority of a platform rather
@@
 limitations of this design are acknowledged and the intent is to ultimately
-spread the IOD layer from the IO nodes into the compute area as well.  This is
-intended both to help address the limitations of the IO bandwidth and compute
-capability of these few nodes for data processing, but also to take advantage
-of new layers in the storage hierarchy. By incorporating NVRAM into compute
-nodes, new options for buffering data prior to being moved to centralized
-storage become available and addresses potential concerns about SSD
-performance. For example, including a small amount of Phase Change memory into
-many or most compute nodes offers a way to move data outside of both the
-compute and IO path for data and communication intensive operations. Other
-projects~\cite{zheng:2010:predata} have shown this will have value, but the
-cost will have to be considered as part of the overall platform budget. This
-lessens the impact of some operators while offering additional options for
-places to store data.
+spread the IOD layer from the IO nodes into the compute area as well.
+%This is
+%intended both to help address the limitations of the IO bandwidth and compute
+%capability of these few nodes for data processing, but also to take advantage
+%of new layers in the storage hierarchy. By incorporating NVRAM into compute
+%nodes, new options for buffering data prior to being moved to centralized
+%storage become available and addresses potential concerns about SSD
+%performance. For example, including a small amount of Phase Change memory into
+%many or most compute nodes offers a way to move data outside of both the
+%compute and IO path for data and communication intensive operations. Other
+%projects~\cite{zheng:2010:predata} have shown this will have value, but the
+%cost will have to be considered as part of the overall platform budget. This
+%lessens the impact of some operators while offering additional options for
+%places to store data.
 
@@
 local data within the IOD layer rather than against the globally shared DAOS
-layer. With the additional desire to support using compute node resources for
-these operations, serious work will be required to make a fully functional
-end-to-end IOD layer implementation for a production system.
+layer.
+%With the additional desire to support using compute node resources for
+%these operations, serious work will be required to make a fully functional
+%end-to-end IOD layer implementation for a production system.
 
@@
 by the VOSD layer.
-
 This is the layer where the container/object model is translated into the
@@
 
-While transactions at the HDF5 and IOD layer use the same term, at the DAOS
-layer the terminology shifts. Instead of transactions, the term {\em epochs}
-is used instead. Rather than attempting to introduce confusion, this is intended
-to help clarify how these concepts are used at different layers of the FFSIO
-stack. In the HDF5 and IOD layer, every operation has a transaction that may
-or may not ultimately be persisted to the DAOS layer. When a transaction is
-persisted to DAOS, it is termed an epoch to reflect that this is a persistent
-version of the container. For simplicity the epoch ID is the same as the
-transaction ID that was persisted.
-
-The process of persisting a transaction to an epoch involves a process called
-``flattening''. Since this stack uses a copy-on-write approach to reduce the
-space requirement for new versions of existing files, when a transaction is
-persisted to a DAOS epoch, all of the changes between the last epoch and the
-current epoch must be combined into a single entry. The process of combining
-the changes from a series of transactions into a single update for an epoch
-is termed flattening.
-
-The current implementation has the DAOS layer map the container/object data
-model onto a directory/file data model used for most existing file systems.
-Should a fully object-based file system be deployed at the VOSD layer, this
-mapping would be unnecessary. The current projections suggest that a standard
-POSIX-like file system will likely be used at the lowest level on each storage
-device requiring the mapping at some level. To perform this mapping, DAOS
-considers the following.
-
-Each container is represented by a directory on some storage device containing
-symbolic links to all of the shards it contains and maintains the epoch ID.
-In particular the Highest Committed Epoch (HCE) is an important concept for
-quickly identifying which version of a shard to retrieve and to block writes to
-older epochs since those have been committed.
+%While transactions at the HDF5 and IOD layer use the same term, at the DAOS
+%layer the terminology shifts. Instead of transactions, the term {\em epochs}
+%is used instead. Rather than attempting to introduce confusion, this is intended
+%to help clarify how these concepts are used at different layers of the FFSIO
+%stack. In the HDF5 and IOD layer, every operation has a transaction that may
+%or may not ultimately be persisted to the DAOS layer. When a transaction is
+%persisted to DAOS, it is termed an epoch to reflect that this is a persistent
+%version of the container. For simplicity the epoch ID is the same as the
+%transaction ID that was persisted.
+
+%The process of persisting a transaction to an epoch involves a process called
+%``flattening''. Since this stack uses a copy-on-write approach to reduce the
+%space requirement for new versions of existing files, when a transaction is
+%persisted to a DAOS epoch, all of the changes between the last epoch and the
+%current epoch must be combined into a single entry. The process of combining
+%the changes from a series of transactions into a single update for an epoch
+%is termed flattening.
+
+%The current implementation has the DAOS layer map the container/object data
+%model onto a directory/file data model used for most existing file systems.
+%Should a fully object-based file system be deployed at the VOSD layer, this
+%mapping would be unnecessary. The current projections suggest that a standard
+%POSIX-like file system will likely be used at the lowest level on each storage
+%device requiring the mapping at some level. To perform this mapping, DAOS
+%considers the following.
+
+%Each container is represented by a directory on some storage device containing
+%symbolic links to all of the shards it contains and maintains the epoch ID.
+%In particular the Highest Committed Epoch (HCE) is an important concept for
+%quickly identifying which version of a shard to retrieve and to block writes to
+%older epochs since those have been committed.
 
@@
 
-At a more detailed level, the design for VOSD is an increment beyond the
-current Lustre Object Storage Device design to incorporate the idea of shards
-and the versioning aspects of transactions/epochs. For every DAOS shard, the
-VOSD has information for storing and accessing the currently committed version,
-the Highest Committed Epoch, as well as a staging dataset representing the
-next version of the object being stored. Both of these are combined in a shard
-root.
-
-For data integrity, an intent log is maintained as part of the underlying file
-system enabling fault recovery.
-
-Beyond the functionality to incorporate and expose the copy-on-write nature
-of the underlying file system and the semantics for storing and processing
-shards and their associated epochs, this is largely an evolution of the
-existing Lustre OSD layer.
+%At a more detailed level, the design for VOSD is an increment beyond the
+%current Lustre Object Storage Device design to incorporate the idea of shards
+%and the versioning aspects of transactions/epochs. For every DAOS shard, the
+%VOSD has information for storing and accessing the currently committed version,
+%the Highest Committed Epoch, as well as a staging dataset representing the
+%next version of the object being stored. Both of these are combined in a shard
+%root.
+%
+%For data integrity, an intent log is maintained as part of the underlying file
+%system enabling fault recovery.
+%
+%Beyond the functionality to incorporate and expose the copy-on-write nature
+%of the underlying file system and the semantics for storing and processing
+%shards and their associated epochs, this is largely an evolution of the
+%existing Lustre OSD layer.
 
@@
 In the subsections that follow, we examine transactions and epochs, metadata
-management, function and analysis shipping, and the arbitrary connected graphs
-support.
+management, and function and analysis shipping.
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+\vspace{-0.05in}
+\end{figure*}
 
@@
 \subsubsection{Transactions}
-To understand how transactions are used in the IOD layer, some terminology and
-concepts must be explained first. At the coarsest grain level is a container.
-Each container provides the single access context through which to access a
-collection of objects. Transactions are the way that a series of modifications
-to the objects within a container are treated atomically. Conceptually,
-containers corresponds to a something akin to an HDF5 file in a traditional
-file system. The objects in each container represent different data within a
-file.  The three initially defined object types are key-value stores,
-multi-dimensional arrays, and blobs.  The easiest way to understand these types
-is to evaluate these from the perspective of an HDF5 file, the initial user
-interface layer. The key-value store represents a collection of attributes or
-groups. The array represents a potentially multi-dimensional array.  The blob
-represents a byte stream of arbitrary contents.  The fundamental difference
-between an array and a blob is that the array has metadata specifying the
-dimension(s). At the physical layer within the IO nodes, all of these objects
-may be striped across multiple IO nodes.  Given this context, the transactions
-come in two forms.
-
-First is a single leader transaction where the IOD manages based on calls from
+%To understand how transactions are used in the IOD layer, some terminology and
+%concepts must be explained first. At the coarsest grain level is a container.
+%Each container provides the single access context through which to access a
+%collection of objects.
+Transactions are the way that a series of modifications
+to the objects within a container are treated atomically.
+%Conceptually,
+%containers corresponds to a something akin to an HDF5 file in a traditional
+%file system. The objects in each container represent different data within a
+%file.  The three initially defined object types are key-value stores,
+%multi-dimensional arrays, and blobs.  The easiest way to understand these types
+%is to evaluate these from the perspective of an HDF5 file, the initial user
+%interface layer. The key-value store represents a collection of attributes or
+%groups. The array represents a potentially multi-dimensional array.  The blob
+%represents a byte stream of arbitrary contents.  The fundamental difference
+%between an array and a blob is that the array has metadata specifying the
+%dimension(s). At the physical layer within the IO nodes, all of these objects
+%may be striped across multiple IO nodes.  Given this context, the
+Transactions are either single or multi-leader.
+
+\textbf{Single leader} transactions let IOD manage based on calls from
 a single client. The underlying assumption is that the client side will manage
@@
 reporting to the IOD how to evolve the transaction state. 
-
-The second form is called multi-leader and has the IOD layer manage the
-transactions. In this case, when the transaction is created, a count of clients
+\textbf{Multi-leader} has the IOD layer manage the
+transactions for all clients. In this case, when the transaction is created, a count of clients
 is provided to the IOD layer. As clients commit their changes to the container,
 the reference count is reduced. Once the count reaches 0, the transaction is
-automatically committed.
+automatically committed. Aggregation into a smaller set of leaders is also
+possible.
 
@@
 
-There are two steps for beginning a transaction on a container. The first step
-is for one or more process to open the container. This handle can be shared
-eliminating the need for every participating process to hit the IOD layer to
-open the file. The second step is a call to determine how many leaders will
-participate in the transaction. In the single leader case, there is no
-aggregation of success/fail statuses to determine the final transaction state.
-Instead, it is assumed that the client will fully manage the transaction. In
-the multi-leader model, some subset from 2 to $n$ where $n$ is the count of all
-processes, declare themselves a leader for this container operation to the IOD
-layer. Any number of processes can participate in modifying container without
-regard to whether or not they are a leader. Once each leader has finished, with
-the assumption that any clients they may be responsible for are finished as
-well, the IOD layer aggregates those responses to either commit or abort the
-transaction.
-
-Ultimately, with the passive detection of faults for transaction leaders, the
-transaction mechanism can work very well. A mostly unstated restriction that is
+%There are two steps for beginning a transaction on a container. The first step
+%is for one or more process to open the container. This handle can be shared
+%eliminating the need for every participating process to hit the IOD layer to
+%open the file. The second step is a call to determine how many leaders will
+%participate in the transaction. In the single leader case, there is no
+%aggregation of success/fail statuses to determine the final transaction state.
+%Instead, it is assumed that the client will fully manage the transaction. In
+%the multi-leader model, some subset from 2 to $n$ where $n$ is the count of all
+%processes, declare themselves a leader for this container operation to the IOD
+%layer. Any number of processes can participate in modifying container without
+%regard to whether or not they are a leader. Once each leader has finished, with
+%the assumption that any clients they may be responsible for are finished as
+%well, the IOD layer aggregates those responses to either commit or abort the
+%transaction.
+
+%Ultimately, with the passive detection of faults for transaction leaders, the
+%transaction mechanism can work very well.
+A mostly unstated restriction that is
 being relaxed is that every sequential transaction on a container is considered
@@
 subsequent five succeed, when the delayed process finally fails, all six
-transactions are rolled back. The thought of using this mechanism to store
-subsequent checkpoint outputs in the same container to both save space, but not
-care if one fails, cannot work in the current form. This has been acknowledged
-and is being relaxed requiring a new parameter to the creation of a transaction
-determining if it will be dependent or not.
+transactions are rolled back.
+%The thought of using this mechanism to store
+%subsequent checkpoint outputs in the same container to both save space, but not
+%care if one fails, cannot work in the current form. This has been acknowledged
+%and is being relaxed requiring a new parameter to the creation of a transaction
+%determining if it will be dependent or not.
 
@@
 \subsection{Metadata Management}
-Metadata management has been a perennial challenge for parallel storage
-systems.  Eliminating metadata management as a special case and instead
-treating it just as data is a central design goal of the Fast Forward project.
-This is a hybrid approach to metadata management that is halfway between
-providing no inherent metadata support and having a fully integrated, but
-separate metadata management system.
-
-Eliminating metadata as a core component of a file system is not new. It has
-been explored as part of the Light Weight File Systems
-project~\cite{oldfield:lwfs}. In LWFS, the metadata service is explicitly
-limited to a user task with the storage layer limited to data
-storage/retrieval, authorization, and authentication. This approach proved
-workable. Using this hybrid approach is less common~\cite{weil:2006:ceph} and
-introduces other issues.
+%Metadata management has been a perennial challenge for parallel storage
+%%systems.  Eliminating metadata management as a special case and instead
+%treating it just as data is a central design goal of the Fast Forward project.
+%FFSIO uses a hybrid approach to metadata management that is halfway between
+%providing no inherent metadata support and having a fully integrated, but
+%separate metadata management system.
+
+%Eliminating metadata as a core component of a file system is not new. It has
+%been explored as part of the Light Weight File Systems
+%project~\cite{oldfield:lwfs}. In LWFS, the metadata service is explicitly
+%limited to a user task with the storage layer limited to data
+%storage/retrieval, authorization, and authentication. This approach proved
+%workable. Using this hybrid approach is less common~\cite{weil:2006:ceph} and
+%introduces other issues.
 
@@
 appropriate. Unfortunately, this still couples the metadata to a single object
-that must serialize access. If the metadata, including information about
-striping and other data layout operations were separated completely from the
-data path, more scalable throughput could be achieved. The real challenge of
+that must serialize access.
+%If the metadata, including information about
+%striping and other data layout operations were separated completely from the
+%data path, more scalable throughput could be achieved.
+The real challenge of
 this is introduced by the IOD, DAOS, and VOSD layers collectively. Each of
@@
 
-\subsection{Arbitrarily Connected Graphs}
-\label{sec:acg}
-
-What people popularly consider Big Data applications fall into two broad
-categories. First, data processing tasks that can fit into the MapReduce model
-where data is tagged and sorted to discover relationships. These sorts of
-applications only require scale out rather than scale up. Scaling out requires
-replicas to process data simultaneously, but do not need to coordinate for that
-data processing. Scaling up, what scientific simulations do, requires sometimes
-serious coordination between the processes for any of them to succeed. In the
-middle are graph applications that, with some replicated data, can be made to
-fit reasonably well into the MapReduce model. The challenge is having access to
-the edge and vertex lists effectively to build partitions for independent
-processing. GraphBuilder~\cite{Jain:2013:GraphBuilder} is a tool to generate
-effective graph partitions reducing the load for using MapReduce to process
-graph data sets. GraphLab~\cite{Low:2012:GraphLab} offers a way to process
-these graphs efficiently for parallel platforms using a minimum of
-communication. Using these tools as motivators, changes to the HDF5 interface
-and the underlying storage infrastructure is proposed.  The following
-illustrates the architecture of both frameworks:
-
-\begin{figure}[htbp]
-\centering
-\vspace{-0.10in}
-\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
-\vspace{-0.30in}
-\caption{GraphLab and GraphBuilder stacks}
-\label{fig:graphlab-graphbuilder}
-\vspace{-0.10in}
-\end{figure}
-
-In order to make both of these tools work on top of the extreme scale stack,
-they both have to be modified. After these modifications are implemented,
-GraphBuilder will be able to write the partitioned graph in the newly proposed
-HDF5 files which will thus be stored in the IOD nodes (or IONs) in a
-parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
-the library to perform at high speeds by benefiting from the new
-features, such as the function shipping. In general both frameworks will be
-modified so that calls to HDFS-based formats are replaced by the proposed HDF5
-equivalents. This is referred to as the HDF Adaptation Layer or HAL and will
-provide, from the GraphBuilder/GraphLab point of view:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  capability for storing the newly proposed HDF5 format
-\item
-  association of network information to vertices/edges
-\item
-  shipping computation to the IONs
-\item
-  asynchronous vertex updates
-\item
-  efficient data sharing among CNs
-\item
-  computation over versioned datasets
-\end{itemize}
-
-The initial phase of this project has determined the necessary changes in the
-HDF5 format to support these features. These identified features will be
-proven during phase two with a demonstration of GraphBuilder and GraphLab.
+%\subsection{Arbitrarily Connected Graphs}
+%\label{sec:acg}
+%
+%What people popularly consider Big Data applications fall into two broad
+%categories. First, data processing tasks that can fit into the MapReduce model
+%where data is tagged and sorted to discover relationships. These sorts of
+%applications only require scale out rather than scale up. Scaling out requires
+%replicas to process data simultaneously, but do not need to coordinate for that
+%data processing. Scaling up, what scientific simulations do, requires sometimes
+%serious coordination between the processes for any of them to succeed. In the
+%middle are graph applications that, with some replicated data, can be made to
+%fit reasonably well into the MapReduce model. The challenge is having access to
+%the edge and vertex lists effectively to build partitions for independent
+%processing. GraphBuilder~\cite{Jain:2013:GraphBuilder} is a tool to generate
+%effective graph partitions reducing the load for using MapReduce to process
+%graph data sets. GraphLab~\cite{Low:2012:GraphLab} offers a way to process
+%these graphs efficiently for parallel platforms using a minimum of
+%communication. Using these tools as motivators, changes to the HDF5 interface
+%and the underlying storage infrastructure is proposed.  The following
+%illustrates the architecture of both frameworks:
+%
+%\begin{figure}[htbp]
+%\centering
+%\vspace{-0.10in}
+%\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
+%\vspace{-0.30in}
+%\caption{GraphLab and GraphBuilder stacks}
+%\label{fig:graphlab-graphbuilder}
+%\vspace{-0.10in}
+%\end{figure}
+%
+%In order to make both of these tools work on top of the extreme scale stack,
+%they both have to be modified. After these modifications are implemented,
+%GraphBuilder will be able to write the partitioned graph in the newly proposed
+%HDF5 files which will thus be stored in the IOD nodes (or IONs) in a
+%parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
+%the library to perform at high speeds by benefiting from the new
+%features, such as the function shipping. In general both frameworks will be
+%modified so that calls to HDFS-based formats are replaced by the proposed HDF5
+%equivalents. This is referred to as the HDF Adaptation Layer or HAL and will
+%provide, from the GraphBuilder/GraphLab point of view:
+%
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%\item
+%  capability for storing the newly proposed HDF5 format
+%\item
+%  association of network information to vertices/edges
+%\item
+%  shipping computation to the IONs
+%\item
+%  asynchronous vertex updates
+%\item
+%  efficient data sharing among CNs
+%\item
+%  computation over versioned datasets
+%\end{itemize}
+%
+%The initial phase of this project has determined the necessary changes in the
+%HDF5 format to support these features. These identified features will be
+%proven during phase two with a demonstration of GraphBuilder and GraphLab.
 
@@
 
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
-\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Number of Hosts}
-\label{fig:eval-hosts}
-\vspace{-0.05in}
-\end{figure*}
-
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
-\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Data Sizes}
-\label{fig:eval-size}
-\vspace{-0.05in}
-\end{figure*}
-
-This stack has an early prototype implementation intended to test concepts
-rather than performance and scalability. It has focused on examining the
-interaction of the different APIs for each layer to flesh out any detailed
-requirements or concerns that may have been missed in the conceptualization of
-this IO stack. To demonstrate the viability of the IO stack described in this
-paper, we show some very early performance results from the untuned prototype.
+%This stack has an early prototype implementation intended to test concepts
+%rather than performance and scalability. It has focused on examining the
+%interaction of the different APIs for each layer to flesh out any detailed
+%requirements or concerns that may have been missed in the conceptualization of
+%this IO stack. To demonstrate the viability of the IO stack described in this
+%paper, we show some very early performance results from the untuned prototype.
 

### 2015-09-17 paper.tex
moved Carlos and Ivo to Acknowledgements

@@
 \author{
-\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2], Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
+%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2], Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
+\alignauthor {Jay Lofstead\footnotemark[1], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
 \footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
-\footnotemark[2]~~University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu\\
+%\footnotemark[2]~~University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu\\
 \footnotemark[3]~~HDF Group koziol@hdfgroup.org\\
@@
 \label{fig:eval-size}
-\vspace{-0.05in}
+\vspace{-0.2in}
 \end{figure*}
@@
 \includegraphics[scale=0.35]{logos/nnsa_logo}
+The authors would like to thank Carlos Maltzahn and Ivo Jimenez for both
+facilitating the connections and motivating the detailed analysis that led to
+this work.
+
 Sandia National Laboratories is a multi-program laboratory managed and operated

### 2016-07-18 paper.tex
added reviews and "plagarism" concerns (fixed via refs). Need to edit a bit more still.

--- 2014-05-13 paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\pdfpagewidth=8.5in
+\documentclass[conference]{IEEEtran} \pdfpagewidth=8.5in
 \pdfpageheight=11in
@@
 \usepackage{listings}
+\usepackage{hyperref}
 \lstset {% general command to set parameter(s)
@@
 
-%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}
-
 \newcommand{\DDT}{D\textsuperscript{2}T~}
@@
 
-\hyphenation{sub-trans-ac-tion}
-
 \addtolength{\parskip}{-0.08in}
@@
 
-%\conferenceinfo{Cluster'14,} {September 17, 2014, Madrid, Spain}
-%\CopyrightYear{2014}
-%\crdata{978-1-4503-1103-8/11/11}
-%\clubpenalty=10000
-%\widowpenalty = 10000
-
-\title{An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications}
+%\title{Fast Forwarding Storage for Extreme Scale Platforms and Big Data Applications}
+\title{Lessons Learned from the Fast Forward Storage and I/O Project}
 
 \author{
+%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
 \IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Lawrence Berkeley National Laboratory koziol@lbl.gov}
+\IEEEauthorblockA{\IEEEauthorrefmark{4}Segate Government Solutions John.Bent@seagategov.com}
 \IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
@@
 \begin{abstract}
+%The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
+%project is going to have significant impact on storage systems design within
+%and beyond the HPC community. With phase two of the project starting, it is an
+%excellent opportunity to explore the complete design and how it will address
+%the needs of extreme scale platforms.  This paper examines the overall design.
+%Each layer of the proposed stack is discussed in some detail along with
+%cross-cutting topics that impact multiple layers, such as transactions and
+%metadata management. We also describe the proposed features to better support
+%some classes of big data applications.
+%
+%This paper not only provides a timely summary of important aspects of the
+%design specifications but also captures the underlying reasoning that is not
+%available elsewhere. With this information, we encourage the broader community
+%to understand the design, intent, and future directions to foster discussion
+%guiding phase two and the ultimate production storage stack based on this work.
+%An initial performance evaluation of the early prototype implementation is also
+%provided to validate the presented design.
+
 The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
 project is going to have significant impact on storage systems design within
-and beyond the HPC community. With phase 1 of the project complete, it is an
+and beyond the HPC community. With phase two of the project starting, it is an
 excellent opportunity to explore the complete design and how it will address
-the needs of extreme scale platforms.  This paper examines the overall design.
-Each layer of the proposed stack is discussed in some detail along with
-cross-cutting topics that impact multiple layers, such as transactions and
-metadata management. We also describe the proposed features to better support
-some classes of big data applications.
+the needs of extreme scale platforms.  This paper examines each layer of the
+proposed stack in some detail along with cross-cutting topics, such as
+transactions and metadata management.
 
@@
 design specifications but also captures the underlying reasoning that is not
-available elsewhere. With this information, we encourage the broader community
-to understand the design, intent, and future directions to foster discussion
-guiding phase 2 and the ultimate production storage stack based on this work.
-An initial performance evaluation of the early prototype implementation is also
-provided to validate the presented design.
+available elsewhere. We encourage the broader community to understand the
+design, intent, and future directions to foster discussion guiding phase two
+and the ultimate production storage stack based on this work.  An initial
+performance evaluation of the early prototype implementation is also provided
+to validate the presented design.
 
@@
 and performance to adequately serve extreme scale science platform
-requirements.  Adding to the problem complexity is the variety of Big Data
-problems users want to address using these platforms. Trying to address both of
-these domains is a daunting problem.
+requirements.  While new hardware, such as non-volatile memory will help, we
+still need a new software stack to incorporate this new hardware as well as
+address the extreme parallelism and performance requirements demanded by
+exascale applications. Adding to the problem complexity is the variety of Big
+Data problems users want to address using these platforms. Unlike the
+centralized storage arrays favored for HPC platforms, big data analytics
+systems have grown up using storage distributed on all of the nodes driving a
+very different software architecture. With post-exascale platforms required to
+address both workloads, a new storage stack is required.
 
@@
 set~\cite{fastforward:2014:docs} for a future IO stack to address the
-identified challenges. The first phase recently completed with a second phase
-currently getting underway as of early 2014. The core focus of the first phase
-was basic functionality and design. While an idealized potential system would
-be the perfect target architecture, the reality of budgets has tempered many of
-the decisions. For example, extensive availability of NVRAN or SSDs on all of
-the compute nodes is currently not economically feasible limiting some of the
+identified challenges. The first phase completed in 2014 with a second phase
+currently getting underway. The core focus of the first phase was basic
+functionality and design. While an idealized potential system would be the
+perfect target architecture, the reality of budgets has tempered many of the
+decisions. For example, extensive availability of NVRAM or SSDs on all of the
+compute nodes is currently not economically feasible limiting some of the
 potential design choices.  With this in mind, the second phase will refine this
@@
 \begin{figure}[htbp]
-\vspace{-0.10in}
+%\vspace{-0.10in}
 \centering
 \includegraphics[width=\columnwidth]{images/arch-mapping}
-\vspace{-0.15in}
+%\vspace{-0.15in}
 \caption{Target Architecture and Component Mapping}
 \label{fig:arch-mapping}
-\vspace{-0.15in}
+%\vspace{-0.15in}
 \end{figure}
@@
 incorporates five layers, some of which have potentially optional components.
-The top layer is generally a high level IO library, such as the demonstration
-HDF5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
-access to the storage stack through such an API to manage the complexity of
-working with the lower layers and to enable advanced functionality that would
-require more direct user intervention. This layer incorporates the necessary
-features for ACGs from a end-user's perspective.
+The top layer is comprised of generally a high level IO library, such as the
+demonstration HDF5 library~\cite{hdf5} and a more complex API for accessing the
+lower level components. This layer is in dark blue. Becuase the system supports
+more complex architectures and supports richer functionality, hiding this
+complexity behind a user-friendly API is the intent. It is possible to 
+access the storage stack through the more complex API, but the additional
+requirements beyond standard POSIX calls will prompt most users to use an IO
+library.  This layer incorporates the necessary features for ACGs from a
+end-user's perspective.
 
@@
 passive data staging processes demonstrated
-previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
-have considerable functionality. The IO dispatcher (IOD) serves as the primary
-storage interface for the IO stack (in green) and offers features like Burst
-Buffers to insulate the persistent storage array from bursty IO workloads.
-Ideally, the IOD layer's functionality can be optional based on available
-hardware and compute power provided on the IO Nodes (IONs). Much of the
-functionality offered at this layer would shift either up or down the stack as
-discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
-layer serves as the persistent storage interface and translation layer between
-the user-visible object model and the requirements of the underlying storage
-infrastructure. It is intended to be the traditional file system-like
+previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. One special function
+of note for this layer is that it is where function shipping will be deployed.
+This is discussed in Section~\ref{sec:end-user}. The next two layers have
+considerable functionality.
+
+The IO Dispatcher (IOD) serves as the primary storage interface for the IO
+stack (in green) and offers features like Burst Buffers to insulate the
+persistent storage array from bursty IO workloads.  Ideally, the IOD layer's
+functionality can be optional based on available hardware and compute power
+provided on the IO Nodes (IONs). Transactions are handled primarily at this
+layer. Much of the functionality offered at this layer would shift either up or
+down the stack as discussed in detail below.
+
+The Distributed Asynchronous Object Storage (DAOS) layer serves as the
+persistent storage interface and translation layer between the user-visible
+object model and the requirements of the underlying storage infrastructure. 
+Transactions work a bit differently at this layer and are called epochs to
+distinguish them. DAOS is intended to be the traditional file system-like
 foundation on which everything else is built with no dependence on any
@@
 properly.  Instead, the DAOS layer can handle all of the IO operations from the
-user API layer, albeit with the potential performance penalty of manipulating
-the shared, persistent storage array. At the bottom is the Versioning Object
-Storage Device (VOSD) (in purple).  It serves as the interface for storing
-objects of all types efficiently for each storage device in the parallel
-storage array. Think of this layer as the physical disk interface layer. In
-terms of Lustre, this would replace the API on individual storage devices with
-an interface friendlier to the containers of objects and transactions/epochs
-concepts used in the higher layers.
+user API layer, albeit with the potential performance and usability penalty of
+manipulating the shared, persistent storage array directly.
+
+At the bottom is the Versioning Object Storage Device (VOSD) (in purple).  It
+serves as the interface for storing objects of all types efficiently for each
+storage device in the parallel storage array. Think of this layer as the
+physical disk interface layer. In terms of Lustre, this would replace the API
+on individual storage devices with an interface friendlier to the containers of
+objects and transactions/epochs concepts used in the higher layers.
 
@@
 information represents information that may or may not have been written down,
-but is the intent of ultimate product.  These ideas are presented to give a
-fuller picture of where the project is going rather than dwelling on any
+but is the intent of ultimate system.  This information was gathered through
+interviews with the core FFSIO team members. These ideas are presented to give
+a fuller picture of where the project is going rather than dwelling on any
 limitations of the published documents. This is most important to illustrate
 how different concepts will work across layers since that information is spread
-across multiple documents and may lack a cohesive overall view.
-
-The rest of the paper is organized as follows. A brief overview of related work
-is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
+across multiple documents and may lack a cohesive overall view. Previously,
+a poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
+consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
+been presented.
+
+The rest of the paper is organized as follows. An overview of related work is
+presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
 discusses the programmatic interface end users will see when interacting with
@@
 Data'' applications with different characteristic IO patterns have somewhat
-complicated the picture. Extreme scale machines will be expected to handle both
-the traditional simulation-related workloads as well as applications more
+complicated the picture. Vendors are shifting products to address the far
+larger market forcing HPC systems to adapt to these different storage
+approaches. Extreme scale machines will be expected to handle both the
+traditional simulation-related workloads as well as applications more
 squarely in the Big Data arena. This will require some adjustments to the
@@
 
+Lustre~\cite{braam:lustre-arch} is the de facto standard on most major clusters
+offering scalable performance and fine-grained end-user and programmatic
+control over how data is placed in the storage system. The broad community
+support has led to a solid code base with sufficient optimizations to serve as
+the low-cost, proven solution. For each installation, system-wide settings that
+apply to all files on the file system are made. The end user, should they have
+different needs can reconfigure these characteristics on a file-by-file basis.
+This becomes an issue because the dominant file size is tiny. In many cases, it
+can be $<$ 4 KB.  To keep from slowing the overall system performance when
+creating and opening these files, most systems are configured to use a 1 MB
+stripe size and a stripe count of 4 meaning only 4 storage targets are used per
+file. This limits the default aggregate bandwidth to the combined speed of four
+storage targets. By reconfiguring on a file-by-file basis, this default can be
+overcome for large, parallel files achieving very high performance. The
+downside is that this setting must be done to take advantage of the full
+parallel file system performance.
+
 Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
-both a POSIX and object interface including features typically found in parallel
-file systems. Ceph's unique striping approach uses pseudo-random numbers with a
-known seed eliminating the need for the metadata service to track where each
-stripe in a parallel file is placed.
-
-PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
-such as a single process opening a file and sharing the handle. It has been
-commercialized in recent years as OrangeFS.
-
-Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
-clusters offering scalable performance and fine-grained end-user and
-programmatic control over how data is placed in the storage system.
-
-GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
+both a POSIX and object interface including features typically found in
+parallel file systems. Ceph's unique striping approach uses pseudo-random
+numbers with a known seed eliminating the need for the metadata service to
+track where each stripe in a parallel file is placed. Ceph's strengths are in
+providing good perforamnce and scalability with the ability to handle failures
+and deploying new storage adapting the system in a live environment. However,
+this failure handling advantage was shown~\cite{wang:2013:ceph} to limit peak
+performance more than other systems like Lustre.  More recently, some of these
+limitations have begun to be addressed by the Ceph team, but no new evaluations
+have been performed to determine if these changes close the gap sufficiently to
+address the extreme HPC performance needs.
+
+PVFS~\cite{carns:pvfs} was built understanding the scalibility bottlenecks
+Lustre suffers. For example, Lustre requires all processes opening a file to
+hit the metadata server to receive a proper file handle. PVFS reduces this load
+by allowing a single process to open a file and sharing the handle with other
+processes participating in the IO operation. There are also other optimizations
+that enhance file system performance. It has been commercialized in recent
+years as OrangeFS.
+
+GPFS~\cite{schmuck:gpfs} offers a highly scalable parallel file system with
+robust functionality to handle both parallel storage, recovery, and
+optimization. It only supports a hands-off approach for providing good
 performance for scaling parallel IO tasks and is used extensively by its owner,
-IBM.
-
-Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
-striping system that detects the need for additional stripes for performance
-and adjusts the file layout as necessary.
+IBM. Unfortunately, the stripe size is fixed introducing potentially false
+sharing, when two processes indpendently write to the same stripe, but without
+overlapping, causing potentially reduced performance. Beyond these sorts of
+fixed parameters, a wide variety of optimizations and features are available
+for additional licensing fees.
+
+Panasas~\cite{panasas:architecture} uses a fundamentally different approach to
+parallel file system performance. When parallel writers simultaneously write to
+a shared file, the system dynamically adapts the number of stripes to maintain
+high performance. This adaption is invisible to the user other than seeing that
+the system maintains high performance no matter what configuration the workload
+exhibits.
+
+This project learns from all of these parallel file system efforts and offers a
+scalable approach that can work well for everything from a small cluster to the
+largest exascale platforms. By understanding what works well and what the
+limitations are for each of the above systems as well as emerging hardware
+architectures, this project addresses the limitations while maintaining the
+advantages of the above systems.
 
@@
 treat the resulting collection of pieces as a single object.  Distributed file
-systems rely on a single client creating a file, but distributing the set of
-files across a wide array of storage devices. The other, popular distributed
-file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
-decades for enterprise file systems.  These other file systems are mainly of
-interest in the context of the ACG features of FFSIO and will be discussed more
-in Section~\ref{sec:acg}.
+systems rely on a single client creating a file typically on a single storage
+device. For performance, the file or object may be replicated.  The other,
+popular distributed file system of note is NFS~\cite{powlowski:1994:nfs3} that
+has been used for decades for enterprise file systems.  NFS is known to support
+a global namespace with data migrating towards users on access and pushed
+towards safer storage based on local platform characteristics. These other file
+systems are mainly of interest in the context of the ACG features of FFSIO and
+will be discussed more in Section~\ref{sec:acg}.
+
+The main alternative from scratch design for a file system is
+Sirocco~\cite{sirocco,curry:2016:sirocco}. Rather than continuing the striped
+design of existing parallel file systems, Sirocco is inspired by peer-to-peer
+and object-based systems and includes features like transactions to protect
+data modification process independence when writing to avoid coordination
+overhead.  The base assumptions are that storage is pervasive and volatile.
+Storage devices and locations may come and go randomly, reminiscent of the
+Google or Ceph assumptions of regularly failing hardware. When data is pushed
+into the system, initial resilience characteristics are guaranteed prior to
+returning control back to the user. Then, as system pressures dictate, data
+will either replicate as demanded by use and/or migrated towards long-term
+resilience requirements.  Unlike the FFSIO project, Sirocco assumes it is
+possible that data may be successfully stored in the system, but it is
+currently inaccessible because all copies are currently offline. There is also
+some potential difficulty in finding data since it will migrate around the
+system. To be fair, Sirocco intends to function as the storage layer for a
+higher level file system API making many of the awkward system features
+invisible to the end user. Sirocco is in the process of being released
+publicly.
 
@@
 interface for programmatically interacting with the FFSIO stack, the team used
-the HDF5 API and leveraged the Virtual Object Layer (VOL) for the initial
+the HDF5 API and leveraged its Virtual Object Layer (VOL) for the initial
 design and implementation demonstration. This also serves as a good test
@@
 \def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
+%\itemsep1pt\parskip0pt\parsep0pt
 
@@
 
-\item\label{item:fs}
+\item
   Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
@@
 \item
-  Analysis Shipping from CN to IONs or DAOS nodes. This is similar
-  to~\ref{item:fs} but instead of returning the result over the network, it
-  gets stored on the nodes and pointers to it are returned.
+  Analysis Shipping from compute nodes to IO Nodes or DAOS nodes. This is
+  similar to function shipping, but instead of returning the result over the
+  network, it is stored on the nodes and pointers to the data are returned.
 
@@
 creating associations between various information items. The main
-components of HDF5 are described below.
+conceptual components for data stored in HDF5 are described below.
 
@@
   could contain multiple other groups, datasets, datatypes or attributes within
-  it.
+  it. Groups are named and then accessed using a standard path notation
+  similar to Linux with a ``/'' separating each group name in the hierarchy
+  from the root to the nested group of interest.
 
@@
   \textbf{Dataset}: HDF5 datasets are objects representing actual data
-  or content. Datasets are arrays with potentially multiple dimensions. A
+  or content. Datasets are typically arrays with potentially multiple
+  dimensions. Other types, such as strings and scalars, are also possible. A
   dataset is characterized by a dataspace and a datatype. The dataspace
@@
   extent in each dimension. The datatype describes the type of its data
-  elements.
+  elements. By default, the entire data set is stored as a single chunk
+  reassembled from all processes. It is possible to use a uniform chunking
+  format where data is stored in fixed sized chunks instead.
 
@@
 \begin{figure}[htbp]
-\vspace{-0.10in}
+%\vspace{-0.10in}
 \centering
 \includegraphics[width=\columnwidth]{images/vol-arch.png}
-\vspace{-0.20in}
+%\vspace{-0.20in}
 \caption{Architectural view of the VOL abstraction mechanism}
 \label{fig:vol-arch}
-\vspace{-0.10in}
+%\vspace{-0.10in}
 \end{figure}
 
-The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
+The IOD VOL plugin serves as the bridge between HDF5 and the IOD Layer
 (Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
@@
 
+Since the IOD layer is optional by design, a second VOL plugin is required to
+access DAOS directly. This additional complexity was deemed acceptable for the
+flexibility it affords. Further, since all end user interactions are intended
+to be through an IO API, having two different plugins is a small amount of
+extra work to provide a single interface that would operate on widely different
+scale deployments (e.g., one too small to have the IOD layer and one at the
+other extreme with a large IOD layer interfacing with a large, shared DAOS
+layer).
+
 \subsection{HDF5 to FFSIO Mapping}
@@
 internal data types, such as datasets, these concepts must be mapped onto the
-proposed FFSIO data storage concepts. At a high level, think of the HDF5 types
-mapping to FFSIO types as follows: file $\rightarrow$ container, dataset $\rightarrow$
-array, and groups and attributes $\rightarrow$ key-value store.
-%This mapping is shown in Table~\ref{tab:mapping}
-
-%\begin{table}[ht]
+proposed FFSIO data storage concepts.  This mapping is shown in
+Table~\ref{tab:mapping}
+
+\begin{table}[ht]
 %    \vspace{-0.10in}
-%    \centering
-%    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
-%    \bigskip
+    \centering
+    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
+    \bigskip
 %    \vspace{-0.15in}
-%
-%    \begin{tabular}{|r|r|}
-%\hline
-%HDF5 & FFSIO\\
-%\hline
-%file & container \\
-%dataset & array \\
-%group & key-value store \\
-%attribute(s) & key-value store \\
-%\hline
-%    \end{tabular}
-%    \label{tab:mapping}
-%\end{table}
-%
+
+    \begin{tabular}{|r|r|}
+\hline
+HDF5 & FFSIO\\
+\hline
+file & container \\
+dataset & array \\
+group & key-value store \\
+attribute(s) & key-value store \\
+\hline
+    \end{tabular}
+    \label{tab:mapping}
+\end{table}
+
 In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.
@@
 The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
-the massive process count on the storage stack. A current trend mixing MPI with
-a threading library like OpenMP is addressing the same issue, but limited to
-handling the parallelism on a single node rather than multiple nodes. Projected
-extreme scale platforms will have far fewer storage stack end-points per
-compute process in which to receive requests and data. By reducing the number
-of simultaneous requests, delays can be reduced. This has been demonstrated for
+the massive process count fan in onto the storage stack. The current trend of
+using both MPI and a node-level threading library like OpenMP, CUDA, or OpenACC
+is addressing the same issue, but limited to handling the parallelism on a
+single node rather than multiple nodes. Projected extreme scale platforms will
+have far fewer storage stack end-points per compute process or even compute
+node in which to receive requests and data. By reducing the number of
+simultaneous requests, delays can be reduced. This has been demonstrated for
 the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
@@
 of those key-value store objects works identically.  Attributes are stored in a
-key-value store object.
-%, but use the
-%multi-dimensional array and blob objects to store the values for the
-%attributes.
+key-value store object, but use the
+multi-dimensional array and blob objects to store the values for the
+attributes.
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
 It is possible to request the transformation of an object's physical
-layout to other formats, having multiple copies of the same objects in
+layout to other formats resulting in multiple copies of the same objects in
 multiple formats if desired. Also, the user can pre-fetch objects from
@@
 \centering
-\vspace{-0.10in}
+%\vspace{-0.10in}
 \includegraphics[width=\columnwidth]{images/exa-arch.png}
-\vspace{-0.20in}
+%\vspace{-0.20in}
 \caption{Extreme Scale Architecture}
 \label{fig:exa-arch}
-\vspace{-0.10in}
+%\vspace{-0.10in}
 \end{figure}
@@
 
-\noindent\textbf{Design Philosophy}
+\subsection{Data Versioning}
+
+Since space is limited in expensive, in compute area storage resources, a
+copy-on-write approach is used for new versions of the same data. For example,
+for a checkpoint/restart file, multiple versions will be written. The only
+parts of this container that must be replicated are those that have changed
+since the last write. With potentially many transactions written to the IOD
+layer because it is fast, this approach will enable additional output to be
+stored while reducing the space overhead. The inherent dependencies this
+introduces into the data are a lesser issue for the generally transient data
+in the IOD layer. For data intended for persistence in the DAOS layer, it may
+expose all versions of the data to corruption unnecessarily. This is explored
+in more detail in the Section~\ref{sec:daos}.
+
+\subsection{Design Philosophy}
 
@@
 spread the IOD layer from the IO nodes into the compute area as well.  This is
-intended both to help address the limitations of the IO bandwidth and compute
-capability of these few nodes for data processing, but also to take advantage
+intended to help address the limitations of the IO bandwidth and compute
+capability of these few nodes for data processing and also to take advantage
 of new layers in the storage hierarchy. By incorporating NVRAM into compute
@@
 local data within the IOD layer rather than against the globally shared DAOS
-layer. With the additional desire to support using compute node resources for
-these operations, serious work will be required to make a fully functional
+layer that will likely still be disk for at least a couple more generations of
+platforms. With the additional desire to support using compute node resources
+for these operations, serious work will be required to make a fully functional
 end-to-end IOD layer implementation for a production system.
@@
 version of the container. For simplicity the epoch ID is the same as the
-transaction ID that was persisted.
-
-The process of persisting a transaction to an epoch involves a process called
-``flattening''. Since this stack uses a copy-on-write approach to reduce the
-space requirement for new versions of existing files, when a transaction is
-persisted to a DAOS epoch, all of the changes between the last epoch and the
-current epoch must be combined into a single entry. The process of combining
-the changes from a series of transactions into a single update for an epoch
-is termed flattening.
+transaction ID that was persisted. Unlike transaction IDs, epoch IDs generally
+are not consecutive reflecting that not all transactions will be persisted to
+DAOS.
+
+To deal with the potentially missing data versions between epochs because not
+all transactions are persisted, a special procedure must be followed. The
+``flattening'' process combines multiple copy-on-write versions of a
+transaction into a single epoch.  Since this stack uses a copy-on-write
+approach to reduce the space requirement for new versions of existing files all
+of the changes between the last epoch and the current epoch must be combined
+into a single entry. While not a cost free operation, it is generally
+considered inexpensive since a backwards combining of transaction blocks can
+be made ignoring any block that is already part of the combined changes.
 
@@
 symbolic links to all of the shards it contains and maintains the epoch ID.
-In particular the Highest Committed Epoch (HCE) is an important concept for
+In particular the Highest Committed Epoch is an important concept for
 quickly identifying which version of a shard to retrieve and to block writes to
@@
 does not support multiple applications from the same or different platforms
-using a shared DAOS layer to write to the same container at the same time.
+using a shared DAOS layer to write to the same container at the same time. This
+functionality is not supported by popular existing paralell file systems
+either.
+
+\subsection{Design Philosophy}
+
+The DAOS layer is the key storage management layer for this system. By handling
+the translation between user-level concepts and the underlying hardware,
+performance and functionality are both important. The choice of an object
+interface is influenced by the performance gains achieved by the data analytics
+community for non-shared data access. With the system design favoring requiring
+this operation mode, using an object interface fits naturally. With the broad
+array of object-based storage devices hitting the market, this layer may thin
+outsourcing much of the object creation and management to these speciailzed
+devices.
+
+Since this is the layer at which a storage system will be shared by multiple
+platforms, consistency is also a concern. By shifting to an object model and
+moving away from a POSIX-style directory tree, maintaining consistency will be
+easier. No longer will a consistent view of a particular set of files
+(containers) be required. Instead, only a single container need be consistent.
+With container sharing between platforms generally being limited to downstream
+analysis routines, waiting for a new epoch to be persisted can serve as an
+analysis trigger.
+
+Issues related to the handling of transactions and epochs are discussed in
+Section~\ref{sec:transactions}. Maintaining storage system scalability with
+this functionality will be challenging.
 
@@
 user API level down into the VOSD layer. While each layer affects the concept,
-it is best to look at it across all of the layers.
+it is best to look at each concept across all of the layers.
 
@@
 used to judge whether or not a set of distributed, asynchronous modifications
-across a set of related objects is complete or not.  It is also used to control
-access by treating the transaction ID of committed transaction as a version
-identifier.  At the DAOS layer and below, they are called epochs and represent
-persisted (durable) transactions from the IOD layer. Each of these offers
-different functionality, but are connected as is explained below.
+across a set of related objects (i.e., within a container) is complete or not.
+It is also used to control access by treating the transaction ID of committed
+transaction as a version identifier.  At the DAOS layer and below, they are
+called epochs and represent persisted (durable) transactions from the IOD
+layer. Each of these offers different functionality, but are connected as is
+explained below.
 
@@
 to the objects within a container are treated atomically. Conceptually,
-containers corresponds to a something akin to an HDF5 file in a traditional
+containers correspond to a something akin to an HDF5 file in a traditional
 file system. The objects in each container represent different data within a
@@
 a single client. The underlying assumption is that the client side will manage
-the transactional operations itself and the single client is capable of
-reporting to the IOD how to evolve the transaction state.
+the transactional operations itself and the single client is capable of and
+responsible for reporting to the IOD how to evolve the transaction state.
 
@@
 
-\noindent\textbf{Design Philosophy}
-
-Undocumented, but inherent in the design of these transactions is how faults
-are detected. The initial design assumes the current Lustre fault detection
-mechanism that can determine if a process or node is no longer reachable. This
-detection happens at the DAOS layer and when a fault is detected, the rollback
-process is pushed up to the IOD layer for all non-persisted or non-committed
-transactions. This defines how a fault will be detected and what will trigger a
-passive fault recovery (i.e., transaction abort).
-
-There are two steps for beginning a transaction on a container. The first step
-is for one or more process to open the container. This handle can be shared
-eliminating the need for every participating process to hit the IOD layer to
-open the file. The second step is a call to determine how many leaders will
-participate in the transaction. In the single leader case, there is no
-aggregation of success/fail statuses to determine the final transaction state.
-Instead, it is assumed that the client will fully manage the transaction. In
-the multi-leader model, some subset from 2 to $n$ where $n$ is the count of all
-processes, declare themselves a leader for this container operation to the IOD
-layer. Any number of processes can participate in modifying container without
-regard to whether or not they are a leader. Once each leader has finished, with
-the assumption that any clients they may be responsible for are finished as
-well, the IOD layer aggregates those responses to either commit or abort the
-transaction.
-
-Ultimately, with the passive detection of faults for transaction leaders, the
-transaction mechanism can work very well. A mostly unstated restriction that is
-being relaxed is that every sequential transaction on a container is considered
-dependent on the earlier transaction. Should one output be delayed and the
-subsequent five succeed, when the delayed process finally fails, all six
-transactions are rolled back. The thought of using this mechanism to store
-subsequent checkpoint outputs in the same container to both save space, but not
-care if one fails, cannot work in the current form. This has been acknowledged
-and is being relaxed requiring a new parameter to the creation of a transaction
-determining if it will be dependent or not.
-
 \subsubsection{Epochs}
+
 The Epoch mechanism differs from transactions. Instead of focusing on when a
@@
 
+\subsubsection{Design Philosophy}
+
+Undocumented, but inherent in the design of these transactions is how faults
+are detected. The initial design assumes the current Lustre fault detection
+mechanism that can determine if a process or node is no longer reachable. This
+detection happens at the DAOS layer and when a fault is detected, the rollback
+process is pushed up to the IOD layer for all non-persisted or non-committed
+transactions. This defines how a fault will be detected and what will trigger a
+passive fault recovery (i.e., transaction abort). The challenge with this
+approach will be scalability. Existing Lustre systems can use the IO node
+status as a proxy for compute area status. Since the DAOS layer must now know
+the state of every node, if not every process on every node, to properly
+handle transactions, some scalable status tracking mechanism is required.
+
+There are two steps for beginning a transaction on a container. The first step
+is for one or more process to open the container. This handle can be shared
+eliminating the need for every participating process to hit the IOD layer to
+open the file. The second step is a call to determine how many leaders will
+participate in the transaction. In the single leader case, there is no
+IOD-side aggregation of success/fail statuses to determine the final
+transaction state.  Instead, it is assumed that the client will fully manage
+the transaction. In the multi-leader model, some subset from 2 to $n$ where $n$
+is the count of all processes, declare themselves a leader for this container
+operation to the IOD layer. Any number of processes can participate in
+modifying container without regard to whether or not they are a leader. Once
+each leader has finished, with the assumption that any clients a leader may be
+responsible for are finished as well, the IOD layer aggregates those responses
+to either commit or abort the transaction. For scalability and performance,
+phase two is favoring single leader transactions. With libraries like
+D\textsuperscript{2}T~\cite{lofstead:2012:txn,lofstead:2013:pdsw-txn,lofstead:2014:txn}
+to ease implementing client-side transactions, this burden is lessened.
+
+Ultimately, with the passive detection of faults for transaction leaders, the
+transaction mechanism can work very well. A mostly unstated restriction that is
+being relaxed for phase two is that every sequential transaction on a container
+is considered dependent on the earlier transaction. Should one output be
+delayed and the subsequent five succeed, when the delayed process finally
+fails, all six transactions are rolled back. The thought of using this
+mechanism to store subsequent checkpoint outputs in the same container to both
+save space, but not care if one fails, cannot work in the current form. This
+has been acknowledged and is being relaxed requiring a new parameter to the
+creation of a transaction determining if it will be dependent or not. The
+downside to supporting this functionality is the reduced ability to use
+copy-on-write to reduce space pressures. If a transaction is allowed to fail
+and not affect subsequent transactions, the beginning state for subsequent
+transactions must the committed version prior to the current transaction. For
+example, if transaction ID 5 is marked as independent and 4 was previously
+committed, then transaction ID 6 will have to use version 4 as the base for
+copy-on-write.  A more in depth discussion of how transactions and checkpoint
+restart work is presented in BAD Check~\cite{bent:2015:pdsw}.
+
 \subsection{Metadata Management}
@@
 
-\noindent\textbf{Design Philosophy}
+\subsubsection{Design Philosophy}
 
@@
 this is introduced by the IOD, DAOS, and VOSD layers collectively. Each of
-these requires some different metadata storage and the migration is transparent
+these requires some different metadata storage and the migration is invisible
 to the user.  Supporting fully independent metadata with this model is
@@
 will be re-evaluated for phase two to provide more dynamic functionality.
+
+\subsubsection{Design Philosophy}
+
+The advantages of in transit and in situ processing is well documented. For
+example, PreDatA~\cite{zheng:2010:predata} evaluates different operator
+placement decisions and when each is advantageous. Inherent in this approach is
+the need to ship arbitrary code to run in various locations. Security
+considerations aside, something like C-on-demand~\cite{abbasi:2011:c-on-demand}
+or some other dynamic code generation and execution system will be required.
+Lightweight containers~\cite{merkel:2014:docker} could also be used, but with
+potentially higher management overheads.
+
+\begin{figure}[htbp]
+\centering
+%\vspace{-0.10in}
+\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
+%\vspace{-0.30in}
+\caption{GraphLab and GraphBuilder stacks}
+\label{fig:graphlab-graphbuilder}
+%\vspace{-0.10in}
+\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\vspace{-0.10in}
-\includegraphics[width=\columnwidth]{images/graphlab-and-graphbuilder.png}
-\vspace{-0.30in}
-\caption{GraphLab and GraphBuilder stacks}
-\label{fig:graphlab-graphbuilder}
-\vspace{-0.10in}
-\end{figure}
-
 In order to make both of these tools work on top of the extreme scale stack,
@@
 
+\begin{figure*}[htbp!]
+\centering
+%\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+%\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+%\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+%\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+%\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+%\vspace{-0.05in}
+\end{figure*}
+
 \section{Demonstration}
 \label{sec:evaluation}
-
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
-\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Number of Hosts}
-\label{fig:eval-hosts}
-\vspace{-0.05in}
-\end{figure*}
-
-\begin{figure*}[htbp!]
-\centering
-\vspace{-0.10in}
-\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
-\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
-\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Data Sizes}
-\label{fig:eval-size}
-\vspace{-0.05in}
-\end{figure*}
 
@@
 
-All of the tests are performed on the Buffy Cray XC30-AC at LANL.  It consists
+All of the tests are performed on
+a blinded for review machine.
+%the Buffy Cray XC30-AC at LANL.
+It consists
 of 64 compute nodes each with dual, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz
@@
 GB of RAM.  There are also a metadata, login, and 2 boot nodes.  The storage
-array has disk and SSD paritions. One is a DDN Lustre system with 192 TB disk
+array has disk and SSD partitions. One is a DDN Lustre system with 192 TB disk
 usable with a minimum of 5 GB/sec performance. The rest are SSDs consisting of
@@
 Figure~\ref{fig:eval-hosts} show the reading and writing performance for
-different number of hosts. Each read or write is 4 GB against the x-axis number
-of hosts. The second set in Figure~\ref{fig:eval-size} show the performance of
-reading and writing different sizes for 56 clients, the smallest client count
-when performance stabilizes in the number of hosts tests.  The performance of
-both of these tests are reported to give a very rough idea of the overhead that
-might be involved. Rather than a true overhead, this should be considered the
-maximum overhead that should be expected once an optimized, fully functional IO
-stack is deployed without relying on translating to an underlying parallel file
-system.
+different numbers of hosts. Each read or write is 4 GB against the x-axis
+number of hosts. The second set in Figure~\ref{fig:eval-size} show the
+performance of reading and writing different sizes for 56 clients, the smallest
+client count when performance stabilizes in the number of hosts tests.  The
+performance of both of these tests are reported to give a very rough idea of
+the overhead that might be involved. Rather than a true overhead, this should
+be considered the maximum overhead that should be expected once an optimized,
+fully functional IO stack is deployed without relying on translating to an
+underlying parallel file system.
 
@@
 
-\includegraphics[scale=0.07]{logos/doe_logo}
-\includegraphics[scale=0.30]{logos/snl_logo}
-\includegraphics[scale=0.35]{logos/nnsa_logo}
+%\includegraphics[scale=0.07]{logos/doe_logo}
+%\includegraphics[scale=0.30]{logos/snl_logo}
+%\includegraphics[scale=0.35]{logos/nnsa_logo}
 Sandia National Laboratories is a multi-program laboratory managed and operated

### 2016-07-18 paper_submitted.tex
added reviews and "plagarism" concerns (fixed via refs). Need to edit a bit more still.

--- 2016-07-18 paper.tex
@@
 \author{
-%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
-\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Lawrence Berkeley National Laboratory koziol@lbl.gov}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Segate Government Solutions John.Bent@seagategov.com}
-\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
+\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
+%\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
 }
@@
 how different concepts will work across layers since that information is spread
-across multiple documents and may lack a cohesive overall view. Previously,
-a poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
-consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
-been presented.
+across multiple documents and may lack a cohesive overall view.
 
@@
 The main alternative from scratch design for a file system is
-Sirocco~\cite{sirocco,curry:2016:sirocco}. Rather than continuing the striped
-design of existing parallel file systems, Sirocco is inspired by peer-to-peer
-and object-based systems and includes features like transactions to protect
-data modification process independence when writing to avoid coordination
-overhead.  The base assumptions are that storage is pervasive and volatile.
-Storage devices and locations may come and go randomly, reminiscent of the
-Google or Ceph assumptions of regularly failing hardware. When data is pushed
-into the system, initial resilience characteristics are guaranteed prior to
-returning control back to the user. Then, as system pressures dictate, data
-will either replicate as demanded by use and/or migrated towards long-term
-resilience requirements.  Unlike the FFSIO project, Sirocco assumes it is
-possible that data may be successfully stored in the system, but it is
-currently inaccessible because all copies are currently offline. There is also
-some potential difficulty in finding data since it will migrate around the
-system. To be fair, Sirocco intends to function as the storage layer for a
-higher level file system API making many of the awkward system features
-invisible to the end user. Sirocco is in the process of being released
-publicly.
+Sirocco~\cite{sirocco}. Rather than continuing the striped design of existing
+parallel file systems, Sirocco is inspired by peer-to-peer and object-based
+systems and includes features like transactions to protect data modification
+process independence when writing to avoid coordination overhead.  The base
+assumptions are that storage is pervasive and volatile. Storage devices and
+locations may come and go randomly, reminiscent of the Google or Ceph
+assumptions of regularly failing hardware. When data is pushed into the system,
+initial resilience characteristics are guaranteed prior to returning control
+back to the user. Then, as system pressures dictate, data will either replicate
+as demanded by use and/or migrated towards long-term resilience requirements.
+Unlike the FFSIO project, Sirocco assumes it is possible that data may be
+successfully stored in the system, but it is currently inaccessible because all
+copies are currently offline. There is also some potential difficulty in
+finding data since it will migrate around the system. To be fair, Sirocco
+intends to function as the storage layer for a higher level file system API
+making many of the awkward system features invisible to the end user. Sirocco
+is in the process of being released publicly.
 
@@
 
+blinded for review
+
 %\includegraphics[scale=0.07]{logos/doe_logo}
@@
 %\includegraphics[scale=0.35]{logos/nnsa_logo}
-Sandia National Laboratories is a multi-program laboratory managed and operated
-by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
-Corporation, for the U.S. Department of Energy's National Nuclear Security
-Administration under contract DE-AC04-94AL85000.
+%Sandia National Laboratories is a multi-program laboratory managed and operated
+%by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
+%Corporation, for the U.S. Department of Energy's National Nuclear Security
+%Administration under contract DE-AC04-94AL85000.
 

### 2016-07-18 paper.tex

@@
 
-%\title{Fast Forwarding Storage for Extreme Scale Platforms and Big Data Applications}
-\title{Lessons Learned from the Fast Forward Storage and I/O Project}
+\title{DAOS and Friends: A Proposal for\\an Exascale Storage System}
 
 \author{
-%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
 \IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
@@
 
-To address these challenges, a joint effort between the US Department of
+%To address these challenges,
+A joint effort between the US Department of
 Energy's Office of Advanced Simulation and Computing and Advanced Scientific
@@
   asynchronous operations, creating and managing transactions, end-to-end data
-  integrity, and data type and functions to support the ACG functionality more
+  integrity, and data type and functions to support the big data oriented functionality more
   efficiently than the current API.
@@
 %    \vspace{-0.15in}
-
     \begin{tabular}{|r|r|}
@@
   fashion across the IO nodes.
-\item
-  \textbf{chunked}. same as above but with irregular (sparse) chunking.
-\item
-  \textbf{user-defined}. either contiguous or chunked, but user specifies
+%\item
+  \textbf{(2)chunked}. same as above but with irregular (sparse) chunking.
+%\item
+  \textbf{(3)user-defined}. either contiguous or chunked, but user specifies
   where to place each individual shard.
@@
 In the subsections that follow, we examine transactions and epochs, metadata
-management, function and analysis shipping, and the arbitrary connected graphs
-support.
+management, and function and analysis shipping.
+
+\begin{figure*}[htbp!]
+\centering
+%\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+%\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+%\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+%\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+%\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+%\vspace{-0.2in}
+\end{figure*}
 
@@
 the reference count is reduced. Once the count reaches 0, the transaction is
-automatically committed.
+automatically committed. Aggregation into a smaller set of leaders is also
+possible.
 
@@
 \section{Acknowledgements}
-
-%\includegraphics[scale=0.07]{logos/doe_logo}
-%\includegraphics[scale=0.30]{logos/snl_logo}
-%\includegraphics[scale=0.35]{logos/nnsa_logo}
 Sandia National Laboratories is a multi-program laboratory managed and operated

### 2016-07-26 paper.tex
added reviews and completed final version edits.

--- 2016-07-18 paper.tex
@@
 \pdfpageheight=11in
-%\usepackage{subfig}
 \usepackage{subfigure}
@@
 \usepackage{graphicx}
-%\usepackage{todonotes}
-\usepackage{listings}
 \usepackage{hyperref}
-\lstset {% general command to set parameter(s)
-         language=C,
-     basicstyle=\footnotesize,               % print whole listing small
-%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
-%     identifierstyle =\color{black},  % nothing happens
-%     commentstyle=\color{black}\emph, % white comments
-     %stringstyle=\ttfamily,          % typewriter type for strings
-%     stringstyle=\color{black},       % typewriter type for strings
-         tabsize=4,
-         showtabs=false,
-     showstringspaces=false}
-%can't figure this one out for particles bandwidth
-%\usepackage[caption,label]{subfig}
-
-\newcommand{\DDT}{D\textsuperscript{2}T~}
-\newcommand{\DDTns}{D\textsuperscript{2}T}
 
@@
 
+\IEEEoverridecommandlockouts
+\IEEEpubid{\makebox[\columnwidth]{978-1-4673-8815-3/16/\$31.00 ©2016 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 \begin{document}
 
-%\title{Fast Forwarding Storage for Extreme Scale Platforms and Big Data Applications}
-\title{Lessons Learned from the Fast Forward Storage and I/O Project}
+\title{DAOS and Friends: A Proposal for\\an Exascale Storage System}
 
 \author{
-%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
 \IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
@@
 \begin{abstract}
-%The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
-%project is going to have significant impact on storage systems design within
-%and beyond the HPC community. With phase two of the project starting, it is an
-%excellent opportunity to explore the complete design and how it will address
-%the needs of extreme scale platforms.  This paper examines the overall design.
-%Each layer of the proposed stack is discussed in some detail along with
-%cross-cutting topics that impact multiple layers, such as transactions and
-%metadata management. We also describe the proposed features to better support
-%some classes of big data applications.
-%
-%This paper not only provides a timely summary of important aspects of the
-%design specifications but also captures the underlying reasoning that is not
-%available elsewhere. With this information, we encourage the broader community
-%to understand the design, intent, and future directions to foster discussion
-%guiding phase two and the ultimate production storage stack based on this work.
-%An initial performance evaluation of the early prototype implementation is also
-%provided to validate the presented design.
-
 The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
@@
 
-To address these challenges, a joint effort between the US Department of
+A joint effort between the US Department of
 Energy's Office of Advanced Simulation and Computing and Advanced Scientific
@@
 identified challenges. The first phase completed in 2014 with a second phase
-currently getting underway. The core focus of the first phase was basic
+underway. The first phase focused primarily on basic
 functionality and design. While an idealized potential system would be the
@@
 compute nodes is currently not economically feasible limiting some of the
-potential design choices.  With this in mind, the second phase will refine this
-design incorporating fault recovery and other features missing from the first
-phase.
+potential design choices.  With this in mind, the second phase is 
+incorporating fault recovery and other missing features.
 
 The complete design seeks to offer high availability, byte-granular,
-multi-version concurrency control. Through the use of a copy-on-write style
-mechanism reducing storage space and writing time requirements, multiple
-versions of an object can be stored efficiently. By assuming the client
+multi-version concurrency control. Multiple versions of an object are stored
+efficiently by using a copy-on-write mechanism. By assuming the client
 interface will be through an IO library, a more complicated interface offering
@@
 containers of objects. This shift avoids the bottlenecks related to the POSIX
-files and directories structure such as the serialization of file creation, the
-limitation and impact of the number of files in a directory, and the limited
+files and directories structure such as file creation serialization, the
+file count in a directory limitation and impact, and the limited
 semantics of a byte stream. Instead, the new interface focuses on high-level
@@
 these computing models in addition to the typically bursty IO loads of more
-traditional HPC applications.
+traditional HPC applications. These features are not discussed for space
+reasons.
 
@@
 
-At a more detailed view, the various layers of the IO stack each contribute
+The IO stack layers each contribute
 different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
@@
 The top layer is comprised of generally a high level IO library, such as the
-demonstration HDF5 library~\cite{hdf5} and a more complex API for accessing the
+demonstration HDF5 library~\cite{folk:2011:hdf5} and a more complex API for accessing the
 lower level components. This layer is in dark blue. Becuase the system supports
@@
 
-The Distributed Asynchronous Object Storage (DAOS) layer serves as the
+The Distributed Application Object Storage (DAOS) layer serves as the
 persistent storage interface and translation layer between the user-visible
@@
 
-Along with analysis of the published design documents, a discussion of the
-design philosophy representing the overall intent is presented. This
-information represents information that may or may not have been written down,
-but is the intent of ultimate system.  This information was gathered through
-interviews with the core FFSIO team members. These ideas are presented to give
-a fuller picture of where the project is going rather than dwelling on any
-limitations of the published documents. This is most important to illustrate
-how different concepts will work across layers since that information is spread
-across multiple documents and may lack a cohesive overall view. Previously,
-a poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
+This paper presents an analysis of the published design documents along with a
+discussion of the design philosophy representing the overall intent. The design
+philospohy conveys information that may or may not have been written down, but
+aids understanding for the total design.  This information was gathered through
+interviews with the core FFSIO team members. These ideas are presented to
+reveal the project future rather than dwelling on any limitations of the
+published designs. This is most important to illustrate how different concepts
+will work across layers since that information is spread across multiple
+documents and may lack a cohesive overall view. Previously, a
+poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
 consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
-been presented.
+been published.
+
+With DAOS and friends being groomed as the next generation for the Lustre
+parallel file system, the information and analysis presented here can help
+users determine how to adapt their thinking about storage as well as hopefully
+influence the second phase.
+
+The key contributions of this paper are the syntesized architectural overview
+and the discussion and analysis motivating the architectural descisions. With
+this information, community members can be better informed about an important
+future storage system as well as potentially influence the evolution of the
+design from prototype into production system.
 
@@
 the storage array.  This will be discussed in the context of the HDF5 based
-example library used for the functionality demonstration. Section~\ref{sec:iof}
-briefly discusses the motivation and proposal for the IO forwarding layer.
-Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
-functionality it offers. This will detail the pieces of the layer that are
-potentially optional and mention the cross-cutting features discussed in a
-later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
-layer functions. As with the IOD layer, the cross-cutting features will be
-mentioned, but discussed more fully in the cross-cutting section. The VOSD
+example library used for the functionality demonstration.
+Section~\ref{sec:iof} briefly discusses the motivation and proposal for the IO
+forwarding layer.  Section~\ref{sec:iod} describes the IO Dispatcher layer and
+the broad functionality it offers.  This will detail the pieces of the layer
+that are potentially optional and mention the cross-cutting features discussed
+in a later, cross-cutting section.  Section~\ref{sec:daos} discusses how the
+DAOS layer functions.  As with the IOD layer, the cross-cutting features will
+be mentioned, but discussed more fully in the cross-cutting section.  The VOSD
 layer is discussed in Section~\ref{sec:vosd}.  In particular, the mapping
 between the DAOS and VOSD layers are explored as it pertains to the physical
-storage. Next is an exploration of cross-cutting features like transactions and
-metadata management in Section~\ref{sec:broader}. Since these and other
+storage.  Next is an exploration of cross-cutting features like transactions
+and metadata management in Section~\ref{sec:broader}.  Since these and other
 features are spread across multiple layers, it makes more sense to discuss them
@@
 presented.  A demonstration of the functionality is presented in
-Section~\ref{sec:evaluation}. This shows that the prototype system based on the
-proposed design can function. The paper is concluded in
-Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
-paper.
+Section~\ref{sec:evaluation}.  This shows that the prototype system based on
+the proposed design can function.  Section~\ref{sec:conclusion} concludes the
+paper with a summary of the broad issues.
 
@@
 numbers with a known seed eliminating the need for the metadata service to
-track where each stripe in a parallel file is placed. Ceph's strengths are in
+track where each piece in a striped file is placed. Ceph's strengths are in
 providing good perforamnce and scalability with the ability to handle failures
@@
   asynchronous operations, creating and managing transactions, end-to-end data
-  integrity, and data type and functions to support the ACG functionality more
-  efficiently than the current API.
+  integrity, and data type and functions to support the big data oriented
+  functionality more efficiently than the current API.
 
@@
 
-HDF5~\cite{hdf5} has a versatile data model offering complex data objects
+HDF5~\cite{folk:2011:hdf5} has a versatile data model offering complex data objects
 and metadata. Its information set is a collection of datasets, groups,
@@
 The Virtual Object Layer is an abstraction mechanism internal to the HDF5
-library~\cite{hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
+library~\cite{folk:2011:hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
 just below the public API. The VOL exports an interface that allows writing
@@
 offers without requiring all of the functionality be exposed to users. For
-example, the containers and objects concept is transparently mapped to the
-files and datasets existing HDF5 users are familiar with.  This reduces the
-difficulty porting applications to the new IO stack.
+example, the containers and objects concept is mapped to the files and datasets
+existing HDF5 users are familiar with.  This reduces the difficulty porting
+applications to the new IO stack.
 
@@
 
-Since HDF5 has traditionally offered an interface focused on files and the
-internal data types, such as datasets, these concepts must be mapped onto the
-proposed FFSIO data storage concepts.  This mapping is shown in
-Table~\ref{tab:mapping}
+Since HDF5~\cite{folk:2011:hdf5} offers an interface focused on files and the internal
+data types, such as datasets, these concepts must be mapped onto the proposed
+FFSIO data storage concepts.  This mapping is shown in Table~\ref{tab:mapping}
 
@@
 
-In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.
+In Section~\ref{sec:iod}, the FFSIO types are described in more detail.
 
@@
 
-The core idea for IOD is to provide a way to manage IO load that is
-separate from the compute nodes and the storage array. Communication intensive
-activities, such as data rearrangement, can be moved to the IOD layer
-reducing the number of participants and message count. IOD has three main
-purposes. First, the burst buffers work as a fast cache absorbing write
-operations that then trickles out to the central storage array. It can also
-be used to retrieve objects from the central storage array for more efficient
-read operations and offers data filtering to make client reads more efficient.
-Second, it offers the transaction mechanism for controlling data set visibility
-and to manage faults that could expose an incomplete or corrupt data set to
-users. These transactions are local to the IOD layer until persisted to the
-DAOS layer eliminating the need for burdening the persistent storage with
-transient data.  Third, data processing operations can be placed in the IOD.
-These operations are intended to offer functionality like data rearrangement
-and filtering prior to data reaching the central storage array.
+The core idea for IOD is to provide a way to manage IO load that is separate
+from the compute nodes and the storage array. Communication intensive
+activities, such as data rearrangement, can be moved to the IOD layer reducing
+the number of participants and message count. The IOD has three main purposes.
+First, the burst buffers work as a fast cache absorbing write operations that
+then trickles out to the central storage array or pre-staging read operations.
+It can also be used to retrieve objects from the central storage array for more
+efficient read operations and offers data filtering to make client reads more
+efficient.  Second, it offers the transaction mechanism for controlling data
+set visibility and to manage faults that could expose an incomplete or corrupt
+data set to users.  These transactions are local to the IOD layer until
+persisted to the DAOS layer eliminating the need for burdening the persistent
+storage with transient data.  Third, data processing operations can be placed
+in the IOD.  These operations are intended to offer functionality like data
+rearrangement and filtering prior to data reaching the central storage array.
 
@@
 
-\subsubsection{Container}
+\textbf{Container}
 As mentioned above, the concept of a container is similar to that of a file in
@@
 each container essentially is stored in a hash-space allowing direct access to
-any container without regard to the current organizational context of the
-file system. For example, there is no need to navigate a directory hierarchy
-to name a particular container.
+any container without regard to the current organizational context of the file
+system.  For example, there is no need to navigate a directory hierarchy to
+name a particular container.
 
@@
 of those key-value store objects works identically.  Attributes are stored in a
-key-value store object, but use the
-multi-dimensional array and blob objects to store the values for the
-attributes.
-
-\subsubsection{Multi-Dimensional Array}
+key-value store object, but use the multi-dimensional array and blob objects to
+store the values for the attributes.
+
+\textbf{Multi-Dimensional Array}
 By treating arrays as a special case separate from blobs, additional
@@
 performance of a subsequent data analytics task. A common scenario where this
-is common is a Fortran code (column-major) writes data for a C code (row-major)
+is useful is a Fortran code (column-major) writes data for a C code (row-major)
 to analyze.  The IOD API supports the following sharding strategies:
@@
 \item
-  \textbf{contiguous}. fixed chunking, distributed in a round-robin
+  \textbf{contiguous}. Fixed chunking, distributed in a round-robin
   fashion across the IO nodes.
 \item
-  \textbf{chunked}. same as above but with irregular (sparse) chunking.
-\item
-  \textbf{user-defined}. either contiguous or chunked, but user specifies
+  \textbf{chunked}. Same as above but with irregular (sparse) chunking.
+\item
+  \textbf{user-defined}. Either contiguous or chunked, but user specifies
   where to place each individual shard.
@@
 layout to other formats resulting in multiple copies of the same objects in
-multiple formats if desired. Also, the user can pre-fetch objects from
-the storage cluster into the IO nodes or read them directly from the
-storage cluster. At the semantic level (HDF5), indices can be created
-for datasets resulting in being able to read through an index instead of
-directly from the base array.
+multiple formats. Also, the user can pre-fetch objects from the storage cluster
+into the IO nodes or read them directly from the storage cluster. At the
+semantic level (HDF5), indices can be created for datasets resulting in being
+able to read through an index instead of directly from the base array.
 
@@
 
-The Distributed Asynchronous Object Storage layer serves as the traditional
-parallel file system interface layer for the storage devices. This is the
-consistent, global view of the underlying devices represented in this stack
-by the VOSD layer.
-
-This is the layer where the container/object model is translated into the
-physical storage requirements dictated by the physical storage underneath (the
-VOSD layer). The two key design elements of this layer are the handling of
-epochs and the mapping of containers and objects to the underlying storage.
+The Distributed Application Object Storage (DAOS) layer serves as the
+traditional parallel file system interface layer for the storage devices. This
+is the consistent, global view of the underlying devices represented in this
+stack by the VOSD layer.  This is the layer where the container/object model is
+translated into the physical storage requirements dictated by the physical
+storage underneath (the VOSD layer). The two key design elements of this layer
+are the handling of epochs and the mapping of containers and objects to the
+underlying storage.
 
@@
 In the subsections that follow, we examine transactions and epochs, metadata
-management, function and analysis shipping, and the arbitrary connected graphs
-support.
+management, and function and analysis shipping.
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
+\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Number of Hosts}
+\label{fig:eval-hosts}
+\vspace{-0.05in}
+\end{figure*}
+
+\begin{figure*}[htbp!]
+\centering
+\vspace{-0.10in}
+\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
+\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
+\vspace{-0.10in}
+\caption{Functionality Demonstration Validation for Data Sizes}
+\label{fig:eval-size}
+\vspace{-0.2in}
+\end{figure*}
 
@@
 \label{sec:transactions}
+
+In our previous work~\cite{lofstead:2014:ffsio-consistency}, we focused
+strongly on not just exploring, but also critiquing using transactions, as
+proposed, for consistency control. This section recaps some of that information
+to give a more self-contained presentation.
 
@@
 treating it just as data is a central design goal of the Fast Forward project.
-This is a hybrid approach to metadata management that is halfway between
+FFSIO uses a hybrid approach to metadata management that is halfway between
 providing no inherent metadata support and having a fully integrated, but
@@
 
-\begin{figure*}[htbp!]
-\centering
-%\vspace{-0.10in}
-\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
-\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
-%\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Number of Hosts}
-\label{fig:eval-hosts}
-%\vspace{-0.05in}
-\end{figure*}
-
-\begin{figure*}[htbp!]
-\centering
-%\vspace{-0.10in}
-\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
-\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
-%\vspace{-0.10in}
-\caption{Functionality Demonstration Validation for Data Sizes}
-\label{fig:eval-size}
-%\vspace{-0.05in}
-\end{figure*}
-
 \section{Demonstration}
@@
 
-All of the tests are performed on
-a blinded for review machine.
-%the Buffy Cray XC30-AC at LANL.
-It consists
+All of the tests are performed on the Buffy Cray XC30-AC at LANL.  It consists
 of 64 compute nodes each with dual, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz
@@
 such as fault detection and recovery, can be designed and tested.  These and
-other activities for phase 2 will ultimately generate what is likely to be the
-next generation storage stack for extreme scale platforms.
+other activities for phase two will ultimately generate what is likely to be
+the next generation storage stack for extreme scale platforms.
 
 \section{Acknowledgements}
-
-%\includegraphics[scale=0.07]{logos/doe_logo}
-%\includegraphics[scale=0.30]{logos/snl_logo}
-%\includegraphics[scale=0.35]{logos/nnsa_logo}
 Sandia National Laboratories is a multi-program laboratory managed and operated
